---
title: "Asembo enteric antibody analysis of age-dependent antibody curves"
output: 
  html_notebook:
    theme: default
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

#Summary
This script estimates age-dependent mean antibody curves using cubic splines (in a GAM model from the `mgcv` package) and using ensemble machine learning (using super learning from the `tmleAb` package). 

# Script preamble
```{r preamble}
#-----------------------------
# preamble
#-----------------------------
rm(list=ls())
library(tidyverse)
library(mixtools)

# set up for parallel computing
# configure for a laptop (use only 3 cores)
library(foreach)
library(doParallel)
registerDoParallel(cores=3)

# bright color blind palette:  https://personal.sron.nl/~pault/ 
cblack <- "#000004FF"
cblue <- "#3366AA"
cteal <- "#11AA99"
cgreen <- "#66AA55"
cchartr <- "#CCCC55"
cmagent <- "#992288"
cred <- "#EE3333"
corange <- "#EEA722"
cyellow <- "#FFEE33"
cgrey <- "#777777"

```


#Load the formatted data

```{r load data}
#-----------------------------
# load the formatted data
# with seropositivity indicators
# created with:
# asembo-enteric-ab-data-format.Rmd -->
# asembo-enteric-ab-distributions.Rmd
#-----------------------------
dl <- readRDS("~/enterics-immuno-epi/data/asembo_analysis2.rds")

# list the enteric antigens and make formatted labels for them
mbavars <- c("vsp3","vsp5","cp17","cp23","leca","rotavirus","salb","sald","p18","p39","etec","cholera")

mbalabs <- c("Giardia VSP-3","Giardia VSP-5","Cryptosporidium Cp17","Cryptosporidium Cp23","E. histolytica LecA","Rotavirus","Salmonella LPS B","Salmonella LPS D","Campylobacter p18","Campylobacter p39","ETEC toxin beta subunit","Cholera toxin beta subunit")
```

#Estimate age-dependent mean antibody levels

##Estimate curves using splines in a generalized additive model (GAM)

```{r age-curve mean gam }
#----------------------------------
# simulataneous CIs for GAMs
# estimated by resampling the 
# Baysian posterior estimates of
# the variance-covariance matrix
# assuming that it is multivariate normal
# the function below also estimates 
# the unconditional variance-covariance
# matrix, Vb=vcov(x,unconditional=TRUE), 
# which allows for undertainty in the actual
# estimated mean as well 
# (Marra & Wood 2012 Scandinavian Journal of Statistics, 
#  Vol. 39: 53–74, 2012, doi: 10.1111/j.1467-9469.2011.00760.x )
# simultaneous CIs provide much better coverage than pointwise CIs
# see: http://www.fromthebottomoftheheap.net/2016/12/15/simultaneous-interval-revisited/
#----------------------------------

gamCI <- function(m,newdata,nreps=10000) {
  require(mgcv)
  require(dplyr)
  Vb <- vcov(m,unconditional = TRUE)
  pred <- predict(m, newdata, se.fit = TRUE)
  fit <- pred$fit
  se.fit <- pred$se.fit
  BUdiff <- MASS::mvrnorm(n=nreps, mu = rep(0, nrow(Vb)), Sigma = Vb)
  Cg <- predict(m, newdata, type = "lpmatrix")
  simDev <- Cg %*% t(BUdiff)
  absDev <- abs(sweep(simDev, 1, se.fit, FUN = "/"))
  masd <- apply(absDev, 2L, max)
  crit <- quantile(masd, prob = 0.95, type = 8)
  pred <- data.frame(newdata,fit=pred$fit,se.fit=pred$se.fit)
  pred <- mutate(pred,
                 uprP = fit + (2 * se.fit),
                 lwrP = fit - (2 * se.fit),
                 uprS = fit + (crit * se.fit),
                 lwrS = fit - (crit * se.fit)
  )
  return(pred)
}


# fit cubic spline over age, separately for each antigen
# estimate approximate simultaneous confidence intervals
# for the fits
library(mgcv)
gamfits <- foreach(ab=levels(dl$antigenf),.combine=rbind) %dopar% {
    pd <- filter(dl, antigenf==ab)
    gfit <- gam(logmfi~s(age, bs="cr"),data=pd)
    gsci <- gamCI(m=gfit,newdata=pd,nreps=1000)
    gsci$antigen <- ab
    return(gsci)
  }
```

##Estimate curves using ensemble machine learning
```{r age-curve mean ensemble, echo=TRUE, message=FALSE, warning=FALSE}
#-----------------------------
# estimate curves with 
# ensemble machine learning
# including a library with
# GLM, GAM, Loess, and 
# extreme gradient boosting
#-----------------------------

# detach the mgcv package because
# the ensemble algorithm uses gam()
# from the gam package and there are
# conflicts between them
detach("package:mgcv", unload=TRUE)

# load the tmleAb package, at this time
# only available through github: 
# devtools::install_github("ben-arnold/tmleAb")
library(tmleAb)
set.seed(12345)
slfits <-  foreach(ab=levels(dl$antigenf),.combine=rbind) %dopar% {
  pd <- filter(dl, antigenf==ab)
  slfit <- agecurveAb(Y=pd$logmfi,
                      Age=pd$age,
                      id=pd$childid,
                      SL.library = c("SL.glm","SL.gam","SL.loess","SL.xgboost"))
  sld <- data.frame(age=slfit$Age,fit=slfit$pY)
  sld <- sld %>% group_by(age) %>% filter(row_number()==1) %>% ungroup()
  slres <- left_join(pd,sld,by=c("age"))
  return(slres)
}


```

##Plot age-dependent mean antibody curves

The spline fits are colored and the highly adaptive ensemble fits are grey. The approximate simultaneous confidence bands for the spline fit are shaded.

```{r age-curve mean figure, fig.height=12, fig.width=6}

log10labs <- c( 
  expression(10^0),
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)

# bright color-blind palette defined in an earlier code chunk
pcols <- c(cred,corange,cgreen,cgrey,cteal,cteal,cblue,cmagent)

pcurve <- ggplot(data=gamfits,aes(x=age,group=pathogen,color=pathogen,fill=pathogen)) +
  # approximate simultaneous CI from spline fit
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA,fill=cgrey) +
  # actual measurements
  geom_jitter(aes(y=logmfi),width=0.2,height=0,color="black",alpha=0.4,size=0.3) +
  # spline fit
  geom_line(aes(y=fit),color=cgrey) +
  # ensemble fit
  geom_line(data=slfits,aes(x=age,y=fit)) +
  # plot aesthetics
  facet_wrap(~antigenf,nrow=6,ncol=2) +
  scale_x_continuous(breaks=seq(4,18,by=2))+
  scale_y_continuous(breaks=0:4,labels=log10labs)+
  coord_cartesian(ylim=c(0,4.5))+
  labs(x="Age in months",y="Luminex Response (MFI-bg)") +
  scale_fill_manual(values=pcols) +
  scale_color_manual(values=pcols) +
  theme_minimal() +
  theme(legend.position = "none")

pcurve

# save a PDF version 
ggsave("~/enterics-immuno-epi/figs/asembo-ab-agecurves-mean.pdf",plot=pcurve,device="pdf",width=7,height=14)

```

#Estimate age-dependent seroprevalence

##Estimate seroprevalence curves using splines in a generalized additive model (GAM)

```{r age-curve seroprevalence gam }

#-----------------------------
# subset the data to antigens
# with seropositivity estimates
#-----------------------------

dlp <- dl %>%
  mutate(seropos=ifelse(!is.na(posroc),posroc,posmix)) %>%
  filter(!is.na(seropos)) %>%
  mutate(antigenf=factor(antigenf))


#-----------------------------
# estimate seroprevalence curves with 
# splines
#-----------------------------


# fit cubic spline over age, separately for each antigen
# estimate approximate simultaneous confidence intervals
# for the fits
library(mgcv)
pgamfits <- foreach(ab=levels(dlp$antigenf),.combine=rbind) %dopar% {
    pd <- filter(dlp, antigenf==ab)
    gfit <- mgcv::gam(seropos~s(age, bs="cr"),data=pd,family="binomial")
    gsci <- gamCI(m=gfit,newdata=pd,nreps=1000)
    gsci$antigenf <- ab
    return(gsci)
}

# use antilogit to get back to prevalence from the logit link
pgamfits <- pgamfits %>%
  mutate(fit  = 1/(1+exp(-fit)),
         uprP = 1/(1+exp(-uprP)),
         lwrP = 1/(1+exp(-lwrP)),
         uprS = 1/(1+exp(-uprS)),
         lwrS = 1/(1+exp(-lwrS))) %>%
  mutate(antigenf=factor(antigenf,levels=levels(dlp$antigenf)))
```

##Estimate seroprevalence curves using ensemble machine learning
```{r age-curve seroprevalence ensemble, echo=TRUE, message=FALSE, warning=FALSE}
#-----------------------------
# estimate curves with 
# ensemble machine learning
# including a library with
# GLM, GAM, Loess, and 
# extreme gradient boosting
#-----------------------------

# detach the mgcv package because
# the ensemble algorithm uses gam()
# from the gam package and there are
# conflicts between them
detach("package:mgcv", unload=TRUE)

set.seed(12345)
pslfits <-  foreach(ab=levels(dlp$antigenf),.combine=rbind) %dopar% {
  pd <- filter(dlp, antigenf==ab)
  slfit <- agecurveAb(Y=pd$seropos,
                      Age=pd$age,
                      id=pd$childid,
                      family="binomial",
                      SL.library = c("SL.glm","SL.gam","SL.xgboost"))
  sld <- data.frame(age=slfit$Age,fit=slfit$pY)
  sld <- sld %>% group_by(age) %>% filter(row_number()==1) %>% ungroup()
  slres <- left_join(pd,sld,by=c("age"))
  return(slres)
}

```

##Plot age-dependent seroprevalence  curves

The spline fits are colored and the highly adaptive ensemble fits are grey. The approximate simultaneous confidence bands for the spline fit are shaded.

```{r age-curve seroprevalence figure, fig.height=9, fig.width=6}


# bright color-blind palette defined in an earlier code chunk
# (remove cteal since there are no curves for ETEC or cholera)
pcols <- c(cred,corange,cgreen,cgrey,cblue,cmagent)

ppcurve <- ggplot(data=pgamfits,aes(x=age,group=pathogen,color=pathogen,fill=pathogen)) +
  # approximate simultaneous CI from spline fit
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA,fill=cgrey) +
  # actual measurements
  geom_jitter(aes(y=seropos),width=0.2,height=0,color="black",alpha=0.4,size=0.3) +
  # spline fit
  geom_line(aes(y=fit),color=cgrey) +
  # ensemble fit
  geom_line(data=pslfits,aes(x=age,y=fit)) +
  # plot aesthetics
  facet_wrap(~antigenf,nrow=5,ncol=2) +
  scale_x_continuous(breaks=seq(4,18,by=2))+
  scale_y_continuous(breaks=seq(0,1,by=0.2),labels=sprintf("%1.0f",seq(0,1,by=0.2)*100))+
  coord_cartesian(ylim=c(0,1))+
  labs(x="Age in months",y="Seroprevalence (%)") +
  scale_fill_manual(values=pcols) +
  scale_color_manual(values=pcols) +
  theme_minimal() +
  theme(legend.position = "none")

ppcurve

# save a PDF version
ggsave("~/enterics-immuno-epi/figs/asembo-ab-agecurves-seroprev.pdf",plot=ppcurve,device="pdf",width=7,height=14)

```

# Save fitted estimates for later use
```{r save estimates}
# spline fits of geometric means
saveRDS(gamfits,file="~/enterics-immuno-epi/output/asembo-enteric-ab-gamfits-means.rds")
# ensemble fits of geometric means
saveRDS(slfits,file="~/enterics-immuno-epi/output/asembo-enteric-ab-slfits-means.rds")
# spline fits of age-dependent seroprevalence
saveRDS(pgamfits,file="~/enterics-immuno-epi/output/asembo-enteric-ab-gamfits-seroprev.rds")
# spline fits of age-dependent seroprevalence
saveRDS(pslfits,file="~/enterics-immuno-epi/output/asembo-enteric-ab-slfits-seroprev.rds")
```

# Session Info
```{r session info}
sessionInfo()
```

