---
title: Enteropathogen seroepidemiology among children in low-resource settings
subtitle: Tanzan antibody distributions and seropositivity cutoffs, Figure 1 - supplements 1 and 2
output:
  html_document:
    highlight: haddock
    theme: default
    code_folding: show
    df_print: paged
    toc: yes
    toc_depth: '3'
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Summary

This script summarizes the distribution of enteric antibody responses in the Kongwa trachoma study. 

It fits Gaussian mixture models to the distributions, and where possible, estimates seropositivity cutoffs based on the mixture model fits.


# Script preamble
```{r preamble}
#-----------------------------
# preamble
#-----------------------------
library(here)
here()

# load packages
library(tidyverse)
library(viridis)
library(mixtools)

# set up for parallel computing
# configure for a laptop (use only 3 cores)
library(foreach)
library(doParallel)
registerDoParallel(cores=3)

# bright color blind palette:  https://personal.sron.nl/~pault/ 
cblack <- "#000004FF"
cblue <- "#3366AA"
cteal <- "#11AA99"
cgreen <- "#66AA55"
cchartr <- "#CCCC55"
cmagent <- "#992288"
cred <- "#EE3333"
corange <- "#EEA722"
cyellow <- "#FFEE33"
cgrey <- "#777777"

# viridis color palette
vircols <- viridis(n=7,alpha=1,begin=0.2,end=0.97)

```


#Load the formatted data

```{r data format}
#-----------------------------
# load the formatted data
# created with 
# asembo-enteric-ab-data-format.Rmd
#-----------------------------
dl <- readRDS(here("data","kongwa_analysis.rds"))

# list the enteric antigens and formatted labels for them
mbavars <- c("vsp3","vsp5","cp17","cp23","leca","salb","sald","etec","cholera","p18","p39","pgp3")

mbalabs <- c("Giardia VSP-3","Giardia VSP-5","Cryptosporidium Cp17","Cryptosporidium Cp23","E. histolytica LecA","Salmonella LPS group B","Salmonella LPS group D","ETEC LT β subunit","Cholera toxin β subunit","Campylobacter p18","Campylobacter p39","C. trachomatis pgp3")

dl <- dl %>%
  mutate(antigen=as.character(antigen))
```

#Distribution of antibody levels

## All ages 1-9 years
```{r Ab distribution figure, fig.height=12, fig.width=6}

# custom color blind color palette is in the preamble chunck
# pcols <- c(cred,corange,cgreen,cteal,cblue,cmagent,cgrey,cchartr)
pcols <- c(vircols[c(1:5)],cgrey,vircols[7:6])


# custom log labels
log10labs <- c( 
  expression(10^0),
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)

p <- ggplot(data=dl,aes(x=logmfi,group=pathogen,color=pathogen,fill=pathogen)) +
  facet_wrap(~antigenf,nrow=6,ncol=2,scales="free_y") +
  geom_density(color="black",size=0.24,alpha=0.7) +
  scale_x_continuous(limits = c(0,4.5),breaks = 0:4,labels=log10labs) +
  scale_fill_manual(values=pcols) +
  scale_color_manual(values=pcols) +
  labs(x="Luminex Response (MFI-bg)") +
  theme_minimal(base_size=16) +
  theme(legend.position = "none")

p

```



#Seropositivity cutoff values 

Fit mixture models among children age 1 

```{r roc cutoffs, echo=TRUE, message=FALSE, warning=FALSE}
#----------------------------
# Add ROC cutoffs 
# Estimated by Diana Martin
# for years 1 and 2-4 for antigens
# included in multiple years
# included in the file:
# "Key for Ben Arnold.xlsx"
#----------------------------
dl$roccut <- NA
dl$roccut[dl$antigen %in% "vsp3"] <- log10(116) 
dl$roccut[dl$antigen %in% "vsp5"] <- log10(109) 
dl$roccut[dl$antigen %in% "cp23"] <- log10(1185)
dl$roccut[dl$antigen %in% "cp17"] <- log10(321)
dl$roccut[dl$antigen %in% "leca"] <- log10(153)
dl$roccut[dl$antigen %in% "pgp3"] <- log10(818)


# different cutoffs for bead set used in years 2-4
dl$roccut[dl$antigen %in% "vsp3"  & dl$svy >1] <- log10(170)
dl$roccut[dl$antigen %in% "vsp5"  & dl$svy >1] <- log10(141)
dl$roccut[dl$antigen %in% "cp23"  & dl$svy >1] <- log10(378) 
dl$roccut[dl$antigen %in% "cp17"  & dl$svy >1] <- log10(180) 
dl$roccut[dl$antigen %in% "leca"  & dl$svy >1] <- log10(94)
dl$roccut[dl$antigen %in% "pgp3"  & dl$svy >1] <- log10(882)

```

```{r mixture cutoffs, echo=TRUE, message=FALSE, warning=FALSE}
#-----------------------------
# mixture models
#-----------------------------
fitmix <- function(x,lambda,k) {
  mixfit <- normalmixEM(x,lambda=lambda,k=k)
  mixcut <- mixfit$mu[order(mixfit$mu)][1]+3*mixfit$sigma[order(mixfit$mu)][1]
  cat(summary(mixfit),"\nCutoff value:",mixcut,"(log10 MFI), or:", round(10^mixcut),"(MFI)\n\n")
  # pull out fitted densities
  denmat <- matrix(NA,nrow=length(x),ncol=k)
  for(i in 1:k) {
    denmat[,i] <- mixfit$lambda[i] * dnorm(x,mean=mixfit$mu[i],sd=mixfit$sigma[i])
  }
  denmat <- data.frame(denmat)
  colnames(denmat) <- paste("den",1:k,sep="")
  # return original values plus fitted densities in a dataframe 
  # also return the cutoff value and normalmixEM object
  xden <- data.frame(x=x,denmat)
  list(xden=xden,mixcut=mixcut,mixfit=mixfit)
  
}


## mixture model fits
# store densities of mixture distributions
# and estimated cutoffs
# mixdens <- foreach(ab=c("vsp3","vsp5","cp17","cp23","etec","cholera","leca"),.combine=rbind) %do% {
set.seed(123)
mixdens <- foreach(ab=unique(dl$antigen),.combine=rbind) %do% {
  cat("\n\n Mixture model fit for ",ab,"\n\n")
  mixf <- fitmix(x=dl$logmfi[dl$antigen==ab & dl$age==1 & !is.na(dl$logmfi)],lambda=0.5,k=2)
  di <- mixf$xden
  di$mixcut <- mixf$mixcut
  di$antigen=ab
  di <- di %>% arrange(x) %>% select(antigen,mixcut,everything())
  di
}


# The only realistic cutoff values using a 2-component mixture are 
# for Giardia VSP-3 and VSP-5 and for E. hist LecA
# set all the others to missing
mixcuts <- mixdens %>%
  group_by(antigen) %>% filter(row_number()==1) %>% select(antigen,mixcut) %>% ungroup()
mixcuts$mixcut[ !mixcuts$antigen %in% c("vsp3","vsp5","leca","pgp3")] <- NA

#-----------------------------
# merge in the mixture cutoffs
# to the main dataset
#-----------------------------
dl <- left_join(dl,mixcuts,by=c("antigen"))

```

## Figure of antibody mixture distributions

```{r Ab mixture distribution figure, fig.height=12, fig.width=6}

# Get the mixture density distributions
# Distributions for p18, ETEC and Cholera did not have reasonable
# fits to set them to missing
# set the pathogen field to an arbitrary value (Giardia) to 
# mesh with the global ggplot aesthetics
plotmixdens <- mixdens %>%
  mutate(den1p = ifelse(!antigen %in% c("vsp3","vsp5","leca","pgp3"),NA,den1),
         den2p = ifelse(!antigen %in% c("vsp3","vsp5","leca","pgp3"),NA,den2)) %>%
  mutate(antigenf=factor(antigen,levels=mbavars,labels=mbalabs),
         pathogen="Giardia"
         )

# custom color blind color palette (pcols)
# and custom log10 labels (log10labs)
# defined in a previous code chunk
# pmix <- ggplot(data=filter(dl,age==1 & !antigen %in% c("p18","p39","salb","sald")),
pmix <- ggplot(data=filter(dl,age==1),
               aes(x=logmfi,group=pathogen,color=pathogen,fill=pathogen)) +
  facet_wrap(~antigenf,nrow=6,ncol=2,scales="free_y") +
  geom_density(color=NA,alpha=0.8) +
  # plot fitted mixture distributions
  geom_line(data=plotmixdens,aes(x=x,y=den1p),color=cblack) +
  geom_line(data=plotmixdens,aes(x=x,y=den2p),color=cblack) +
  # add vertical lines for the cutoff values
  geom_vline(aes(xintercept=roccut)) +
  geom_vline(aes(xintercept=mixcut),linetype=2) +
  # labels and formatting
  scale_x_continuous(limits = c(0,4.5),breaks = 0:4,labels=log10labs) +
  coord_cartesian(ylim=c(0,2)) +
  labs(x="Luminex Response (MFI-bg)") +
  scale_fill_manual(values=pcols) +
  scale_color_manual(values=pcols) +
  theme_minimal(base_size = 16) +
  theme(legend.position = "none")

pmix


```

## Figure 1 - supp 1: age-stratified antibody distributions, with cutoffs

**This is Figure 1 - supplement 1 in the article**

```{r age stratified distribution figure, fig.width=12,fig.height=6}
# create age categories, display a single ROC cutoff, put
# E. hist back in its correct place
dlplot <- dl %>%
  mutate(agecat = cut(age,breaks=c(0,1,2,3,4,10),
                      labels=c("1 year","2 years","3 years","4 years","5-9 years")),
         antigenf=factor(antigenf,levels=mbalabs)
         ) %>%
  filter(antigen!="pgp3") %>% mutate(antigenf=droplevels(antigenf)) %>%
  group_by(antigen) %>%
  mutate(plotroc=min(roccut))


# custom color blind color palette is in the preamble chunck
# pcols <- c(cred,corange,cgreen,cteal,cblue,cgrey,cmagent)
# pcols <- c(corange,cred,cmagent,cblue,cteal,cgreen,cgrey)
pcols <- vircols



# custom log labels
log10labs <- c( 
  expression(10^0),
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)

# faceted over age and antigen
p2 <- ggplot(data=dlplot,aes(x=logmfi,fill=pathogen,color=pathogen)) +
  facet_grid(agecat~antigenf,scales='free')+
  geom_density(aes(y=..density..),color="black",size=0.2,alpha=1)+
  geom_vline(aes(xintercept = plotroc)) +
  geom_vline(aes(xintercept = mixcut),linetype=2) +
  # scale_y_continuous(expand=c(0,0),breaks=seq(0,40,by=10)) +
  coord_cartesian(ylim=c(0, 1.25))+
  scale_x_continuous(expand=c(0,0),limits=c(0,5),breaks=0:4,labels=log10labs) +
  # scale_y_continuous(expand=c(0,0),breaks=c(0,0.5,1,1.5),labels=c("0","","1","")) +
  scale_fill_manual(values=pcols,guide_legend(title="Age category"))+
  labs(title="",x="Luminex Response (MFI-bg)",y="Density") +
  theme_grey()+
  theme(
    strip.text.x = element_text(size=9),
    strip.text.y=element_text(size=12,angle=0),
    # panel.grid.minor.y = element_blank(),
    legend.position="none"
  )

p2

# save PDF and TIFF versions
ggsave(here("figs","Fig1sup1-kongwa-ab-distributions-byage.pdf"),plot=p2,device=cairo_pdf,width=18,height=6)
ggsave(here("figs","Fig1sup1-kongwa-ab-distributions-byage.TIFF"),plot=p2,device="tiff",width=18,height=6)

```
## Figure 1 - supp 2: age-stratified antibody distributions of Giardia and pgp3

**This is Figure 1 - supplement 2 in the article**

```{r giar trachoma fig}
# create age categories, display a single ROC cutoff
# limit to Giardia VSP-3 and Trachoma pgp3
dlplot <- dl %>%
  mutate(agecat = cut(age,breaks=c(0,1,2,3,4,10),
                      labels=c("1 year","2 years","3 years","4 years","5-9 years")),
         antigenf=factor(antigenf,levels=mbalabs)
         ) %>%
  filter(antigen %in% c("vsp3","pgp3")) %>% mutate(antigenf=droplevels(antigenf)) %>%
  group_by(antigen) %>%
  mutate(plotroc=min(roccut))


# custom color blind color palette is in the preamble chunck
# pcols <- c(cred,cgrey)
pcols <- vircols[c(1,7)]

# custom log labels
log10labs <- c( 
  expression(10^0),
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)

# faceted over age and antigen
p3 <- ggplot(data=dlplot,aes(x=logmfi,fill=pathogen,color=pathogen)) +
  facet_grid(agecat~antigenf,scales='free')+
  geom_density(aes(y=..density..),color="black",size=0.2,alpha=1)+
  geom_vline(aes(xintercept = plotroc)) +
  geom_vline(aes(xintercept = mixcut),linetype=2) +
  # scale_y_continuous(expand=c(0,0),breaks=seq(0,40,by=10)) +
  coord_cartesian(ylim=c(0, 1.25))+
  scale_x_continuous(expand=c(0,0),limits=c(0,5),breaks=0:4,labels=log10labs) +
  # scale_y_continuous(expand=c(0,0),breaks=c(0,0.5,1,1.5),labels=c("0","","1","")) +
  scale_fill_manual(values=pcols,guide_legend(title="Age category"))+
  labs(title="",x="Luminex Response (MFI-bg)",y="Density") +
  theme_grey()+
  theme(
    strip.text.x = element_text(size=9),
    strip.text.y=element_text(size=12,angle=0),
    # panel.grid.minor.y = element_blank(),
    legend.position="none"
  )

p3

# save PDF and TIFF
ggsave(here("figs","Fig1sup2-kongwa-ab-distributions-byage-vsp3-pgp3.pdf"),plot=p3,device="pdf",width=5,height=6)
ggsave(here("figs","Fig1sup2-kongwa-ab-distributions-byage-vsp3-pgp3.TIFF"),plot=p3,device="tiff",width=5,height=6)

```


# Save data
```{r seropositivity}
dl <- dl %>%
  mutate(posroc = ifelse(logmfi > roccut,1,0),
         posmix = ifelse(logmfi > mixcut,1,0))

```

```{r save data}
#-----------------------------
# save an updated dataset
# that includes cutoff values
# and seropositive indicators
#-----------------------------
saveRDS(dl,file=here("data","kongwa_analysis2.rds"))

```

# Session Info
```{r session info}
sessionInfo()
```

