---
title: "Asembo enteric antibody analysis of age-dependent antibody curves"
output: 
  html_notebook:
    theme: default
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

#Summary
This script estimates age-dependent mean antibody curves using cubic splines (in a GAM model from the `mgcv` package) and using ensemble machine learning (using super learning from the `tmleAb` package). 

# Script preamble
```{r preamble}
#-----------------------------
# preamble
#-----------------------------
rm(list=ls())
library(tidyverse)
library(mixtools)

# set up for parallel computing
# configure for a laptop (use only 3 cores)
library(foreach)
library(doParallel)
registerDoParallel(cores=3)

# bright color blind palette:  https://personal.sron.nl/~pault/ 
cblack <- "#000004FF"
cblue <- "#3366AA"
cteal <- "#11AA99"
cgreen <- "#66AA55"
cchartr <- "#CCCC55"
cmagent <- "#992288"
cred <- "#EE3333"
corange <- "#EEA722"
cyellow <- "#FFEE33"
cgrey <- "#777777"

```


#Load the formatted data

```{r load data}
#-----------------------------
# load the formatted data
# with seropositivity indicators
# created with:
# asembo-enteric-ab-data-format.Rmd -->
# asembo-enteric-ab-distributions.Rmd
#-----------------------------
dl <- readRDS("~/enterics-seroepi/data/asembo_analysis2.rds")

```

#Estimate age-dependent mean antibody levels

##Estimate curves using splines in a generalized additive model (GAM)

```{r age-curve mean gam }
#----------------------------------
# simulataneous CIs for GAMs
# estimated by resampling the 
# Baysian posterior estimates of
# the variance-covariance matrix
# assuming that it is multivariate normal
# the function below also estimates 
# the unconditional variance-covariance
# matrix, Vb=vcov(x,unconditional=TRUE), 
# which allows for undertainty in the actual
# estimated mean as well 
# (Marra & Wood 2012 Scandinavian Journal of Statistics, 
#  Vol. 39: 53–74, 2012, doi: 10.1111/j.1467-9469.2011.00760.x )
# simultaneous CIs provide much better coverage than pointwise CIs
# see: http://www.fromthebottomoftheheap.net/2016/12/15/simultaneous-interval-revisited/
#----------------------------------

gamCI <- function(m,newdata,nreps=10000) {
  require(mgcv)
  require(dplyr)
  Vb <- vcov(m,unconditional = TRUE)
  pred <- predict(m, newdata, se.fit = TRUE)
  fit <- pred$fit
  se.fit <- pred$se.fit
  BUdiff <- MASS::mvrnorm(n=nreps, mu = rep(0, nrow(Vb)), Sigma = Vb)
  Cg <- predict(m, newdata, type = "lpmatrix")
  simDev <- Cg %*% t(BUdiff)
  absDev <- abs(sweep(simDev, 1, se.fit, FUN = "/"))
  masd <- apply(absDev, 2L, max)
  crit <- quantile(masd, prob = 0.95, type = 8)
  pred <- data.frame(newdata,fit=pred$fit,se.fit=pred$se.fit)
  pred <- mutate(pred,
                 uprP = fit + (2 * se.fit),
                 lwrP = fit - (2 * se.fit),
                 uprS = fit + (crit * se.fit),
                 lwrS = fit - (crit * se.fit)
  )
  return(pred)
}


# fit cubic spline over age, separately for each antigen
# estimate approximate simultaneous confidence intervals
# for the fits
library(mgcv)
gamfits <- foreach(ab=levels(dl$antigenf),.combine=rbind) %dopar% {
    pd <- filter(dl, antigenf==ab)
    gfit <- gam(logmfi~s(age, bs="cr"),data=pd)
    gsci <- gamCI(m=gfit,newdata=pd,nreps=1000)
    gsci$antigen <- ab
    return(gsci)
  }
```

##Estimate curves using ensemble machine learning
```{r age-curve mean ensemble, echo=TRUE, message=FALSE, warning=FALSE}
#-----------------------------
# estimate curves with 
# ensemble machine learning
# including a library with
# GLM, GAM, Loess, and 
# extreme gradient boosting
#-----------------------------

# detach the mgcv package because
# the ensemble algorithm uses gam()
# from the gam package and there are
# conflicts between them
detach("package:mgcv", unload=TRUE)

# load the tmleAb package, at this time
# only available through github: 
# devtools::install_github("ben-arnold/tmleAb")
library(tmleAb)
set.seed(12345)
slfits <-  foreach(ab=levels(dl$antigenf),.combine=rbind) %dopar% {
  pd <- filter(dl, antigenf==ab)
  slfit <- agecurveAb(Y=pd$logmfi,
                      Age=pd$age,
                      id=pd$childid,
                      SL.library = c("SL.glm","SL.gam","SL.loess","SL.xgboost"))
  sld <- data.frame(age=slfit$Age,fit=slfit$pY)
  sld <- sld %>% group_by(age) %>% filter(row_number()==1) %>% ungroup()
  slres <- left_join(pd,sld,by=c("age"))
  return(slres)
}


#-----------------------------
# estimate means by age in 
# months (4-17)
# to get 95% CIs through the
# influence curve
# there is 1 measurement
# at age 18 months that is
# excluded from this calculation
#-----------------------------

tmlemeans <- foreach(ab=levels(dl$antigenf),.combine=rbind) %:%
  foreach(agei=4:17,.combine=rbind) %dopar% {
    pd <- filter(dl,antigenf==ab & floor(age)==agei)
    mu <- tmleAb(Y=pd$logmfi,id=pd$id,SL.library="SL.mean")
    di <- data.frame(pathogen=pd$pathogen[1],antigenf=ab,age=agei,mean=mu$psi,se=mu$se,lb=mu$lb,ub=mu$ub)
    di
  }


```

##Plot age-dependent mean antibody curves

The spline fits are colored and the highly adaptive ensemble fits are grey. The approximate simultaneous confidence bands for the spline fit are shaded.

```{r age-curve mean figure all, fig.height=12, fig.width=6}

log10labs <- c( 
  expression(10^0),
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)

# bright color-blind palette defined in an earlier code chunk
pcols <- c(cred,corange,cteal,cblue,cmagent,cgrey,cgreen)
# pcols <- c(corange,cred,cmagent,cchartr,cblue,cteal,cgreen,cgrey)


pcurve <- ggplot(data=gamfits,aes(x=age,group=pathogen,color=pathogen,fill=pathogen)) +
  # approximate simultaneous CI from spline fit
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA,fill=cgrey) +
  # ensemble fit
  geom_line(data=slfits,aes(x=age,y=fit),color=cgrey) +
  # spline fit
  geom_line(aes(y=fit)) +
  # empirical means 95% CI from the influence curve (averaged for each age)
  geom_pointrange(data=tmlemeans,aes(x=age,y=mean,ymin=lb,ymax=ub),size=0.2,color=cblack)+
  # plot aesthetics
  facet_wrap(~antigenf,nrow=6,ncol=2) +
  scale_x_continuous(breaks=seq(4,18,by=2))+
  scale_y_continuous(breaks=0:4,labels=log10labs)+
  coord_cartesian(ylim=c(0,4.5))+
  labs(x="Age in months",y="Luminex Response (MFI-bg)") +
  scale_fill_manual(values=pcols) +
  scale_color_manual(values=pcols) +
  theme_minimal() +
  theme(legend.position = "none")

pcurve

# save a PDF version 
ggsave("~/enterics-seroepi/figs/asembo-ab-agecurves-mean-allestimates.pdf",plot=pcurve,device="pdf",width=7,height=14)

```
Create a more parsimonius plot, that only includes the spline fits but also includes the actual points. Only plot through age 17 months (since n=1 at 18 months)

```{r age-curve mean figure, fig.height=12, fig.width=6}

log10labs <- c( 
  expression(10^0),
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)

# bright color-blind palette defined in an earlier code chunk
pcols <- c(cred,corange,cteal,cblue,cmagent,cgrey,cgreen)
# pcols <- c(corange,cred,cmagent,cblue,cteal,cgreen,cgrey)


pcurve <- ggplot(data=filter(gamfits,age<=17),aes(x=age,group=pathogen,color=pathogen,fill=pathogen)) +
  # approximate simultaneous CI from spline fit
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA,fill=cgrey) +
  # actual measurements
  geom_jitter(aes(y=logmfi),width=0.2,height=0,alpha=0.4,size=0.3) +
  # spline fit
  geom_line(aes(y=fit),color="black") +
  # plot aesthetics
  facet_wrap(~antigenf,nrow=6,ncol=2) +
  scale_x_continuous(breaks=seq(4,18,by=2))+
  scale_y_continuous(breaks=0:4,labels=log10labs)+
  scale_fill_manual(values=pcols) +
  scale_color_manual(values=pcols) +
  coord_cartesian(ylim=c(0,4.5))+
  labs(x="Age in months",y="Luminex Response (MFI-bg)") +
  theme_minimal() +
  theme(legend.position = "none")

pcurve

# save a PDF version 
ggsave("~/enterics-seroepi/figs/asembo-ab-agecurves-mean.pdf",plot=pcurve,device="pdf",width=7,height=14)
```

#Estimate age-dependent seroprevalence

##Estimate seroprevalence curves using splines in a generalized additive model (GAM)


```{r composite antigens}
#-----------------------------
# create composite 
# seropositivity indicators
# that use information from
# multiple antigens
#-----------------------------

dgi <- dl %>% 
  filter(antigen %in% c("vsp3","vsp5")) %>%
  select(childid,time,tr,sex,age,agediff,pathogen,antigen,seropos) %>%
  spread(antigen,seropos) %>%
  mutate(seropos=ifelse(vsp3==1|vsp5==1,1,0),
         antigen="vsp3vsp5",
         antigenf="Giardia VSP-3 or VSP-5") %>%
  select(-vsp3,-vsp5)

dcr <- dl %>% 
  filter(antigen %in% c("cp17","cp23")) %>%
  select(childid,time,tr,sex,age,agediff,pathogen,antigen,seropos) %>%
  spread(antigen,seropos) %>%
  mutate(seropos=ifelse(cp17==1|cp23==1,1,0),
         antigen="cp17cp23",
         antigenf="Cryptosporidium Cp17 or Cp23") %>%
  select(-cp17,-cp23)

ds <- dl %>% 
  filter(antigen %in% c("salb","sald")) %>%
  select(childid,time,tr,sex,age,agediff,pathogen,antigen,seropos) %>%
  spread(antigen,seropos) %>%
  mutate(seropos=ifelse(salb==1|sald==1,1,0),
         antigen="salbsald",
         antigenf="Salmonella LPS groups B or D") %>%
  select(-salb,-sald)

dcp <- dl %>% 
  filter(antigen %in% c("p18","p39")) %>%
  select(childid,time,tr,sex,age,agediff,pathogen,antigen,seropos) %>%
  spread(antigen,seropos) %>%
  mutate(seropos=ifelse(p18==1|p39==1,1,0),
         antigen="p18p39",
         antigenf="Campylobacter p18 or p39") %>%
  select(-p18,-p39)



dlp <- bind_rows(dl,dgi,dcr,ds,dcp) %>%
  mutate(antigenf=factor(antigenf,
                         levels=c("Giardia VSP-3 or VSP-5",
                                  "Cryptosporidium Cp17 or Cp23",
                                  "E. histolytica LecA",
                                  "Salmonella LPS groups B or D",
                                  "ETEC toxin beta subunit",
                                  "Cholera toxin beta subunit",
                                  "Campylobacter p18 or p39")) ) %>%
  filter(!is.na(antigenf)) %>%
  select(pathogen,antigen,antigenf,
         childid,time,tr,sex,age,agediff,
         mfi,logmfi,logmfidiff,
         roccut,mixcut,unexp,unexpcut,serocut,serocut_desc,
         posroc,posmix,posunex,seropos) %>%
  arrange(pathogen,antigenf,childid,time)
```

```{r age-curve seroprevalence gam }
#-----------------------------
# estimate seroprevalence curves with 
# splines
#-----------------------------


# fit cubic spline over age, separately for each antigen
# estimate approximate simultaneous confidence intervals
# for the fits
library(mgcv)
pgamfits <- foreach(ab=levels(dlp$antigenf),.combine=rbind) %dopar% {
    pd <- filter(dlp, antigenf==ab)
    gfit <- mgcv::gam(seropos~s(age, bs="cr"),data=pd,family="binomial")
    gsci <- gamCI(m=gfit,newdata=pd,nreps=1000)
    gsci$antigenf <- ab
    return(gsci)
}

# use antilogit to get back to prevalence from the logit link
pgamfits <- pgamfits %>%
  mutate(fit  = 1/(1+exp(-fit)),
         uprP = 1/(1+exp(-uprP)),
         lwrP = 1/(1+exp(-lwrP)),
         uprS = 1/(1+exp(-uprS)),
         lwrS = 1/(1+exp(-lwrS))) %>%
  mutate(antigenf=factor(antigenf,levels=levels(dlp$antigenf)))
```

##Estimate seroprevalence curves using ensemble machine learning
```{r age-curve seroprevalence ensemble, echo=TRUE, message=FALSE, warning=FALSE}
#-----------------------------
# estimate curves with 
# ensemble machine learning
# including a library with
# GLM, GAM, Loess, and 
# extreme gradient boosting
#-----------------------------

# detach the mgcv package because
# the ensemble algorithm uses gam()
# from the gam package and there are
# conflicts between them
detach("package:mgcv", unload=TRUE)

set.seed(12345)
pslfits <-  foreach(ab=levels(dlp$antigenf),.combine=rbind) %dopar% {
  pd <- filter(dlp, antigenf==ab)
  slfit <- agecurveAb(Y=pd$seropos,
                      Age=pd$age,
                      id=pd$childid,
                      family="binomial",
                      SL.library = c("SL.glm","SL.gam","SL.xgboost"))
  sld <- data.frame(age=slfit$Age,fit=slfit$pY)
  sld <- sld %>% group_by(age) %>% filter(row_number()==1) %>% ungroup()
  slres <- left_join(pd,sld,by=c("age"))
  return(slres)
}

#-----------------------------
# estimate prevalence by age in 
# years (0-9)
# since prevalence is very sparse
# in some strata, cannot use the 
# tmleAb() function above for
# influence-curve based SEs
# because it uses cross-validation
# and sometimes there are folds
# with no events. So, instead
# use an exact binomial approach
#-----------------------------
exactprev <- function(x) {
  # x : a binary indicator of the outcome (1/0)
  tabx <- table(x)
  if(length(tabx)<2) {
    if(names(tabx)=="1") {
      tabx <- c(0,tabx)
    } else{
      tabx <- c(tabx,0)
    }
  } 
  estx <- binom.test(x=tabx[2],n=sum(tabx))
  res <- c(estx$parameter,estx$statistic,estx$estimate,estx$conf.int)
  names(res) <- c("N","n","mean","min95","max95")
  return(res)
}

prevests <- foreach(ab=levels(dlp$antigenf),.combine=rbind) %:%
  foreach(agei=4:17,.combine=rbind) %dopar% {
      pd <- filter(dlp,antigenf==ab & floor(age)==agei)
      mus <- exactprev(pd$seropos)
      res <- data.frame(pathogen=pd$pathogen[1],antigenf=ab,age=agei,N=mus[1],n=mus[2],mean=mus[3],lb=mus[4],ub=mus[5])
      res
    }
   

```

##Plot age-dependent seroprevalence  curves

The spline fits are colored and the highly adaptive ensemble fits are black. The approximate simultaneous confidence bands for the spline fit are shaded.

```{r age-curve seroprevalence figure all, fig.height=9, fig.width=6}


# bright color-blind palette defined in an earlier code chunk
pcols <- c(cred,corange,cgreen,cteal,cblue,cmagent,cgrey)
# pcols <- c(corange,cred,cmagent,cblue,cteal,cgreen,cgrey)


ppcurve <- ggplot(data=pgamfits,aes(x=age,color=antigenf,fill=antigenf)) +
  # approximate simultaneous CI from spline fit
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA) +
  # actual measurements
  # geom_jitter(aes(y=seropos),width=0.2,height=0,color="black",alpha=0.4,size=0.3) +
  # spline fit
  geom_line(aes(y=fit)) +
  # ensemble fit
  geom_line(data=pslfits,aes(x=age,y=fit),col="black") +
  # empirical means 95% CI from the influence curve (averaged for each age)
  geom_pointrange(data=prevests,aes(x=age,y=mean,ymin=lb,ymax=ub),size=0.2,color=cblack)+
  # plot aesthetics
  facet_wrap(~antigenf,nrow=6,ncol=2) +
  scale_x_continuous(breaks=seq(4,18,by=2))+
  scale_y_continuous(breaks=seq(0,1,by=0.2),labels=sprintf("%1.0f",seq(0,1,by=0.2)*100))+
  coord_cartesian(ylim=c(0,1))+
  labs(x="Age in months",y="Seroprevalence (%)") +
  scale_fill_manual(values=pcols) +
  scale_color_manual(values=pcols) +
  theme_minimal() +
  theme(legend.position = "none")

ppcurve

# save a PDF version
ggsave("~/enterics-seroepi/figs/asembo-ab-agecurves-seroprev-allests.pdf",plot=ppcurve,device="pdf",width=7,height=14)

```
Create a more parsimonius figure restricted to spline fits. 
Plot through age 17 months (since n=1 at 18 months).
Excluded Cholera due to cross-reactivity with ETEC and no suggestion that there was cholera transmission in the population at the time of the study.

```{r age-curve seroprevalence figure, fig.height=9, fig.width=6}


# bright color-blind palette defined in an earlier code chunk
pcols <- c(cred,corange,cgreen,cteal,cblue,cmagent)
# pcols <- c(corange,cred,cmagent,cchartr,cblue,cteal,cgreen,cgrey)


ppcurve2 <- ggplot(data=filter(pgamfits,age<=17 & antigen!="cholera"),aes(x=age,color=antigenf,fill=antigenf)) +
  # approximate simultaneous CI from spline fit
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA) +
  # actual measurements
  geom_jitter(aes(y=seropos),width=0.2,height=0,color="black",alpha=0.4,size=0.3) +
  # spline fit
  geom_line(aes(y=fit)) +
  # plot aesthetics
  facet_wrap(~antigenf,nrow=6,ncol=2) +
  scale_fill_manual(values=pcols) +
  scale_color_manual(values=pcols) +
  scale_x_continuous(breaks=seq(4,18,by=2))+
  scale_y_continuous(breaks=seq(0,1,by=0.2),labels=sprintf("%1.0f",seq(0,1,by=0.2)*100))+
  coord_cartesian(ylim=c(0,1))+
  labs(x="Age in months",y="Seroprevalence (%)") +
  theme_minimal() +
  theme(legend.position = "none")

ppcurve2

# save a PDF version
ggsave("~/enterics-seroepi/figs/asembo-ab-agecurves-seroprev.pdf",plot=ppcurve2,device="pdf",width=7,height=12)

```

Compressed figure (1 panel). grid.draw(), required to avoid cropping at plot margin while using custom labels, does not display in HTML -- only PDF, unfortunately. 

```{r age-curve seroprevalence figure small, fig.height=5, fig.width=5}

plabs <- pgamfits %>%
  filter(age==17 & antigen!="cholera") %>% select(antigenf,age,fit) %>% 
  group_by(antigenf) %>% summarize_all(mean) %>%
  mutate(fit=ifelse(antigenf=="Campylobacter p18 + p39",fit+0.02,fit),
         fit=ifelse(antigenf=="ETEC toxin beta subunit",fit-0.02,fit))

# bright color-blind palette defined in an earlier code chunk
pcols <- c(cred,corange,cgreen,cteal,cblue,cmagent)
# pcols <- c(corange,cred,cmagent,cchartr,cblue,cteal,cgreen,cgrey)


ppcurve3 <- ggplot(data=filter(pgamfits,age<=17 & antigen!="cholera"), aes(x=age,group=antigenf,color=antigenf,fill=antigenf)) +
  # approximate simultaneous CI from spline fit
  # geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA) +
  # spline fit
  geom_line(aes(y=fit)) +
  # pathogen labels
  geom_text(data=plabs,aes(x=age+0.5,y=fit,label=antigenf),
            hjust=0)+
  # plot aesthetics
  # facet_wrap(~antigenf,nrow=6,ncol=2) +
  scale_x_continuous(breaks=seq(4,16,by=2))+
  scale_y_continuous(breaks=seq(0,1,by=0.2),labels=sprintf("%1.0f",seq(0,1,by=0.2)*100))+
  scale_fill_manual(values=pcols) +
  scale_color_manual(values=pcols) +
  coord_cartesian(ylim=c(0,1),xlim=c(4,17))+
  labs(x="Age in months",y="Seroprevalence (%)") +
  theme_minimal(base_size=14) +
  theme(legend.position = "none",
        plot.margin = unit(c(1,6,1,1),"cm")
        )

# Code to override clipping for labels
ppcurve4 <- ggplot_gtable(ggplot_build(ppcurve3))
ppcurve4$layout$clip[ppcurve4$layout$name == "panel"] <- "off"
library(grid)
grid.draw(ppcurve4)

# save a PDF version
ggsave("~/enterics-seroepi/figs/asembo-ab-agecurves-seroprev-small.pdf",plot=ppcurve4,device="pdf",width=9,height=7)

```

# Save fitted estimates for later use
```{r save estimates}
# spline fits of geometric means
saveRDS(gamfits,file="~/enterics-seroepi/output/asembo-enteric-ab-gamfits-means.rds")
# ensemble fits of geometric means
saveRDS(slfits,file="~/enterics-seroepi/output/asembo-enteric-ab-slfits-means.rds")
# spline fits of age-dependent seroprevalence
saveRDS(pgamfits,file="~/enterics-seroepi/output/asembo-enteric-ab-gamfits-seroprev.rds")
# spline fits of age-dependent seroprevalence
saveRDS(pslfits,file="~/enterics-seroepi/output/asembo-enteric-ab-slfits-seroprev.rds")
```

# Session Info
```{r session info}
sessionInfo()
```

