---
title: Enteropathogen antibody dynamics and force of infection among children in low-resource settings
subtitle: Kenya antibody profiles, Figure 4 - supplements 1 and 3
output: 
  html_document:
    highlight: haddock
    theme: default
    code_folding: hide
    df_print: paged
    toc: yes
    toc_depth: '3'
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Notebook Summary

This script plots individual antibody trajectories in the Asembo, Kenya cohort from enrollment to follow-up. 205 children were measured at both timepoints.

# Script preamble
```{r preamble}
#-----------------------------
# preamble
#-----------------------------
library(here)
here::here()

# load packages
library(tidyverse)

# bright color blind palette:  https://personal.sron.nl/~pault/ 
cblack <- "#000004FF"
cblue <- "#3366AA"
cteal <- "#11AA99"
cgreen <- "#66AA55"
cchartr <- "#CCCC55"
cmagent <- "#992288"
cred <- "#EE3333"
corange <- "#EEA722"
cyellow <- "#FFEE33"
cgrey <- "#777777"

```


```{r load data}
#-----------------------------
# load the formatted data
# created with 
# asembo-enteric-ab-data-format.Rmd -->
# asembo-enteric-ab-distributions.Rmd
#-----------------------------
dl <- readRDS(here::here("data","asembo_analysis2.rds"))
```

# Identify incident changes
Identify children whose status changed  between enrollment and follow-up. Those who changed from negative to positive are seroconverters (`seroi` below), and those who changed from positive to negative are seroreverters (`seror` below).


```{r long incidence}

#-----------------------------
# identify incident 
# seroconversions and reversions
# based on crossing the
# seropositivity cutoff
# and
# based on a 4-fold change in MFI
# to above the cutoff (seroconversion)
# or starting above cutoff (seroreversion)
#-----------------------------
dlp <- dl %>%
  group_by(antigen,antigenf,childid) %>%
  mutate(seroposA=ifelse(time=="A",seropos,NA),
         seroposA=max(seroposA,na.rm=TRUE),
         seroposB=ifelse(time=="B",seropos,NA),
         seroposB=max(seroposB,na.rm=TRUE),
         seroi=ifelse(seroposB==1 & seroposA==0,1,0),
         seror=ifelse(seroposB==0 & seroposA==1,1,0),
         
         mfiA=ifelse(time=="A",logmfi,NA),
         mfiA=max(mfiA,na.rm=TRUE),
         mfiB=ifelse(time=="B",logmfi,NA),
         mfiB=max(mfiB,na.rm=TRUE),
         seroi4fold=ifelse(logmfidiff>log10(4) & mfiB>serocut,1,0),
         seror4fold=ifelse(logmfidiff< - log10(4) & mfiA>serocut,1,0)
  ) 



```

# Antibody MFI trajectories
Longitudinal antibody trajectories of individual children for all antigens measured, colored by the change in antibody levels (increases and decreases)
```{r change in antibody level figure, fig.height=12, fig.width=6}
#-----------------------------
# Plot individual antibody
# trajectories, colored
# by increases and decreases
#-----------------------------

# create an change category factor for plot aesthetics
dlp <- dlp %>%
  mutate(mficat=factor(ifelse(logmfidiff>0,"Increase","Decrease"),levels=c("Increase","Decrease")))

# convert time to numeric (for plotting)
dlp <- dlp %>%
  mutate(svy=ifelse(time=="A",0,1))

log10labs <- c( 
  expression(10^0),
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)

ppq <- ggplot(data=filter(dlp,!is.na(mficat)),aes(x=svy,y=logmfi,group=factor(childid),color=mficat)) +
  # geom_point(alpha=0.2) +
  geom_line(alpha=0.2) +
  # geom_hline(aes(yintercept=mixcut),linetype="dashed") +
  
  scale_x_continuous(limits=c(-0.25,1.25),breaks=c(0,1),labels=c("Enrollment","Follow-up"))+
  scale_y_continuous(breaks=0:4,labels=log10labs) +
  scale_color_manual(values=c(corange,cteal,"gray80"),
                     guide=guide_legend(title="MFI change:",override.aes = list(alpha=1))
                     ) +
  facet_wrap(~antigenf,nrow=6,ncol=2) +
  labs(x="Measurement",y="Luminex response (MFI-bg)") +
  theme_minimal() +
  theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.ticks.y=element_line(color="gray40"),
    legend.position = "top"
    )

ppq

```

Create the same figure of individual antibody trajectories between measurements, but limit it to antigens with seroincidence measures, and color the trajectories by seroconversion/reversion status. 

```{r change in antibody seroincidence figure, fig.height=9, fig.width=6}
#-----------------------------
# pull the incidence info
# and just merge it back
# to the full longidudinal data
#-----------------------------
dl2 <- dlp

# create an incidence category factor for plot aesthetics
# for seroconversion based on seropositivity only
dl2$icat <- factor(NA,levels=c("Seroconversion","Seroreversion","No change"))
dl2$icat[dl2$seroi==1] <- "Seroconversion"
dl2$icat[dl2$seror==1] <- "Seroreversion"
dl2$icat[dl2$seroi==0 & dl2$seror==0] <- "No change"


# create an incidence category factor for plot aesthetics
# for seroconversion based on 4-fold changes in MFI
dl2$i4cat <- factor(NA,levels=c("Seroconversion","Seroreversion","No change"))
dl2$i4cat[dl2$seroi4fold==1] <- "Seroconversion"
dl2$i4cat[dl2$seror4fold==1] <- "Seroreversion"
dl2$i4cat[dl2$seroi4fold==0 & dl2$seror==0] <- "No change"

# create an incidence category factor for plot aesthetics
# for seroconversion comparing both methods
dl2$icat_comp <- factor(NA,levels=c(">4-fold increase, across cutoff",">4-fold decrease",">4-fold increase, above cutoff","<4-fold change"))
dl2$icat_comp[dl2$seroi4fold==1 & dl2$seroi==1] <- ">4-fold increase, across cutoff"
dl2$icat_comp[dl2$seror4fold==1] <- ">4-fold decrease"
dl2$icat_comp[dl2$seroi4fold==1 & dl2$seroi==0] <- ">4-fold increase, above cutoff"
dl2$icat_comp[is.na(dl2$icat_comp)] <- "<4-fold change"
dl2 <- dl2 %>%
  mutate(icat_comp = factor(icat_comp,levels=c(">4-fold increase, across cutoff",
                                               ">4-fold increase, above cutoff",
                                               ">4-fold decrease",
                                               "<4-fold change"))
         )


pp <- ggplot(data=filter(dl2,!is.na(icat)),aes(x=svy,y=logmfi,group=factor(childid),color=icat_comp)) +
  # geom_point(alpha=0.2) +
  geom_hline(aes(yintercept=serocut),linetype="dashed",size=0.3) +
  geom_line(alpha=0.2) +
  
  scale_x_continuous(limits=c(-0.25,1.25),breaks=c(0,1),labels=c("Enrollment","Follow-up"))+
  scale_y_continuous(breaks=0:4,labels=log10labs) +
  scale_color_manual(values=c(corange,cmagent,cteal,"gray80"),
                     guide=guide_legend(title="MFI change:", override.aes = list(alpha=1), ncol=2,nrow=2),
                     ) +
  facet_wrap(~antigenf,nrow=6,ncol=2) +
  labs(x="Measurement",y="Luminex response (MFI-bg)") +
  theme_minimal() +
  theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.ticks.y=element_line(color="gray40"),
    legend.position = "top" 
    )

pp

```

# Figure 4 - supplement 1
```{r change in antibody seroincidence figure subset, fig.width=7,fig.height=12}
#-----------------------------
# final figure, excluding cholera
# due to cross-reactivity with ETEC
#-----------------------------
relev <- levels(dl2$antigenf)[c(1:6,11,7,9,10)]
dl3 <- dl2 %>%
  filter(!is.na(icat) & antigen!="cholera") %>%
  ungroup() %>%
  mutate(antigenf=factor(antigenf,levels=relev))


pp2 <- ggplot(data=dl3,aes(x=svy,y=logmfi,group=factor(childid),color=icat_comp)) +
  # geom_point(alpha=0.2) +
  geom_hline(aes(yintercept=serocut),linetype="dashed",size=0.3) +
  geom_line(alpha=0.2) +
  
  scale_x_continuous(limits=c(-0.25,1.25),breaks=c(0,1),labels=c("Enrollment","Follow-up"))+
  scale_y_continuous(breaks=0:4,labels=log10labs) +
  scale_color_manual(values=c(corange,cmagent,cteal,"gray80"),
                     guide=guide_legend(title="MFI change:", override.aes = list(alpha=1), ncol=2,nrow=2),
                     ) +
  facet_wrap(~antigenf,nrow=6,ncol=2) +
  labs(x="Measurement",y="Luminex response (MFI-bg)") +
  theme_minimal() +
  theme(
    panel.grid.minor.x = element_blank(),
    panel.grid.minor.y = element_blank(),
    panel.grid.major.y = element_blank(),
    axis.ticks.y=element_line(color="gray40"),
    legend.position = "top" 
    )

pp2

# save PDF and TIFF versions
ggsave(here::here("figs","Fig4sup1-asembo-ab-trajectories.pdf"),plot=pp2,device=cairo_pdf,width=5,height=10)
ggsave(here::here("figs","Fig4sup1-asembo-ab-trajectories.TIFF"),plot=pp2,device="tiff",width=5,height=10)

```

# Figure 4 - supplement 3
Estimate prevalence of each category
```{r estimate prevalence of each class , warning=FALSE}
dl4 <- dl3 %>%
  group_by(antigenf,icat_comp, .drop=FALSE) %>%
  tally() %>%
  group_by(antigenf) %>%
  mutate(N=sum(n),
         q=N-n,
         prev=n/N
         )
# exact binomial confidence intervals on prevalence estimates
ci_lb <- apply(dl4[c("n","q")],1,function(x) binom.test(x,alternative="two.sided")$conf.int[1])
ci_ub <- apply(dl4[c("n","q")],1,function(x) binom.test(x,alternative="two.sided")$conf.int[2])
dl4 <- dl4 %>%
  bind_cols(data.frame(ci_lb,ci_ub)) %>%
  arrange(antigenf,icat_comp)
  
```
```{r figure proportion in each class by measurement, fig.width=7,fig.height=12, warning=FALSE}
# compare seroconversion classification prevalences by measurement
sero_comp_p <- ggplot(data=dl4, aes(x=icat_comp, y=prev,color=icat_comp) ) +
    facet_wrap(~antigenf,nrow=6,ncol=2)+
    geom_errorbar(aes(ymin=ci_lb,ymax=ci_ub),width=0.1)+
    geom_point()+
    scale_color_manual(values=c(corange,cmagent,cteal,"gray50"),
                       guide=guide_legend(title="MFI change:",
                                          override.aes = list(alpha=1),
                                          nrow=2,ncol=2
                                          ))+
    xlab("From Enrollment to Follow-up") +
    ylab("Percentage of children (%)") +
    coord_cartesian(ylim=c(0,1),xlim=c(1,4))+
    scale_y_continuous(breaks=seq(0,1,by=0.2),labels=sprintf("%1.0f",seq(0,1,by=0.2)*100)) +
    # scale_x_continuous(breaks=1:5) +
    theme_minimal() +
    theme(
      legend.position="top",
      panel.grid.minor.x=element_blank(),
      axis.text.x = element_blank()
    )
  
sero_comp_p

# save PDF and TIFF versions
ggsave(filename=here::here("figs","Fig4sup3-asembo-ab-trajectory-meas.pdf"),plot = sero_comp_p, device=cairo_pdf, width=5,height=10)
ggsave(filename=here::here("figs","Fig4sup3-asembo-ab-trajectories-meas.TIFF"),plot = sero_comp_p, device='tiff', width=5,height=10)

```

# Session Info
```{r session info}
sessionInfo()
```




