---
title: "Enteropathogen seroepidemiology among children in low-resource settings"
subtitle: "Haiti antibody profiles, Figure 5 - supplement 1, Table 2"
output: 
  html_document:
    highlight: haddock
    theme: default
    code_folding: show
    df_print: paged
    toc: yes
    toc_depth: '3'
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  html_notebook:
    theme: default
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Notebook Summary

This script plots individual antibody trajectories in the Haitian cohort over successive measurement visits and also by age.  Children were measured up to 9 times, but the majority were measured 5 times, so figures by measurement round include up to 6 visits. 

# Script preamble
```{r preamble}
#-----------------------------
# preamble
#-----------------------------
library(here)
here()

# load packages
library(tidyverse)
library(kableExtra)
library(ROCR)


# set up for parallel computing
# configure for a laptop (use only 3 cores)
library(foreach)
library(doParallel)
registerDoParallel(cores=3)

# bright color blind palette:  https://personal.sron.nl/~pault/ 
cblack <- "#000004FF"
cblue <- "#3366AA"
cteal <- "#11AA99"
cgreen <- "#66AA55"
cchartr <- "#CCCC55"
cmagent <- "#992288"
cred <- "#EE3333"
corange <- "#EEA722"
cyellow <- "#FFEE33"
cgrey <- "#777777"

```


```{r load data}
#-----------------------------
# load the formatted data
# created with 
# asembo-enteric-ab-data-format.Rmd
#-----------------------------
dl <- readRDS(here("data","haiti_analysis2.rds"))

# list the enteric antigens and formatted labels for them
mbavars <- c("vsp3","vsp5","cp17","cp23","leca","etec","salb","sald","norogi","norogii")

mbalabs <- c("Giardia VSP-3","Giardia VSP-5","Cryptosporidium Cp17","Cryptosporidium Cp23","E. histolytica LecA","ETEC toxin beta subunit","Salmonella LPS B","Salmonella LPS D","Norovirus GI", "Norovirus GII")
```

# Identify incident changes
Among children, identify those that changed status between enrollment and follow-up. Those who changed from negative to positive are seroconverters (`seroi` below), and those who changed from positive to negative are seroreverters (`seror` below).


```{r long incidence}

#-----------------------------
# identify seropositive measures
# hierarchy of information for 
# cutoffs:
# 1. ROC
# 2. mixture model based
# 3. estimated among presumed unexposed
#
# store the cutoff value used
# for figures
#-----------------------------
dl <- dl %>%
  mutate(seropos=ifelse(!is.na(posroc),posroc,posmix),
         serocut=ifelse(!is.na(posroc),roccut,mixcut),
         serocut_desc=ifelse(!is.na(posroc),"ROC","Mixture Model")) %>%
  mutate(seropos=ifelse(!is.na(seropos),seropos,posunex),
         serocut_desc=ifelse(!is.na(serocut),serocut_desc,"Unexp dist"),
         serocut=ifelse(!is.na(serocut),serocut,unexpcut)) 


#-----------------------------
# identify incident 
# seroconversions and reversions
#-----------------------------

# group the data by child and
# use lags to identify
# time in years between measurements,
# sero-conversions + sero-reversions 
# between measurements
# set the first measurement to 
# missing for the incidence indicators
# (using the is.na(age_diff) to identify them)
dl2 <- dl %>%
  group_by(antigen,id) %>% 
  arrange(antigen,id,sdate,sampleid) %>%
  mutate(age_min  = min(age),
         age_diff = age - lag(age),
         
         logmfi_lag  = lag(logmfi),
         logmfi_lead = lead(logmfi),
         logmfi_dlag  = logmfi - logmfi_lag,
         logmfi_dlead = logmfi_lead - logmfi,
         
         # incident seroconversions and reversions
         # including cumulative numbers
         seropos_lag  = lag(seropos),
         seroi = ifelse(seropos==1 & seropos_lag==0,1,0),
         seroi = ifelse(is.na(age_diff),NA,seroi),
         seroin = cumsum(ifelse(is.na(seroi),0,seroi)),
         seroin = ifelse(seroi==1,seroin,0),
         seror = ifelse(seropos==0 & seropos_lag==1,1,0),
         seror = ifelse(is.na(age_diff),NA,seror),
         serorn = cumsum(ifelse(is.na(seror),0,seror)),
         serorn = ifelse(seror==1,serorn,0)
         ) %>%
  select(
         id,sampleid,sdate,
         starts_with("age"),
         pathogen,antigen,antigenf,
         mfi,logmfi,logmfi_dlag,logmfi_dlead,
         roccut,mixcut,serocut,
         posroc,posmix,seropos,seroi,seroin,seror,serorn,
         -logmfi_lag,-logmfi_lead,-seropos_lag)

# incident seroconversions based on a 4-fold increase in MFI
# with a second measure above the seropositivity cutoff
# incident seroreversions based on a 4-fold decrease in MFI
# with the first measure above the seropositivity cutoff
dl2 <- dl2 %>%
  mutate(seroi4fold = ifelse(logmfi_dlag>log10(4) & logmfi>serocut,1,0),
         seroin4fold = cumsum(ifelse(is.na(seroi4fold),0,seroi4fold)),
         seroin4fold = ifelse(seroi4fold==1,seroin4fold,0),
         
         seror4fold=ifelse(logmfi_dlag< -log10(4) & lag(logmfi)>serocut,1,0),
         serorn4fold = cumsum(ifelse(is.na(seror4fold),0,seror4fold)),
         serorn4fold = ifelse(seror4fold==1,serorn4fold,0)
         )


```

Create labels for different seroconversion patterns
```{r seroconversion patterns}
dl2pp <- dl2 %>%
  arrange(antigen,id,sdate,sampleid) %>%
  group_by(antigen,id) %>%
  mutate(measnum=row_number()) %>%
  mutate(
    logmfi_chng = ifelse(logmfi_dlag>0,"Increasing","Decreasing"),
    logmfi_chng = ifelse(is.na(logmfi_chng) & logmfi_dlead>0,"Increasing",logmfi_chng),
    logmfi_chng = ifelse(is.na(logmfi_chng) & logmfi_dlead<0,"Decreasing",logmfi_chng),
    logmfi_chng = factor(logmfi_chng,levels=c("Increasing","Decreasing")),
    
    sero_chng = ifelse(lead(seroi4fold)==1,">4-fold increase to above cutoff","<4-fold change"),
    sero_chng = ifelse(lead(seror4fold)==1,">4-fold decrease from above cutoff",sero_chng),
    sero_chng = ifelse(is.na(sero_chng),lead(sero_chng),sero_chng),
    sero_chng = ifelse(is.na(sero_chng),lag(sero_chng),sero_chng),
    sero_chng = factor(sero_chng,levels=c(">4-fold increase to above cutoff",">4-fold decrease from above cutoff","<4-fold change")),
    
    sero_comp = ifelse(lead(seroi)==1 & lead(seroi4fold==1),">4-fold increase, across cutoff","<4-fold change"),
    sero_comp = ifelse(lead(seroi)==0 & lead(seroi4fold==1),">4-fold increase, above cutoff",sero_comp),
    sero_comp = ifelse(lead(seror4fold)==1,">4-fold decrease",sero_comp),
    sero_comp = factor(sero_comp,levels=c(">4-fold increase, across cutoff",">4-fold decrease",">4-fold increase, above cutoff","<4-fold change")),
    
    everseroi = max(seroi,na.rm=TRUE)
    
  ) %>%
  mutate(antigenf=factor(antigenf,levels=mbalabs)) # re-level the factor to put Salmonella on same row

# custom log labels
log10labs <- c( 
  expression(10^0),
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)

```

# Longitudinal child trajectories

```{r figure of individual MFI trajectories by age, fig.width=7,fig.height=7}
# Individual trajectories by age
indivp_age <- ggplot(filter(dl2pp,!is.na(sero_chng)), aes(x=age, y=logmfi, group=factor(id),color=sero_chng) ) +
    facet_wrap(~antigenf,nrow=6,ncol=2)+
    geom_line(alpha=0.3) +
    geom_hline(aes(yintercept=serocut),linetype="dashed") +
    scale_color_manual(values=c(corange,cteal,"gray70"),
                       guide=guide_legend(title="MFI change:", 
                                          override.aes = list(alpha=1),
                                          ncol=2,nrow=2
                                          ))+
    xlab("Age in years") +
    ylab("Luminex Response (MFI-bg)") +
    coord_cartesian(ylim=c(0, 4.5),xlim=c(0,11))+
    scale_y_continuous(breaks=0:4,labels=log10labs) +
    scale_x_continuous(breaks=0:11) +
    theme_minimal() +
    theme(
      legend.position="top",
      panel.grid.minor=element_blank(),
      panel.grid.major.y=element_blank(),
      axis.ticks.y=element_line(color="gray40")
    )
  
indivp_age

```

```{r figure of individual MFI trajectories by age compare class, fig.width=7,fig.height=7}
# Individual trajectories by age, comparison of classifications
indivp_age <- ggplot(filter(dl2pp,!is.na(sero_comp)), aes(x=age, y=logmfi, group=factor(id),color=sero_comp) ) +
    facet_wrap(~antigenf,nrow=6,ncol=2)+
    geom_line(alpha=0.3) +
    geom_hline(aes(yintercept=serocut),linetype="dashed") +
    scale_color_manual(values=c(corange,cteal,cmagent,"gray70"),
                       guide=guide_legend(title="MFI change:",
                                          override.aes = list(alpha=1),
                                          nrow=2,ncol=2
                                          ))+
    xlab("Age in years") +
    ylab("Luminex Response (MFI-bg)") +
    coord_cartesian(ylim=c(0, 4.5),xlim=c(0,11))+
    scale_y_continuous(breaks=0:4,labels=log10labs) +
    scale_x_continuous(breaks=0:11) +
    theme_minimal() +
    theme(
      legend.position="top",
      panel.grid.minor=element_blank(),
      panel.grid.major.y=element_blank(),
      axis.ticks.y=element_line(color="gray40")
    )
  
indivp_age

```

```{r figure of individual MFI trajectories by measurement, fig.width=7,fig.height=7}
# Individual trajectories by measurement (1 to 6)
# since only 29 children were measured 7+ times
table(dl2pp$measnum[dl2pp$antigen=="vsp3"])
# summarize age distribution among measurements 1-6
summary(dl2pp$age[dl2pp$antigen=="vsp3" & dl2pp$measnum<=6])
indivp_meas <- ggplot(filter(dl2pp,!is.na(sero_chng) & measnum<=6), aes(x=measnum, y=logmfi, group=factor(id),color=sero_chng) ) +
    facet_wrap(~antigenf,nrow=6,ncol=2)+
    geom_line(alpha=0.3) +
    geom_hline(aes(yintercept=serocut),linetype="dashed") +
    scale_color_manual(values=c(corange,cteal,"gray70"),
                       guide=guide_legend(title="MFI change:",
                                          override.aes = list(alpha=1),
                                          nrow=2,ncol=2
                                          ))+
    xlab("Measurement number") +
    ylab("Luminex Response (MFI-bg)") +
    coord_cartesian(ylim=c(0, 4.5),xlim=c(1,6))+
    scale_y_continuous(breaks=0:4,labels=log10labs) +
    scale_x_continuous(breaks=1:6) +
    theme_minimal(base_size=16) +
    theme(
      legend.position="top",
      panel.grid.minor=element_blank(),
      panel.grid.major.y=element_blank(),
      axis.ticks.y=element_line(color="gray40")
    )
  
indivp_meas

```

# Figure 5 - supplement 1, Individual trajectories by measurement round
```{r figure of individual MFI trajectories by measurement compare class, fig.width=7,fig.height=7}
# Individual trajectories by measurement (1 to 6)
# since only 29 children were measured 7+ times
# compare seroconversion classification
indivp_meas2 <- ggplot(filter(dl2pp,!is.na(sero_comp) & measnum<=6), aes(x=measnum, y=logmfi, group=factor(id),color=sero_comp) ) +
    facet_wrap(~antigenf,nrow=6,ncol=2)+
    geom_line(alpha=0.3) +
    geom_hline(aes(yintercept=serocut),linetype="dashed") +
    scale_color_manual(values=c(corange,cteal,cmagent,"gray70"),
                       guide=guide_legend(title="MFI change:",
                                          override.aes = list(alpha=1),
                                          nrow=2,ncol=2
                                          ))+
    xlab("Measurement number") +
    ylab("Luminex Response (MFI-bg)") +
    coord_cartesian(ylim=c(0, 4.5),xlim=c(1,6))+
    scale_y_continuous(breaks=0:4,labels=log10labs) +
    scale_x_continuous(breaks=1:6) +
    theme_minimal() +
    theme(
      legend.position="top",
      panel.grid.minor=element_blank(),
      panel.grid.major.y=element_blank(),
      axis.ticks.y=element_line(color="gray40")
    )
  
indivp_meas2

# save PDF and TIFF versions
ggsave(filename=here("figs","Fig5sup1-haiti-indiv-ab-trajectories-meas.pdf"),plot = indivp_meas, device='pdf', width=6,height=12)
ggsave(filename=here("figs","Fig5sup1-haiti-indiv-ab-trajectories-meas.TIFF"),plot = indivp_meas, device='tiff', width=6,height=12)

```


```{r figure of individual MFI trajectories}

dl2plot <- dl2 %>% group_by(id)
dl2plot$sid <- group_indices(dl2plot)
dl2plot$sid <- factor(dl2plot$sid)

# color by MFI increases/decreases or seroconversion/reversion statsus
dl2plot <- dl2plot %>%
  mutate(
    logmfi_chng  = ifelse(logmfi_dlag>0,"Increasing","Decreasing"),
    logmfi_chng = ifelse(is.na(logmfi_chng) & logmfi_dlead>0,"Increasing",logmfi_chng),
    logmfi_chng = ifelse(is.na(logmfi_chng) & logmfi_dlead<0,"Decreasing",logmfi_chng),
    logmfi_chng = factor(logmfi_chng,levels=c("Increasing","Decreasing")),
    
    sero_chng = ifelse(lead(seroi)==1,"Seroconversion","No change"),
    sero_chng = ifelse(lead(seror)==1,"Seroreversion",sero_chng),
    sero_chng = ifelse(is.na(sero_chng),lead(sero_chng),sero_chng),
    sero_chng = factor(sero_chng,levels=c("Seroconversion","Seroreversion","No change"))
    
  )


# custom color blind color palette is in the preamble chunck
# pcols <- c(cred,corange,cgreen,cteal,cblue,cmagent)
pcols <- c(corange,cred,cmagent,cblue,cteal,cgreen)

# custom log labels
log10labs <- c( 
  expression(10^0),
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)

# plot all trajectories, colored (doesn't look that great)
# p <- ggplot(dl2plot, aes(x=age, y=logmfi, group=factor(id),color=sero_chng) ) +
#     facet_wrap(~antigenf,nrow=6,ncol=2)+
#     geom_line(alpha=0.1) +
#     guides(colour=FALSE) +
#     guides(alpha=FALSE) +
#     # geom_line(data=filter(dl2plot,sid==iid),aes(x=age,y=logmfi),size=1.2) +
#     scale_color_manual(values=c(corange,cteal,cgrey))+
#     xlab("Child Age (Years)") +
#     ylab("Luminex Response (MFI-bg)") +
#     coord_cartesian(ylim=c(0, 4.5))+
#     scale_y_continuous(breaks=0:4,labels=log10labs) +
#     scale_x_continuous(limits=c(0,10),breaks=seq(0,10,by=2)) +
#     theme_minimal() +
#     theme(
#       # strip.text = element_text(size=16),
#       # axis.text.x=element_text(size=12),
#       # axis.text.y=element_text(size=12),
#       # axis.title.x=element_text(size=14),
#       # axis.title.y=element_text(size=14),
#       legend.position="none"
#     )
# 
# p

# highlight a handfull of example individual trajectories
for(iid in c(20,24,30,34,35,38,43,47,72,76,77,119,135)) {
  p <- ggplot(dl2plot, aes(x=age, y=logmfi, group=factor(id)) ) +
    facet_wrap(~antigenf,nrow=6,ncol=2)+
    geom_line(alpha=0.1,color=cgrey) +
    guides(colour=FALSE) +
    guides(alpha=FALSE) +
    geom_hline(aes(yintercept=serocut),linetype="dashed") +
    geom_line(data=filter(dl2plot,sid==iid),aes(x=age,y=logmfi,color=sero_chng)) +
    scale_color_manual(values=c(corange,cteal,cgrey))+
    xlab("Child Age (Years)") +
    ylab("Luminex Response (MFI-bg)") +
    coord_cartesian(ylim=c(0, 4.5))+
    scale_y_continuous(breaks=0:4,labels=log10labs) +
    scale_x_continuous(limits=c(0,10),breaks=seq(0,10,by=2)) +
    theme_minimal() +
    theme(
      legend.position="none"
    )
  
  p

  ggsave(filename=here("figs",paste("haiti-spaghetti-",iid,".pdf",sep="")),plot = p,device='pdf',width=5,height=14)

}

```



# Method agreement for identifying seroconversion

Compare the agreement in the identification of seroconversions based on two approaches: crossing the seropositivity threshold, and a 4-fold increase in MFI.  Compare the approaches separately for the primary seroconversion and subsequent seroconversions

```{r indicators of seroconversion}
#----------------------------------
# Create a factor that summarizes 
# seroconversion status based 
# on seropositivity cutoff, with 4 categories:
# seroreversion, no change, 1st seroconversion, 
# 2nd or 3rd seroconversion
#----------------------------------
dl3 <- dl2 %>%
  mutate(seroinf=cut(seroin,breaks=c(-1,0,1,5),
                     labels=c("No change", 
                              "Primary seroconversion", 
                              "Secondary seroconversion")),
         serostat = seroinf
  )

dl3$serostat = factor(dl3$serostat,levels=c("Seroreversion",levels(dl3$seroinf))) 
dl3$serostat[dl3$seror==1] <- "Seroreversion"


#----------------------------------
# Create a factor that summarizes 
# seroconversion status based on a 4-fold increase in MFI
# with 4 categories:
# seroreversion, no change, 1st seroconversion, 
# 2nd to 4th seroconversion
#----------------------------------
dl3 <- dl3 %>%
  mutate(seroinf4fold=cut(seroin4fold,breaks=c(-1,0,1,5),
                     labels=c("No change", 
                              "Primary seroconversion", 
                              "Secondary seroconversion")),
         serostat4fold = seroinf4fold
  )

dl3$serostat4fold = factor(dl3$serostat4fold,levels=c("Seroreversion",levels(dl3$seroinf4fold))) 
dl3$serostat4fold[dl3$seror4fold==1] <- "Seroreversion"
```

Compare classification of incident seroconversions
```{r seroconversion agreement}
dclass <- foreach(ab=levels(dl3$antigenf),.combine=rbind) %do% {
  dcompi <- dl3 %>% filter(antigenf==ab)
  tabi <- as.vector(table(dcompi$seroi,dcompi$seroi4fold))
  names(tabi) <- c("sp04f0","sp14f0","sp04f1","sp14f1")
  kappai <- psych::cohen.kappa(table(dcompi$seroi,dcompi$seroi4fold))$kappa
  tabi <- data.frame(t(tabi))
  tabi <- tabi %>%
    mutate(Nobs=sp04f0+sp14f0+sp04f1+sp14f1,
           agreement=(sp04f0+sp14f1)/Nobs,
           kappa=kappai,
           antigenf=ab) %>%
    select(antigenf,Nobs,everything())
  tabi

}


knitr::kable(dclass,digits=3,caption="Classification agreement of seroconversion by based on seropositivity cutoffs and a 4-fold increase in MFI",
             col.names = c("Antigen","N", "Seropos 0\n4-fold incr. 0","Seropos 1\n4-fold incr. 0","Seropos 0\n4-fold incr. 1","Seropos 1\n4-fold incr. 1","Agreement","Kappa")) %>%
  kable_styling(bootstrap_options = "striped",full_width = TRUE)


```

# Distribution of change in MFI

## Distribution of changes in MFI by seroconversion status
```{r change in MFI by episode 4-fold def, fig.width=7,fig.height=7}

dl3plot <- dl3
                          
# calculate Ns in each seroconversion category to print in the figure
epiNs <- dl3plot %>%
  group_by(pathogen,antigenf,serostat4fold) %>%
  summarize(n=n()) %>%
  filter(!is.na(serostat4fold)) %>%
  mutate(nlab=paste("(n=",n,")",sep=""))


# custom color blind color palette is in the preamble chunck
pcols <- c(cred,corange,cgreen,cteal,cblue,cmagent)
# pcols <- c(corange,cred,cmagent,cblue,cteal,cgreen)

set.seed(123)
diffmfi_byepisode <- ggplot(data=filter(dl3plot,!is.na(serostat4fold)),aes(x=serostat4fold,y=logmfi_dlag,fill=pathogen,color=pathogen)) +
  facet_wrap(~antigenf,nrow=6,ncol=4)+ 
  geom_jitter(position = position_jitter(width = 0.2), alpha = 0.2,pch=19,size=0.5) +
  geom_boxplot(notch=FALSE,outlier.color=NA,width=0.5,alpha=0.5,fill=NA,color="grey20")+
  geom_hline(aes(yintercept = log10(4)),lty="dashed") +
  geom_hline(aes(yintercept = -log10(4)),lty="dashed") +

  geom_hline(aes(yintercept = 0)) +
  # add Ns
  geom_text(data=epiNs,aes(x=serostat4fold,y=-3.25,label=nlab),col="gray40",size=2) +
  scale_y_continuous(expand=c(0,0),limits=c(-3.5,4.5),breaks=-3:4) +
  scale_fill_manual(values=pcols)+
  scale_color_manual(values=pcols)+
  labs(title="",x="Seroconversion status based on 4-fold change in MFI",y="Change in log10 Luminex Response (MFI-bg) between measurements") +
  theme_minimal(base_size=12)+
  theme(
    panel.grid.minor.x = element_blank(),
    axis.text.x=element_text(angle=45,hjust=1),
    legend.position="none"
  )

diffmfi_byepisode

# save a PDF version
ggsave(here("figs","haiti-ab-mfidiff-byepisode-4fold.pdf"),plot=diffmfi_byepisode,device="pdf",width=9,height=8)


```

## Distribution of changes in MFI by age category
```{r change in MFI by agecat, fig.width=7,fig.height=7}

# Difference in MFI values - boxplot
dl3plot2 <- dl2 %>%
  mutate(agecat=cut(age,breaks=c(0,0.99,1.99,2.99,3.99,4.99,12),labels=c("0","1","2","3","4","5+")))

# calculate Ns in each age category to print in the figure
ageNs <- dl3plot2 %>%
  group_by(pathogen,antigenf,agecat) %>%
  summarize(n=n()) %>%
  filter(!is.na(agecat)) %>%
  mutate(nlab=paste("(n=",n,")",sep=""))

# custom color blind color palette is in the preamble chunck
# pcols <- c(cred,corange,cgreen,cteal,cblue,cmagent)
pcols <- c(corange,cred,cmagent,cblue,cteal,cgreen)

diffmfi_byage <- ggplot(data=dl3plot2,aes(x=agecat,y=logmfi_dlead,fill=pathogen,color=pathogen)) +
  facet_wrap(~antigenf,nrow=6,ncol=4)+ # ,scales='free_y'
  geom_jitter(position = position_jitter(width = 0.2), alpha = 0.2,pch=19) +
  geom_boxplot(notch=FALSE,outlier.color=NA,width=0.5,alpha=0.5,fill=NA,color="grey20")+
  geom_hline(aes(yintercept = 0)) +
  geom_text(data=ageNs,aes(x=agecat,y=-3.75,label=nlab),col="gray40",size=2) +

  scale_y_continuous(expand=c(0,0),limits=c(-4,4.5),breaks=-3:4) +
  scale_fill_manual(values=pcols)+
  scale_color_manual(values=pcols)+
  labs(title="",x="Age at beginning of period (years completed)",y="Change in log10 Luminex Response (MFI-bg) between measurements") +
  theme_minimal(base_size=12)+
  theme(
    panel.grid.minor.x = element_blank(),
    legend.position="none"
  )

diffmfi_byage

# save a PDF version
ggsave(here("figs","haiti-ab-mfidiff-byage.pdf"),plot=diffmfi_byage,device="pdf",width=9,height=8)


```



# Session Info
```{r session info}
sessionInfo()
```




