---
title: Enteropathogen antibody dynamics and force of infection among children in low-resource settings
subtitle: Adjusting force of infection for sampling interval, Figure 7
output:
  html_document:
    highlight: haddock
    theme: default
    code_folding: hide
    df_print: paged
    toc: yes
    toc_depth: '3'
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Notebook Summary

In the Kenya and Haiti cohorts, plot empirical seroconversion rates alongside simulation results from scenarios that sampled cohorts with a 30-day interval. `SI-File8-sampling-interval-simulation.Rmd (.html)` includes a detailed description of the simulation and more detailed results from the full set of scenarios considered. 

# Script preamble
```{r preamble}
#-----------------------------
# preamble
#-----------------------------
library(here)
here::here()

# load packages
library(tidyverse)

# safe color blind palette
# http://jfly.iam.u-tokyo.ac.jp/color/
# http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

```

# Load rate estimates and simulation results

```{r load files,warning=FALSE}
#-----------------------------
# load the estimates
# created with 
# Fig6-asembo-incidence-analysis.Rmd
# Table2-haiti-incidence-analysis.Rmd
#-----------------------------
dk <- readRDS(here::here("output","asembo-enteric-ab-ests-seroprev-vs-rates.rds")) %>%
  mutate(country = "Kenya") %>%
  filter(method == "Seropositivity cutoff" & pathogen != "V. cholerae")

dh <- readRDS(here::here("output","haiti-enteric-ab-ests-seroprev-vs-rates.rds")) %>%
  mutate(country = "Haiti") %>%
  filter(method == "Seropositivity cutoff")

#-----------------------------
# load simulation results
# created with 
# SI-File8-sampling-interval-simulation.Rmd
#-----------------------------
ksim <- readRDS(here::here("output","asembo-imputation-simulation-results.rds")) %>%
  mutate(country = "Kenya")

hsim <- readRDS(here::here("output","haiti-imputation-simulation-results.rds")) %>%
  mutate(country = "Haiti")

dsim <- bind_rows(ksim,hsim) %>%
  ungroup() %>%
  mutate(antigenf = case_when(
    antigenf == "Giardia VSP-5" ~ "Giardia VSP-3 or VSP-5",
    antigenf == "Cryptosporidium Cp17" ~ "Cryptosporidium Cp17 or Cp23",
    antigenf == "E. histolytica LecA" ~ "E. histolytica LecA",
    antigenf == "Salmonella LPS group B" ~ "Salmonella LPS groups B or D",
    antigenf == "ETEC LT B subunit" ~ "ETEC LT B subunit",
    antigenf == "Campylobacter p18" ~ "Campylobacter p18 or p39",
    antigenf == "Norovirus GI.4" ~ "Norovirus GI.4",
    antigenf == "Norovirus GII.4.NO" ~ "Norovirus GII.4.NO")
  )

```

# Figure 7

**Figure caption:** Empirical seroconversion rates compared with estimates from 100 simulated datasets with daily resolution IgG trajectories that were sampled at a 30-day interval before estimating rates. Vertical lines in the simulation results indicate medians. **A**. In the Kenya cohort, children were aged 4-18 months and empirical IgG measurements were separated by 6 months (approximately 180 days). **B**. In the Haiti cohort, children were aged 0-11 years, and empirical IgG measurements were separated by approximately 1 year. 

```{r kenya empircal and simulation}

# sort pathogens by rates
kratelabs <- c("E. histolytica LecA",
              "Salmonella LPS groups B or D",
              "Giardia VSP-3 or VSP-5",
              "Cryptosporidium Cp17 or Cp23",
              "ETEC LT B subunit",
              "Campylobacter p18 or p39")
dkplot <- dk %>%
  mutate(antigenf=factor(antigenf,levels=kratelabs)) %>%
  arrange(antigenf)
dsim_kplot <- dsim %>%
  ungroup() %>%
  filter(country=="Kenya" & sinterval == "30") %>%
  mutate(antigenf=factor(antigenf,levels=kratelabs))


pcols <- cbPalette[c(6,7)]
kplot_foi <- ggplot(data=dsim_kplot,aes(x=antigenf,y=seroi)) +
  geom_jitter(width=0.2,alpha=0.5,col=cbPalette[7])+
  stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
                geom = "crossbar", width = 0.5,color="black",size=0.2)+
  geom_errorbar(data=dkplot,aes(ymin=seroi_lb,ymax=seroi_ub),width=0.2) +
  geom_point(data=dkplot,size=2.5)+
  labs(x="",y="seroconversion rate per year at risk (force of infection)",tag="A",title="Kenya") +
  scale_y_continuous(breaks=seq(0,16,by=2))+
  annotate("text",x=6.45,y=6,label="empirical")+
  annotate("text",x=4,y=11.5,label="simulation replicates\n(30 day sampling interval)",color=cbPalette[7])+
  coord_flip(ylim=c(0,16))+
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "none",
    axis.title.x = element_text(size=12)
        )

kplot_foi

ggsave(file=here::here("figs","Fig7a-asembo-foi-simulation.pdf"),plot=kplot_foi,device=cairo_pdf,height=4,width=8)
ggsave(file=here::here("figs","Fig7a-asembo-foi-simulation.TIFF"),plot=kplot_foi,device="tiff",height=4,width=8)

```

```{r haiti empircal and simulation}

# sort pathogens by rates
hratelabs <- c("E. histolytica LecA",
               "Norovirus GI.4",
               "Giardia VSP-3 or VSP-5",
               "Salmonella LPS groups B or D",
               "Cryptosporidium Cp17 or Cp23",
               "Norovirus GII.4.NO",
               "ETEC LT B subunit")
dhplot <- dh %>%
  mutate(antigenf=factor(antigenf,levels=hratelabs)) %>%
  arrange(antigenf)
dsim_hplot <- dsim %>%
  ungroup() %>%
  filter(country=="Haiti" & sinterval == "30") %>%
  mutate(antigenf=factor(antigenf,levels=hratelabs))


pcols <- cbPalette[c(6,7)]
hplot_foi <- ggplot(data=dsim_hplot,aes(x=antigenf,y=seroi)) +
  geom_jitter(width=0.2,alpha=0.5,col=cbPalette[7])+
  stat_summary(fun.y = median, fun.ymin = median, fun.ymax = median,
                geom = "crossbar", width = 0.5,color="black",size=0.2)+
  geom_errorbar(data=dhplot,aes(ymin=seroi_lb,ymax=seroi_ub),width=0.2) +
  geom_point(data=dhplot,size=2.5)+
  labs(x="",y="seroconversion rate per year at risk (force of infection)",tag="B",title="Haiti") +
  scale_y_continuous(breaks=seq(0,16,by=2))+
  annotate("text",x=7.45,y=1.1,label="empirical")+
  annotate("text",x=6,y=6.2,label="simulation replicates\n(30-day sampling interval)",color=cbPalette[7])+
  coord_flip(ylim=c(0,10.5))+
  theme_minimal(base_size = 16) +
  theme(
    legend.position = "none",
    axis.title.x = element_text(size=12)
        )

hplot_foi

ggsave(file=here::here("figs","Fig7b-haiti-foi-simulation.pdf"),plot=hplot_foi,device=cairo_pdf,height=4,width=8)
ggsave(file=here::here("figs","Fig7b-haiti-foi-simulation.TIFF"),plot=hplot_foi,device="tiff",height=4,width=8)

```

# Session Info
```{r session info}
sessionInfo()
```




