---
title: "Haiti enteric antibody analysis of age-dependent antibody curves"
output: 
  html_notebook:
    theme: default
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

#Summary
This script estimates age-dependent mean antibody curves using cubic splines (in a GAM model from the `mgcv` package) and using ensemble machine learning (using super learning from the `tmleAb` package). 

# Script preamble
```{r preamble}
#-----------------------------
# preamble
#-----------------------------
rm(list=ls())
library(tidyverse)
library(kableExtra) # devtools::install_github("haozhu233/kableExtra")
library(mixtools)

options(knitr.table.format = "html") 

# set up for parallel computing
# configure for a laptop (use only 3 cores)
library(foreach)
library(doParallel)
registerDoParallel(cores=3)

# bright color blind palette:  https://personal.sron.nl/~pault/ 
cblack <- "#000004FF"
cblue <- "#3366AA"
cteal <- "#11AA99"
cgreen <- "#66AA55"
cchartr <- "#CCCC55"
cmagent <- "#992288"
cred <- "#EE3333"
corange <- "#EEA722"
cyellow <- "#FFEE33"
cgrey <- "#777777"

```


#Load the formatted data

```{r load data}
#-----------------------------
# load the formatted data
# with seropositivity indicators
# created with:
# haiti-enteric-ab-data-format.Rmd -->
# haiti-enteric-ab-distributions.Rmd
#-----------------------------
dl <- readRDS("~/enterics-immuno-epi/data/haiti_analysis2.rds")

# list the enteric antigens and formatted labels for them
# re-arrange the order slightly compared to other figures 
# just to acommodate plot aesthetics
mbavars <- c("vsp3","vsp5","cp17","cp23","leca","etec","salb","sald","norogi","norogii")

mbalabs <- c("Giardia VSP-3","Giardia VSP-5","Cryptosporidium Cp17","Cryptosporidium Cp23","E. histolytica LecA","ETEC toxin beta subunit","Salmonella LPS B","Salmonella LPS D","Norovirus GI", "Norovirus GII")

dl <- dl %>%
  mutate(antigenf=factor(antigenf,levels=mbalabs),
         pathogen=factor(pathogen,levels=c("Giardia","Cryptosporidium","E. histolytica","ETEC","Salmonella","Norovirus")))

```

#Estimate age-dependent mean antibody levels

##Estimate curves using splines in a generalized additive model (GAM)

```{r age-curve mean gam }
#----------------------------------
# simulataneous CIs for GAMs
# estimated by resampling the 
# Baysian posterior estimates of
# the variance-covariance matrix
# assuming that it is multivariate normal
# the function below also estimates 
# the unconditional variance-covariance
# matrix, Vb=vcov(x,unconditional=TRUE), 
# which allows for undertainty in the actual
# estimated mean as well 
# (Marra & Wood 2012 Scandinavian Journal of Statistics, 
#  Vol. 39: 53–74, 2012, doi: 10.1111/j.1467-9469.2011.00760.x )
# simultaneous CIs provide much better coverage than pointwise CIs
# see: http://www.fromthebottomoftheheap.net/2016/12/15/simultaneous-interval-revisited/
#----------------------------------

gamCI <- function(m,newdata,nreps=10000) {
  require(mgcv)
  require(dplyr)
  Vb <- vcov(m,unconditional = TRUE)
  pred <- predict(m, newdata, se.fit = TRUE)
  fit <- pred$fit
  se.fit <- pred$se.fit
  BUdiff <- MASS::mvrnorm(n=nreps, mu = rep(0, nrow(Vb)), Sigma = Vb)
  Cg <- predict(m, newdata, type = "lpmatrix")
  simDev <- Cg %*% t(BUdiff)
  absDev <- abs(sweep(simDev, 1, se.fit, FUN = "/"))
  masd <- apply(absDev, 2L, max)
  crit <- quantile(masd, prob = 0.95, type = 8)
  pred <- data.frame(newdata,fit=pred$fit,se.fit=pred$se.fit)
  pred <- mutate(pred,
                 uprP = fit + (2 * se.fit),
                 lwrP = fit - (2 * se.fit),
                 uprS = fit + (crit * se.fit),
                 lwrS = fit - (crit * se.fit)
  )
  return(pred)
}


# fit cubic spline over age, separately for each antigen
# estimate approximate simultaneous confidence intervals
# for the fits
library(mgcv)
gamfits <- foreach(ab=levels(dl$antigenf),.combine=rbind) %dopar% {
    pd <- filter(dl, antigenf==ab)
    gfit <- gam(logmfi~s(age, bs="cr"),data=pd)
    gsci <- gamCI(m=gfit,newdata=pd,nreps=1000)
    gsci$antigen <- ab
    return(gsci)
  }
```

##Estimate curves using ensemble machine learning
```{r age-curve mean ensemble, echo=TRUE, message=FALSE, warning=FALSE}
#-----------------------------
# estimate curves with 
# ensemble machine learning
# including a library with
# GLM, GAM, Loess, and 
# extreme gradient boosting
#-----------------------------

# detach the mgcv package because
# the ensemble algorithm uses gam()
# from the gam package and there are
# conflicts between them
detach("package:mgcv", unload=TRUE)

# load the tmleAb package, at this time
# only available through github: 
# devtools::install_github("ben-arnold/tmleAb")
library(tmleAb)
set.seed(123)
slfits <-  foreach(ab=levels(dl$antigenf),.combine=rbind) %dopar% {
  pd <- filter(dl, antigenf==ab)
  slfit <- agecurveAb(Y=pd$logmfi,
                      Age=pd$age,
                      id=pd$id,
                      SL.library = c("SL.glm","SL.gam","SL.loess","SL.xgboost"))
  sld <- data.frame(age=slfit$Age,fit=slfit$pY)
  sld <- sld %>% group_by(age) %>% filter(row_number()==1) %>% ungroup()
  slres <- left_join(pd,sld,by=c("age"))
  return(slres)
}

#-----------------------------
# estimate means by age in 
# years (0-9)
# to get 95% CIs through the
# influence curve
# there are 19 measurements
# among children 10-12 that
# are excluded for these estimates
# due to small sample sizes
#-----------------------------

tmlemeans <- foreach(ab=levels(dl$antigenf),.combine=rbind) %:%
  foreach(agei=0:9,.combine=rbind) %dopar% {
    pd <- filter(dl,antigenf==ab & floor(age)==agei)
    mu <- tmleAb(Y=pd$logmfi,id=pd$id,SL.library="SL.mean")
    di <- data.frame(pathogen=pd$pathogen[1],antigenf=ab,age=agei,mean=mu$psi,se=mu$se,lb=mu$lb,ub=mu$ub)
    di
  }


```

##Plot age-dependent mean antibody curves

Plot all of the estimates together. The spline fits are colored and the highly adaptive ensemble fits are grey. Empirical means with 95% CIs (averaged by year) are black. The approximate simultaneous confidence bands for the spline fit are shaded. The plots are trimed at age 10 years to reduce edge effects because there are only 19 measurements older than 10 years.

```{r age-curve mean figure all, fig.height=12, fig.width=6}

log10labs <- c( 
  expression(10^0),
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)

# custom color blind color palette is in the preamble chunck
pcols <- c(cred,corange,cgreen,cblue,cteal,cmagent)

pcurveall <- ggplot(data=gamfits,aes(x=age,group=pathogen,color=pathogen,fill=pathogen)) +
  # approximate simultaneous CI from spline fit
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA,fill=cgrey) +
  # actual measurements
  # geom_jitter(aes(y=logmfi),width=0.2,height=0,color="black",alpha=0.4,size=0.3) +
  # ensemble fit
  geom_line(data=slfits,aes(x=age,y=fit),color=cgrey) +
  # spline fit
  geom_line(aes(y=fit)) +
  # empirical means 95% CI from the influence curve (averaged for each age)
  geom_pointrange(data=tmlemeans,aes(x=age+0.5,y=mean,ymin=lb,ymax=ub),size=0.2,color=cblack)+
  # plot aesthetics
  facet_wrap(~antigenf,nrow=6,ncol=2) +
  scale_x_continuous(limits=c(0,10),breaks=seq(0,10,by=2))+
  scale_y_continuous(breaks=0:4,labels=log10labs)+
  coord_cartesian(ylim=c(0,4.5))+
  labs(x="Age in years",y="Luminex Response (MFI-bg)") +
  scale_fill_manual(values=pcols) +
  scale_color_manual(values=pcols) +
  theme_minimal() +
  theme(legend.position = "none")

pcurveall

# save a PDF version 
ggsave("~/enterics-immuno-epi/figs/haiti-ab-agecurves-mean-allestimates.pdf",plot=pcurveall,device="pdf",width=7,height=14)

```


Create a more parsimonius plot, that only includes the spline fits but also includes the actual points.
```{r age-curve mean figure, fig.height=12, fig.width=6}

log10labs <- c( 
  expression(10^0),
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)

# custom color blind color palette is in the preamble chunck
pcols <- c(cred,corange,cgreen,cblue,cteal,cmagent)

pcurve <- ggplot(data=gamfits,aes(x=age,group=pathogen,color=pathogen,fill=pathogen)) +
  # approximate simultaneous CI from spline fit
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA,fill=cgrey) +
  # actual measurements
  geom_jitter(aes(y=logmfi),width=0.2,height=0,alpha=0.4,size=0.3) +
  # ensemble fit
  # geom_line(data=slfits,aes(x=age,y=fit),color=cgrey) +
  # spline fit
  geom_line(aes(y=fit),color="black") +
  # plot aesthetics
  facet_wrap(~antigenf,nrow=6,ncol=2) +
  scale_x_continuous(limits=c(0,10),breaks=seq(0,10,by=2))+
  scale_y_continuous(breaks=0:4,labels=log10labs)+
  coord_cartesian(ylim=c(0,4.5))+
  labs(x="Age in years",y="Luminex Response (MFI-bg)") +
  scale_fill_manual(values=pcols) +
  scale_color_manual(values=pcols) +
  theme_minimal() +
  theme(legend.position = "none")

pcurve

# save a PDF version 
ggsave("~/enterics-immuno-epi/figs/haiti-ab-agecurves-mean.pdf",plot=pcurve,device="pdf",width=7,height=14)

```


#Estimate age-dependent seroprevalence

##Estimate seroprevalence curves using splines in a generalized additive model (GAM)

```{r seroprevalence and composite antigens, warning=FALSE}
#-----------------------------
# subset the data to antigens
# with seropositivity estimates
# hierarchy of information for 
# cutoffs:
# 1. ROC
# 2. mixture model based
# 3. estimated among presumed unexposed
#
# store the cutoff value used
# for figures
#-----------------------------
dlp <- dl %>%
  mutate(seropos=ifelse(!is.na(posroc),posroc,posmix),
         serocut=ifelse(!is.na(posroc),log10(roccut),mixcut),
         serocut_desc=ifelse(!is.na(posroc),"ROC","Mixture Model")) %>%
  mutate(seropos=ifelse(!is.na(seropos),seropos,posunex),
         serocut_desc=ifelse(!is.na(serocut),serocut_desc,"Unexp dist"),
         serocut=ifelse(!is.na(serocut),serocut,unexpcut)) %>%
  filter(!is.na(seropos))

#-----------------------------
# create composite 
# seropositivity indicators
# that use information from
# multiple antigens
#-----------------------------

dgi <- dlp %>% 
  filter(antigen %in% c("vsp3","vsp5")) %>%
  select(id,sampleid,sdate,age,pathogen,antigen,seropos) %>%
  spread(antigen,seropos) %>%
  mutate(seropos=ifelse(vsp3==1|vsp5==1,1,0),
         antigen="vsp3vsp5",
         antigenf="Giardia VSP-3 + VSP-5") %>%
  select(-vsp3,-vsp5)

dcr <- dlp %>% 
  filter(antigen %in% c("cp17","cp23")) %>%
  select(id,sampleid,sdate,age,pathogen,antigen,seropos) %>%
  spread(antigen,seropos) %>%
  mutate(seropos=ifelse(cp17==1|cp23==1,1,0),
         antigen="cp17cp23",
         antigenf="Cryptosporidium Cp17 + Cp23") %>%
  select(-cp17,-cp23)

ds <- dlp %>% 
  filter(antigen %in% c("salb","sald")) %>%
  select(id,sampleid,sdate,age,pathogen,antigen,seropos) %>%
  spread(antigen,seropos) %>%
  mutate(seropos=ifelse(salb==1|sald==1,1,0),
         antigen="salbsald",
         antigenf="Salmonella LPS groups B + D") %>%
  select(-salb,-sald)



dlp <- bind_rows(dlp,dgi,dcr,ds) %>%
  mutate(antigenf=factor(antigenf,
                         levels=c("Giardia VSP-3 + VSP-5",
                                  "Cryptosporidium Cp17 + Cp23",
                                  "E. histolytica LecA",
                                  "ETEC toxin beta subunit",
                                  "Salmonella LPS groups B + D",
                                  "Norovirus GI", 
                                  "Norovirus GII")) ) %>%
  filter(!is.na(antigenf)) %>%
  select(pathogen,antigen,antigenf,
         id,sampleid,sdate,age,
         seropos) %>%
  arrange(pathogen,antigenf,id,sdate)
```


```{r age-curve seroprevalence gam }
#-----------------------------
# estimate seroprevalence curves with 
# splines
#-----------------------------

# fit cubic spline over age, separately for each antigen
# estimate approximate simultaneous confidence intervals
# for the fits
library(mgcv)
pgamfits <- foreach(ab=levels(dlp$antigenf),.combine=rbind) %dopar% {
    pd <- filter(dlp, antigenf==ab)
    gfit <- mgcv::gam(seropos~s(age, bs="cr"),data=pd,family="binomial")
    gsci <- gamCI(m=gfit,newdata=pd,nreps=1000)
    gsci$antigenf <- ab
    return(gsci)
}

# use antilogit to get back to prevalence from the logit link
pgamfits <- pgamfits %>%
  mutate(fit  = 1/(1+exp(-fit)),
         uprP = 1/(1+exp(-uprP)),
         lwrP = 1/(1+exp(-lwrP)),
         uprS = 1/(1+exp(-uprS)),
         lwrS = 1/(1+exp(-lwrS))) %>%
  mutate(antigenf=factor(antigenf,levels=levels(dlp$antigenf)))
```

##Estimate seroprevalence curves using ensemble machine learning
```{r age-curve seroprevalence ensemble, echo=TRUE, message=FALSE, warning=FALSE}
#-----------------------------
# estimate curves with 
# ensemble machine learning
# including a library with
# GLM, GAM, Loess, and 
# extreme gradient boosting
#-----------------------------

# detach the mgcv package because
# the ensemble algorithm uses gam()
# from the gam package and there are
# conflicts between them
detach("package:mgcv", unload=TRUE)

set.seed(12345)
pslfits <-  foreach(ab=levels(dlp$antigenf),.combine=rbind) %dopar% {
  pd <- filter(dlp, antigenf==ab)
  slfit <- agecurveAb(Y=pd$seropos,
                      Age=pd$age,
                      id=pd$id,
                      family="binomial",
                      SL.library = c("SL.glm","SL.gam","SL.xgboost"))
  sld <- data.frame(age=slfit$Age,fit=slfit$pY)
  sld <- sld %>% group_by(age) %>% filter(row_number()==1) %>% ungroup()
  slres <- left_join(pd,sld,by=c("age"))
  return(slres)
}

#-----------------------------
# estimate prevalence by age in 
# years (0-9)
# since prevalence is very sparse
# in some strata, cannot use the 
# tmleAb() function above for
# influence-curve based SEs
# because it uses cross-validation
# and sometimes there are folds
# with no events. So, instead
# use an exact binomial approach
#-----------------------------
exactprev <- function(x) {
  # x : a binary indicator of the outcome (1/0)
  tabx <- table(x)
  if(length(tabx)<2) {
    if(names(tabx)=="1") {
      tabx <- c(0,tabx)
    } else{
      tabx <- c(tabx,0)
    }
  } 
  estx <- binom.test(x=tabx[2],n=sum(tabx))
  res <- c(estx$parameter,estx$statistic,estx$estimate,estx$conf.int)
  names(res) <- c("N","n","mean","min95","max95")
  return(res)
}

prevests <- foreach(ab=levels(dlp$antigenf),.combine=rbind) %:%
  foreach(agei=0:9,.combine=rbind) %dopar% {
      pd <- filter(dlp,antigenf==ab & floor(age)==agei)
      mus <- exactprev(pd$seropos)
      res <- data.frame(pathogen=pd$pathogen[1],antigenf=ab,age=agei,N=mus[1],n=mus[2],mean=mus[3],lb=mus[4],ub=mus[5])
      res
    }
   


```

##Plot age-dependent seroprevalence curves

Plot all of the estimates together. The spline fits are colored and the highly adaptive ensemble fits are grey. Empirical means with 95% CIs (averaged by year) are black. The approximate simultaneous confidence bands for the spline fit are shaded. The plots are trimed at age 10 years to reduce edge effects because there are only 19 measurements older than 10 years.

```{r age-curve seroprevalence figure all, fig.height=9, fig.width=6}

# bright color-blind palette defined in an earlier code chunk
pcols <- c(cred,corange,cgreen,cteal,cblue,cmagent)

ppcurve <- ggplot(data=pgamfits,aes(x=age,group=pathogen,color=pathogen,fill=pathogen)) +
  # approximate simultaneous CI from spline fit
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA,fill=cgrey) +
  # actual measurements
  # geom_jitter(aes(y=posmix),width=0.2,height=0,color="black",alpha=0.4,size=0.3) +
   # ensemble fit
  geom_line(data=pslfits,aes(x=age,y=fit),color=cgrey) +
  # spline fit
  geom_line(aes(y=fit)) +
  # empirical means 95% CI from the influence curve (averaged for each age)
  geom_pointrange(data=prevests,aes(x=age+0.5,y=mean,ymin=lb,ymax=ub),size=0.2,color=cblack)+
  # plot aesthetics
  facet_wrap(~antigenf,nrow=6,ncol=2) +
  scale_x_continuous(limits=c(0,10),breaks=seq(0,10,by=2))+
  scale_y_continuous(breaks=seq(0,1,by=0.2),labels=sprintf("%1.0f",seq(0,1,by=0.2)*100))+
  coord_cartesian(ylim=c(0,1))+
  labs(x="Age in years",y="Seroprevalence (%)") +
  scale_fill_manual(values=pcols) +
  scale_color_manual(values=pcols) +
  theme_minimal() +
  theme(legend.position = "none")

ppcurve

# save a PDF version
ggsave("~/enterics-immuno-epi/figs/haiti-ab-agecurves-seroprev-allests.pdf",plot=ppcurve,device="pdf",width=7,height=14)

```

Create a more parsimonius plot, that only includes the spline fits but also includes the actual points.

```{r age-curve seroprevalence figure, fig.height=9, fig.width=6}

  # set ETEC confidence band to NA after it flatlines at 100%
dplot <- pgamfits %>%
  mutate(lwrS=ifelse(antigen=="etec" & fit==1,NA,lwrS))


# bright color-blind palette defined in an earlier code chunk
pcols <- c(cred,corange,cgreen,cteal,cblue,cmagent)

ppcurve2 <- ggplot(data=dplot,aes(x=age,group=pathogen,color=pathogen,fill=pathogen)) +
  # approximate simultaneous CI from spline fit
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA) +
  # actual measurements
  geom_point(aes(y=seropos),color="black",alpha=0.4,size=0.3) +
  # spline fit
  geom_line(aes(y=fit)) +
  # plot aesthetics
  facet_wrap(~antigenf,nrow=6,ncol=2) +
  scale_x_continuous(limits=c(0,10),breaks=seq(0,10,by=2))+
  scale_y_continuous(breaks=seq(0,1,by=0.2),labels=sprintf("%1.0f",seq(0,1,by=0.2)*100))+
  coord_cartesian(ylim=c(0,1))+
  labs(x="Age in years",y="Seroprevalence (%)") +
  scale_fill_manual(values=pcols) +
  scale_color_manual(values=pcols) +
  theme_minimal() +
  theme(legend.position = "none")

ppcurve2

# save a PDF version
ggsave("~/enterics-immuno-epi/figs/haiti-ab-agecurves-seroprev.pdf",plot=ppcurve2,device="pdf",width=7,height=14)

```

## Numerical seroprevalence estimates to report in the manuscript. 


```{r seroprev numerical estimates}
# grab predicted means from cubic splines very close to each year of age
pests <- pgamfits %>%
  mutate(agey = floor(age)) %>%
  filter(age+0.02 > agey & age-0.02 < agey) %>%
  group_by(antigenf) %>%
  filter(!duplicated(agey)) %>%
  arrange(antigenf,agey) %>%
  select(antigenf,agey,prev=fit,prev_min95=lwrS,prev_max95=uprS)

pests


```

# Save fitted estimates for later use
```{r save estimates}
# spline fits of geometric means
saveRDS(gamfits,file="~/enterics-immuno-epi/output/haiti-enteric-ab-gamfits-means.rds")
# ensemble fits of geometric means
saveRDS(slfits,file="~/enterics-immuno-epi/output/haiti-enteric-ab-slfits-means.rds")
# spline fits of age-dependent seroprevalence
saveRDS(pgamfits,file="~/enterics-immuno-epi/output/haiti-enteric-ab-gamfits-seroprev.rds")
# spline fits of age-dependent seroprevalence
saveRDS(pslfits,file="~/enterics-immuno-epi/output/haiti-enteric-ab-slfits-seroprev.rds")
```

# Session Info
```{r session info}
sessionInfo()
```

