---
title: "Haiti enteric antibody analysis of longitudinal change"
output: 
  html_document:
    highlight: haddock
    theme: default
    code_folding: show
    df_print: paged
    toc: yes
    toc_depth: '3'
    toc_float:
      collapsed: yes
      smooth_scroll: yes
  html_notebook:
    theme: default
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Summary

This script summarizes the change in antibody levels within individual children and also estimates the incidence proportion of seroconversion and seroreversion over different age windows.


# Script preamble
```{r preamble}
#-----------------------------
# preamble
#-----------------------------
library(here)
here()

# load packages
library(tidyverse)

# set up for parallel computing
# configure for a laptop (use only 3 cores)
library(foreach)
library(doParallel)
registerDoParallel(cores=3)

# bright color blind palette:  https://personal.sron.nl/~pault/ 
cblack <- "#000004FF"
cblue <- "#3366AA"
cteal <- "#11AA99"
cgreen <- "#66AA55"
cchartr <- "#CCCC55"
cmagent <- "#992288"
cred <- "#EE3333"
corange <- "#EEA722"
cyellow <- "#FFEE33"
cgrey <- "#777777"

```


```{r load data}
#-----------------------------
# load the formatted data
# created with 
# asembo-enteric-ab-data-format.Rmd
#-----------------------------
dl <- readRDS(here("data","haiti_analysis2.rds"))

# list the enteric antigens and formatted labels for them
mbavars <- c("vsp3","vsp5","cp17","cp23","leca","etec","salb","sald","norogi","norogii")

mbalabs <- c("Giardia VSP-3","Giardia VSP-5","Cryptosporidium Cp17","Cryptosporidium Cp23","E. histolytica LecA","ETEC toxin beta subunit","Salmonella LPS B","Salmonella LPS D","Norovirus GI", "Norovirus GII")
```

# Identify incident changes
Among children, identify those that changed status between enrollment and follow-up. Those who changed from negative to positive are seroconverters (`seroi` below), and those who changed from positive to negative are seroreverters (`seror` below).


```{r long incidence}

#-----------------------------
# identify seropositive measures
# hierarchy of information for 
# cutoffs:
# 1. ROC
# 2. mixture model based
# 3. estimated among presumed unexposed
#
# store the cutoff value used
# for figures
#-----------------------------
dl <- dl %>%
  mutate(seropos=ifelse(!is.na(posroc),posroc,posmix),
         serocut=ifelse(!is.na(posroc),roccut,mixcut),
         serocut_desc=ifelse(!is.na(posroc),"ROC","Mixture Model")) %>%
  mutate(seropos=ifelse(!is.na(seropos),seropos,posunex),
         serocut_desc=ifelse(!is.na(serocut),serocut_desc,"Unexp dist"),
         serocut=ifelse(!is.na(serocut),serocut,unexpcut)) 


#-----------------------------
# identify incident 
# seroconversions and reversions
#-----------------------------

# group the data by child and
# use lags to identify
# time in years between measurements,
# sero-conversions + sero-reversions 
# between measurements
# set the first measurement to 
# missing for the incidence indicators
# (using the is.na(age_diff) to identify them)
dl2 <- dl %>%
  group_by(antigen,id) %>% 
  arrange(antigen,id,sdate,sampleid) %>%
  mutate(age_min  = min(age),
         age_diff = age - lag(age),
         
         logmfi_lag  = lag(logmfi),
         logmfi_lead = lead(logmfi),
         logmfi_dlag  = logmfi - logmfi_lag,
         logmfi_dlead = logmfi_lead - logmfi,
         
         # incident seroconversions and reversions
         # including cumulative numbers
         seropos_lag  = lag(seropos),
         seroi = ifelse(seropos==1 & seropos_lag==0,1,0),
         seroi = ifelse(is.na(age_diff),NA,seroi),
         seroin = cumsum(ifelse(is.na(seroi),0,seroi)),
         seroin = ifelse(seroi==1,seroin,0),
         seror = ifelse(seropos==0 & seropos_lag==1,1,0),
         seror = ifelse(is.na(age_diff),NA,seror),
         serorn = cumsum(ifelse(is.na(seror),0,seror)),
         serorn = ifelse(seror==1,serorn,0)
         ) %>%
  select(
         id,sampleid,sdate,
         starts_with("age"),
         pathogen,antigen,antigenf,
         mfi,logmfi,logmfi_dlag,logmfi_dlead,
         roccut,mixcut,serocut,
         posroc,posmix,seropos,seroi,seroin,seror,serorn,
         -logmfi_lag,-logmfi_lead,-seropos_lag)


```

# Longitudinal child trajectories

```{r figure of individual MFI trajectories, echo=FALSE}

dl2plot <- dl2 %>% group_by(id)
dl2plot$sid <- group_indices(dl2plot)
dl2plot$sid <- factor(dl2plot$sid)

# color by MFI increases/decreases or seroconversion/reversion statsus
dl2plot <- dl2plot %>%
  mutate(
    logmfi_chng  = ifelse(logmfi_dlag>0,"Increasing","Decreasing"),
    logmfi_chng = ifelse(is.na(logmfi_chng) & logmfi_dlead>0,"Increasing",logmfi_chng),
    logmfi_chng = ifelse(is.na(logmfi_chng) & logmfi_dlead<0,"Decreasing",logmfi_chng),
    logmfi_chng = factor(logmfi_chng,levels=c("Increasing","Decreasing")),
    
    sero_chng = ifelse(lead(seroi)==1,"Seroconversion","No change"),
    sero_chng = ifelse(lead(seror)==1,"Seroreversion",sero_chng),
    sero_chng = ifelse(is.na(sero_chng),lead(sero_chng),sero_chng),
    sero_chng = factor(sero_chng,levels=c("Seroconversion","Seroreversion","No change"))
    
  )


# custom color blind color palette is in the preamble chunck
# pcols <- c(cred,corange,cgreen,cteal,cblue,cmagent)
pcols <- c(corange,cred,cmagent,cblue,cteal,cgreen)

# custom log labels
log10labs <- c( 
  expression(10^0),
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)

# plot all trajectories, colored (doesn't look that great)
# p <- ggplot(dl2plot, aes(x=age, y=logmfi, group=factor(id),color=sero_chng) ) +
#     facet_wrap(~antigenf,nrow=6,ncol=2)+
#     geom_line(alpha=0.1) +
#     guides(colour=FALSE) +
#     guides(alpha=FALSE) +
#     # geom_line(data=filter(dl2plot,sid==iid),aes(x=age,y=logmfi),size=1.2) +
#     scale_color_manual(values=c(corange,cteal,cgrey))+
#     xlab("Child Age (Years)") +
#     ylab("Luminex Response (MFI-bg)") +
#     coord_cartesian(ylim=c(0, 4.5))+
#     scale_y_continuous(breaks=0:4,labels=log10labs) +
#     scale_x_continuous(limits=c(0,10),breaks=seq(0,10,by=2)) +
#     theme_minimal() +
#     theme(
#       # strip.text = element_text(size=16),
#       # axis.text.x=element_text(size=12),
#       # axis.text.y=element_text(size=12),
#       # axis.title.x=element_text(size=14),
#       # axis.title.y=element_text(size=14),
#       legend.position="none"
#     )
# 
# p

# highlight a handfull of example individual trajectories
foreach(iid=c(20,24,30,34,35,38,43,47,72,76,77,119,135)) %do% {
  p <- ggplot(dl2plot, aes(x=age, y=logmfi, group=factor(id)) ) +
    facet_wrap(~antigenf,nrow=6,ncol=2)+
    geom_line(alpha=0.1,color=cgrey) +
    guides(colour=FALSE) +
    guides(alpha=FALSE) +
    geom_hline(aes(yintercept=serocut),linetype="dashed") +
    geom_line(data=filter(dl2plot,sid==iid),aes(x=age,y=logmfi,color=sero_chng)) +
    scale_color_manual(values=c(corange,cteal,cgrey))+
    xlab("Child Age (Years)") +
    ylab("Luminex Response (MFI-bg)") +
    coord_cartesian(ylim=c(0, 4.5))+
    scale_y_continuous(breaks=0:4,labels=log10labs) +
    scale_x_continuous(limits=c(0,10),breaks=seq(0,10,by=2)) +
    theme_minimal() +
    theme(
      legend.position="none"
    )

  ggsave(filename=here("figs",paste("haiti-spaghetti-",iid,".pdf",sep="")),plot = p,device='pdf',width=5,height=14)

}

```

# Analysis of change in MFI



## Fig 3 Distribution of changes in MFI by seroconversion status
```{r change in MFI by episode, fig.width=7,fig.height=7}

# Create a factor that summarizes seroconversion status, with 4 categories:
# seroreversion, no change, 1st seroconversion, 2nd or 3rd seroconversion
dl2plot <- dl2 %>%
  mutate(seroinf=cut(seroin,breaks=c(-1,0,1,5),
                     labels=c("No change", "1st\nseroconversion", "2nd or 3rd\nseroconversion")),
         serostat = seroinf
  )

dl2plot$serostat = factor(dl2plot$serostat,levels=c("Seroreversion",levels(dl2plot$seroinf))) 
dl2plot$serostat[dl2plot$seror==1] <- "Seroreversion"
                          
# calculate Ns in each seroconversion category to print in the figure
epiNs <- dl2plot %>%
  group_by(pathogen,antigenf,serostat) %>%
  summarize(n=n()) %>%
  filter(!is.na(serostat)) %>%
  mutate(nlab=paste("(n=",n,")",sep=""))


# custom color blind color palette is in the preamble chunck
# pcols <- c(cred,corange,cgreen,cteal,cblue,cmagent)
pcols <- c(corange,cred,cmagent,cblue,cteal,cgreen)

set.seed(123)
diffmfi_byepisode <- ggplot(data=filter(dl2plot,!is.na(serostat)),aes(x=serostat,y=logmfi_dlag,fill=pathogen,color=pathogen)) +
  facet_wrap(~antigenf,nrow=6,ncol=4)+ 
  geom_jitter(position = position_jitter(width = 0.2), alpha = 0.2,pch=19) +
  geom_boxplot(notch=FALSE,outlier.color=NA,width=0.5,alpha=0.5,fill=NA,color="grey20")+
  geom_hline(aes(yintercept = log10(2)),lty="dashed") +
  geom_hline(aes(yintercept = 0)) +
  # add Ns
  geom_text(data=epiNs,aes(x=serostat,y=-3.25,label=nlab),col="gray40",size=2) +
  scale_y_continuous(expand=c(0,0),limits=c(-3.5,4.5),breaks=-3:4) +
  scale_fill_manual(values=pcols)+
  scale_color_manual(values=pcols)+
  labs(title="",x="Seroconversion status",y="Change in log10 Luminex Response (MFI-bg) between measurements") +
  theme_minimal(base_size=12)+
  theme(
    panel.grid.minor.x = element_blank(),
    axis.text.x=element_text(angle=45,hjust=1),
    legend.position="none"
  )

diffmfi_byepisode

# save a PDF version
ggsave(here("figs","haiti-ab-mfidiff-byepisode.pdf"),plot=diffmfi_byepisode,device="pdf",width=9,height=8)


```

## Distribution of changes in MFI by age category
```{r change in MFI by agecat, fig.width=7,fig.height=7}

# Difference in MFI values - boxplot
dl2plot <- dl2 %>%
  mutate(agecat=cut(age,breaks=c(0,0.99,1.99,2.99,3.99,4.99,12),labels=c("0","1","2","3","4","5+")))

# calculate Ns in each age category to print in the figure
ageNs <- dl2plot %>%
  group_by(pathogen,antigenf,agecat) %>%
  summarize(n=n()) %>%
  filter(!is.na(agecat)) %>%
  mutate(nlab=paste("(n=",n,")",sep=""))

# custom color blind color palette is in the preamble chunck
# pcols <- c(cred,corange,cgreen,cteal,cblue,cmagent)
pcols <- c(corange,cred,cmagent,cblue,cteal,cgreen)

diffmfi_byage <- ggplot(data=dl2plot,aes(x=agecat,y=logmfi_dlead,fill=pathogen,color=pathogen)) +
  facet_wrap(~antigenf,nrow=6,ncol=4)+ # ,scales='free_y'
  geom_jitter(position = position_jitter(width = 0.2), alpha = 0.2,pch=19) +
  geom_boxplot(notch=FALSE,outlier.color=NA,width=0.5,alpha=0.5,fill=NA,color="grey20")+
  geom_hline(aes(yintercept = 0)) +
  geom_text(data=ageNs,aes(x=agecat,y=-3.75,label=nlab),col="gray40",size=2) +

  scale_y_continuous(expand=c(0,0),limits=c(-4,4.5),breaks=-3:4) +
  scale_fill_manual(values=pcols)+
  scale_color_manual(values=pcols)+
  labs(title="",x="Age at beginning of period (years completed)",y="Change in log10 Luminex Response (MFI-bg) between measurements") +
  theme_minimal(base_size=12)+
  theme(
    panel.grid.minor.x = element_blank(),
    legend.position="none"
  )

diffmfi_byage

# save a PDF version
ggsave(here("figs","haiti-ab-mfidiff-byage.pdf"),plot=diffmfi_byage,device="pdf",width=9,height=8)


```


# Session Info
```{r session info}
sessionInfo()
```




