---
title: Enteropathogen antibody dynamics and force of infection among children in low-resource settings
subtitle: Supplementary Information File 3. Joint distributions of antibody response.
output: 
  html_document:
    theme: default
    highlight: haddock
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Notebook summary

This notebook summarizes the joint distribution of antibody responses in each cohort. The first section creates main text Figure 2, which summarizes comparisons between different antigens for the same pathogen plus a comparison between ETEC and cholera, where the beta toxin subunit is known to elicit cross-reactivity.  Each scatter plot also includes the spearman's rank correlation coefficient ($\rho$).  After creating Figure 2, the notebook creates supplemental figures that include pairs plots for the joint distribution of every combination of enteric antibody responses in each cohort.

# Script preamble
```{r preamble}
#-----------------------------
# preamble
#-----------------------------
# set to local workspace
library(here)
here()

# load packages
library(tidyverse)
library(scales)
library(viridis)
library(ellipse)
library(RColorBrewer)

# set up for parallel computing
# configure for a laptop (use only 3 cores)
library(foreach)
library(doParallel)
registerDoParallel(cores=3)

# bright color blind palette:  https://personal.sron.nl/~pault/ 
cblack <- "#000004FF"
cblue <- "#3366AA"
cteal <- "#11AA99"
cgreen <- "#66AA55"
cchartr <- "#CCCC55"
cmagent <- "#992288"
cred <- "#EE3333"
corange <- "#EEA722"
cyellow <- "#FFEE33"
cgrey <- "#777777"

# custom log labels
log10labs <- c( 
  expression(10^0),
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)

```



# Load and format data
```{r load and format data}

#--------------------------------
# load the various datasets
#--------------------------------
dh <- readRDS(here("data","haiti_analysis.rds"))
dk <- readRDS(here("data","asembo_analysis.rds"))
dt <- readRDS(here("data","kongwa_analysis.rds"))

#--------------------------------
# subset to common variables
# and append
#--------------------------------
dh <- dh %>% 
  mutate(country="Haiti") %>%
  select(country,id,sid=sampleid,antigen,antigenf,logmfi)
dk <- dk %>% 
  mutate(country="Kenya",sid = ifelse(time=="A","1","2")) %>%
  select(country,id=childid,sid,antigen,antigenf,logmfi)
  
dt <- dt %>% 
  mutate(country="Tanzania",sid="1") %>%
  select(country,id,sid,antigen,antigenf,logmfi)

dall <- bind_rows(dh,dk,dt)

#--------------------------------
# create antigen groupings for
# comparisons
# drop obs not contributing
#--------------------------------
d <- dall %>%
  mutate(comp= ifelse(antigen %in% c("vsp3","vsp5"),"Giardia",NA),
         comp= ifelse(antigen %in% c("cp17","cp23"),"Cryptosporidium",comp),
         comp= ifelse(antigen %in% c("p18","p39"),"Campylobacter",comp),
         comp= ifelse(antigen %in% c("sald","salb"),"Salmonella",comp),
         comp= ifelse(antigen %in% c("cholera","etec"),"ETEC\nV. cholerae",comp),
         comp= ifelse(antigen %in% c("norogi","norogii"),"Norovirus",comp),
         comp= ifelse(country=="Haiti" & antigen =="etec",NA,comp),
         comp= factor(comp,levels=c("Giardia","Cryptosporidium","Campylobacter","Salmonella","ETEC\nV. cholerae","Norovirus"))
         ) %>% 
  mutate(xlab=ifelse(antigen %in% c("vsp3","vsp5"),"VSP-3",NA),
         ylab=ifelse(antigen %in% c("vsp3","vsp5"),"VSP-5",NA),
         xlab=ifelse(antigen %in% c("cp17","cp23"),"Cp17",xlab),
         ylab=ifelse(antigen %in% c("cp17","cp23"),"Cp23",ylab),
         xlab=ifelse(antigen %in% c("p18","p39"),"p18",xlab),
         ylab=ifelse(antigen %in% c("p18","p39"),"p39",ylab),
         xlab=ifelse(antigen %in% c("salb","sald"),"LPS Group B",xlab),
         ylab=ifelse(antigen %in% c("salb","sald"),"LPS Group D",ylab),
         xlab=ifelse(antigen %in% c("cholera","etec"),"Cholera toxin B subunit",xlab),
         ylab=ifelse(antigen %in% c("cholera","etec"),"ETEC LT B subunit",ylab),
         xlab=ifelse(antigen %in% c("norogi","norogii"),"GII.4.NO",xlab),
         ylab=ifelse(antigen %in% c("norogi","norogii"),"GI.4",ylab)
         ) %>%
  filter(!is.na(comp))

#--------------------------------
# for antigen each pair, 
# label it as "x" or "y" to 
# spread it to wide format
#--------------------------------
dw <- d %>%
  mutate(xy=ifelse(antigen %in% c("vsp3","cp17","sald","p18","cholera","norogii"),"x","y")) %>%
  select(country,comp,xlab,ylab,id,sid,xy,logmfi) %>%
  spread(xy,logmfi) %>%
  mutate(country=factor(country,levels=c("Haiti","Kenya","Tanzania")))

```
# Estimate correlations and smooths

```{r est correlations}
#--------------------------------
# estimate spearman's correlation
# within each country, comparison
#--------------------------------
dcorr <- dw %>%
  group_by(country,comp) %>%
  mutate(corxy=cor(x,y,method="spearman",use="pairwise.complete.obs") ) %>%
  summarize(corxy=max(corxy,na.rm=T))
```

```{r est trimmed smooths}
#--------------------------------
# estimate smooths, trimmed to
# drop the bottom and top 1% of
# data in each comparison
# to avoid edge effects
#--------------------------------
dsmooths <- foreach(countryi=levels(dw$country),.combine=rbind) %:%
                      foreach(compi=levels(dw$comp),.combine=rbind) %do% {
                        pd <- filter(dw,country==countryi & comp==compi)
                        if(nrow(pd)>0) {
                          xqs <- quantile(pd$x,probs=c(0.01,0.99),na.rm=TRUE)
                          newd <- data.frame(x=seq(xqs[1],xqs[2],by=0.01))
                          lfit <- loess(y~x,data=pd)
                          return(data.frame(country=countryi,comp=compi,x=newd,y=predict(lfit,newdata=newd)))
                        }
                        
  
}

```


# Figure 2
Summary composite figure, **Figure 2 in the main text**
```{r composite figure,fig.width=15,fig.height=15 }

vircols <- viridis(n=4,alpha=1,begin=0.2,end=0.97)
pcol <- vircols[2]

# grab labels
dlabs <- dw %>% select(country,comp,xlab,ylab) %>% group_by(country,comp) %>% slice(1)

complot <- ggplot(data=dw,aes(x=x,y=y)) +
  facet_grid(comp~country) +
  geom_point(pch=19,color=pcol,alpha=0.1) +
  geom_line(data=dsmooths,aes(x=x,y=y),col="black",size=1.2)+
  geom_text(data=dcorr,
            aes(x=0.5,y=4.3,label=paste("rho ==",sprintf("%1.2f",corxy)) ),
            parse=TRUE, col="black")   +
  geom_text(data=dlabs,aes(x=2.3,y=0.1,label=xlab),color="gray40",angle=0)+
  geom_text(data=dlabs,aes(x=0.1,y=2.3,label=ylab),color="gray40",angle=90)+
  scale_x_continuous(limits=c(0,4.6),breaks=0:4,labels = log10labs)+
  scale_y_continuous(limits=c(0,4.6),breaks=0:4,labels = log10labs)+
  coord_equal() +
  labs(x="Luminex Response (MFI-bg)",y="Luminex Response (MFI-bg)") +
  theme_minimal(base_size=12) +
  theme(
    strip.text.x=element_text(size=12),
    strip.text.y=element_text(size=12,angle=0),
    legend.position="none"
  )


complot

# save PDF and TIFF versions
ggsave(here("figs","Fig2-ab-scatter-composite.pdf"),plot=complot,device=cairo_pdf,width=13,height=15)
ggsave(here("figs","Fig2-ab-scatter-composite.TIFF"),plot=complot,device="tiff",width=13,height=15)
```

# Individual country pairs plots
The above figure was created as a synthesis across individual country pairs plots. Below, the script creates each pairs plot that shows the joint relationship between every combination of antigens in each cohort.

```{r pairs functions}
#----------------------------------
# correlation ellipse
#----------------------------------
myellipse<-function(x,y,...){
  maxx <- max(x,na.rm=TRUE)
  minx <- min(x,na.rm=TRUE)
  maxy <- max(y,na.rm=TRUE)
  miny <- min(y,na.rm=TRUE)
  midx <- (maxx+minx)/2
  midy <- (maxy+miny)/2
  corxy <- cor(x,y,method="spearman",use="pairwise.complete.obs")
  colgroup<-cut(corxy,breaks=seq(-0.1,1,length=20),labels=F)
  viridiscols <- viridis(20)
  cols<-viridiscols[colgroup]
  xyc <-sprintf("%1.2f",corxy)
  xyc[grep("NA",xyc)]<-""
  exy <- ellipse(corxy,centre=c(midx,midy),scale=c((maxx-minx)/6,(maxy-miny)/6))
  polygon(exy,col=alpha(cols,alpha=0.5))
  lines(exy)
  if(!is.na(corxy)) {
    if(corxy<0.8) {
      text(midx,midy,xyc,cex=0.8)
    } else{
      text(maxx,midy-((maxy-miny)/3),xyc,cex=0.8,adj=1)
    }
  }
  
}


#----------------------------------
# scatter plot with loess fit
# (trimmed to reduce edge effects)
#----------------------------------
scatterloess<-function(x,y,cex=0.4,...){
  ld <- data.frame(x,y)
  ld <- ld[complete.cases(ld),]
  if(nrow(ld)>0) {
    points(ld$x,ld$y,pch=19,cex=cex,col=alpha('black',alpha=0.2))
    viridiscols <- viridis(11)
    lfit <- loess(y~x,data=ld)
    xqs <- quantile(x,probs=c(0.01,0.99),na.rm=TRUE)
    px <- seq(xqs[1],xqs[2],by=0.01)
    py <- predict(lfit,newdata=data.frame(x=px))
    lines(px,py,col=viridiscols[1],lwd=1.5)
  }
  
}


```


## Haiti
```{r haiti pairs plot,fig.width=7,fig.height=7}

# list the enteric antigens in Haiti and formatted labels for them
mbavars <- c("vsp3","vsp5","cp17","cp23","leca","salb","sald","etec","norogi","norogii")
mbalabs <- c("Giardia\nVSP-3","Giardia\nVSP-5","Cryptosporidium\nCp17","Cryptosporidium\nCp23","E. histolytica\nLecA","Salmonella\nLPS B","Salmonella\nLPS D","ETEC\nLT B subunit","Norovirus\nGI", "Norovirus\nGII")

hmat <- dall %>%
  filter(country=="Haiti") %>%
  select(id,sid,antigen,logmfi) %>%
  spread(antigen,logmfi) 

pairs(hmat[mbavars],labels=mbalabs,cex=0.1,las=1, 
      upper.panel=scatterloess,
      lower.panel=myellipse
)

```

## Kenya
```{r kenya pairs plot,fig.width=7,fig.height=7}
# list the enteric antigens in Asembo Kenya and formatted labels for them
mbavars <- c("vsp3","vsp5","cp17","cp23","leca","salb","sald","etec","cholera","p18","p39")

mbalabs <- c("Giardia\nVSP-3","Giardia\nVSP-5","Cryptosporidium\nCp17","Cryptosporidium\nCp23","E. histolytica\nLecA","Salmonella\nLPS B","Salmonella\nLPS D","ETEC\nLT B subunit","Cholera\ntoxin B subunit","Campylobacter\np18","Campylobacter\np39")

kmat <- dall %>%
  filter(country=="Kenya") %>%
  select(id,sid,antigen,logmfi) %>%
  spread(antigen,logmfi)

pairs(kmat[mbavars],labels=mbalabs,cex=0.1,las=1,
      upper.panel=scatterloess,
      lower.panel=myellipse
)

```

## Tanzania

There are a few blank panels in this figure. The reason is that some antigens were included only in year 1, and the cholera beta toxin was only included in years 2-4.  Table 1 and the Methods of the article include additional details.
```{r tanzania pairs plot,fig.width=7,fig.height=7}

# list the enteric antigens in Kongwa, Tanzania and formatted labels for them
mbavars <- c("vsp3","vsp5","cp17","cp23","leca","salb","sald","etec","cholera","p18","p39")

mbalabs <- c("Giardia\nVSP-3","Giardia\nVSP-5","Cryptosporidium\nCp17","Cryptosporidium\nCp23",
             "E. histolytica\nLecA","Salmonella\nLPS B","Salmonella\nLPS D","ETEC\nLT B subunit","Cholera\ntoxin B subunit","Campylobacter\np18","Campylobacter\np39")

tmat <- dall %>%
  filter(country=="Tanzania") %>%
  select(id,sid,antigen,logmfi) %>%
  spread(antigen,logmfi)
pairs(tmat[mbavars],labels=mbalabs,cex=0.1,las=1,
      upper.panel=scatterloess,
      lower.panel=myellipse
)

```

# Session Info
```{r session info}
sessionInfo()
```

