---
title: Enteropathogen antibody dynamics and force of infection among children in low-resource settings
subtitle: Supplementary Information File 1. No effect of intervention or bead lot on enteropathogen antibody response in Kenya and Tanzania.
output: 
  html_document:
    highlight: haddock
    theme: default
    code_folding: show
    df_print: paged
    toc: yes
    toc_depth: '3'
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Notebook summary

This notebook compares antibody measurements among children assigned to different treatment arms in the Asembo, Kenya household water filter trial and in the Kongwa, Tanzania annual azithromycin distribution trial. For additional details on the interventions and design, please see the main text and previously published articles related to the trials:

Morris, J. F., Murphy, J., Fagerli, K., Schneeberger, C., Jaron, P., Moke, F., Juma, J., Ochieng, J. B., Omore, R., Roellig, D., Xiao, L., Priest, J. W., Narayanan, J., Montgomery, J. M., Hill, V., Mintz, E., Ayers, T. L. & O’Reilly, C. E. A Randomized Controlled Trial to Assess the Impact of Ceramic Water Filters on Prevention of Diarrhea and Cryptosporidiosis in Infants and Young Children-Western Kenya, 2013. _Am. J. Trop. Med. Hyg._ 98, 1260–1268 (2018). https://www.ncbi.nlm.nih.gov/pubmed/29611500

Wilson, N., Goodhew, B., Mkocha, H., Joseph, K., Bandea, C., Black, C., Igietseme, J., Munoz, B., West, S. K., Lammie, P., Kasubi, M. & Martin, D. L. Evaluation of a Single Dose of Azithromycin for Trachoma in Low-Prevalence Communities. _Ophthalmic Epidemiol._ 26, 1–6 (2019). https://www.ncbi.nlm.nih.gov/pubmed/30543311

The results show that the interventions did not effect enteropathogen antibody response. Based on these results, we pooled data across treatment arms in all analyses.

The final section of this notebook stratifies the analyses from Tanzania in years 1 versus years 2-4. The Tanania study  used different bead lots in year 1 versus years 2-4 of the study. Different bead lots could theoretically have different antigen coupling efficiency, which could introduce differences in the MFI-bg measurements, since they are on an arbitrary scale that depends on how much antibody from the sample binds to the beads. Results of the stratified analyses show that the patterns observed were qualitatively and quantitatively similar for year 1 (n=902) versus years 2-4 (n=4,087). Owing to the much larger number of samples from later years, the results of this analysis also show that the pooled results match almost exactly the estimates from years 2-4.

# Script preamble
```{r preamble}
#-----------------------------
# preamble
#-----------------------------
library(here)
here()

# load packages
library(tidyverse)
library(tmleAb) # development package: devtools::install_github("ben-arnold/tmleAb")
library(mgcv)
library(knitr)
library(kableExtra)
  options(knitr.table.format = "html") 

# set up for parallel computing
# configure for a laptop (use only 3 cores)
library(foreach)
library(doParallel)
registerDoParallel(cores=3)


# bright color blind palette:  https://personal.sron.nl/~pault/ 
cblack <- "#000004FF"
cblue <- "#3366AA"
cteal <- "#11AA99"
cgreen <- "#66AA55"
cchartr <- "#CCCC55"
cmagent <- "#992288"
cred <- "#EE3333"
corange <- "#EEA722"
cyellow <- "#FFEE33"
cgrey <- "#777777"

```


#Asembo, Kenya Intervention Effects


## Load the data and format
```{r load data}
#-----------------------------
# load the formatted data
# with seropositivity indicators
# created with:
# asembo-enteric-ab-data-format.Rmd -->
# asembo-enteric-ab-distributions.Rmd
#-----------------------------
dl <- readRDS(here("data","asembo_analysis2.rds"))

# list the enteric antigens and make formatted labels for them
mbavars <- c("vsp3","vsp5","cp17","cp23","leca","salb","sald","p18","p39","etec","cholera")

mbalabs <- c("Giardia VSP-3","Giardia VSP-5","Cryptosporidium Cp17","Cryptosporidium Cp23","E. histolytica LecA","Salmonella LPS B","Salmonella LPS D","Campylobacter p18","Campylobacter p39","ETEC LT β subunit","Cholera toxin β subunit")
```

```{r composite antigens, warning=FALSE}
#-----------------------------
# create composite 
# seropositivity indicators
# that use information from
# multiple antigens
#-----------------------------

dgi <- dl %>% 
  filter(antigen %in% c("vsp3","vsp5")) %>%
  select(childid,time,tr,sex,age,agediff,pathogen,antigen,seropos) %>%
  spread(antigen,seropos) %>%
  mutate(seropos=ifelse(vsp3==1|vsp5==1,1,0),
         antigen="vsp3vsp5",
         antigenf="Giardia VSP-3 or VSP-5") %>%
  select(-vsp3,-vsp5)

dcr <- dl %>% 
  filter(antigen %in% c("cp17","cp23")) %>%
  select(childid,time,tr,sex,age,agediff,pathogen,antigen,seropos) %>%
  spread(antigen,seropos) %>%
  mutate(seropos=ifelse(cp17==1|cp23==1,1,0),
         antigen="cp17cp23",
         antigenf="Cryptosporidium Cp17 or Cp23") %>%
  select(-cp17,-cp23)

ds <- dl %>% 
  filter(antigen %in% c("salb","sald")) %>%
  select(childid,time,tr,sex,age,agediff,pathogen,antigen,seropos) %>%
  spread(antigen,seropos) %>%
  mutate(seropos=ifelse(salb==1|sald==1,1,0),
         antigen="salbsald",
         antigenf="Salmonella LPS groups B or D") %>%
  select(-salb,-sald)

dcp <- dl %>% 
  filter(antigen %in% c("p18","p39")) %>%
  select(childid,time,tr,sex,age,agediff,pathogen,antigen,seropos) %>%
  spread(antigen,seropos) %>%
  mutate(seropos=ifelse(p18==1|p39==1,1,0),
         antigen="p18p39",
         antigenf="Campylobacter p18 or p39") %>%
  select(-p18,-p39)



dlp <- bind_rows(dl,dgi,dcr,ds,dcp) %>%
  mutate(antigenf=factor(antigenf,
                         levels=c("Giardia VSP-3 or VSP-5",
                                  "Cryptosporidium Cp17 or Cp23",
                                  "E. histolytica LecA",
                                  "Salmonella LPS groups B or D",
                                  "ETEC LT β subunit",
                                  "Cholera toxin β subunit",
                                  "Campylobacter p18 or p39")) ) %>%
  filter(!is.na(antigenf)) %>%
  select(pathogen,antigen,antigenf,
         childid,time,tr,sex,age,agediff,
         mfi,logmfi,logmfidiff,
         roccut,mixcut,unexp,unexpcut,serocut,serocut_desc,
         posroc,posmix,posunex,seropos) %>%
  arrange(pathogen,antigenf,childid,time)

#-----------------------------
# create a dummy variable equal
# to 1 for all obs as a trick to
# get marginal predictions by
# age while still including REs
# for child
#-----------------------------
dlp <- dlp %>%
  mutate(dummy=1)
```

## Estimate curves in each treatment arm
```{r age-curve by arm }
#----------------------------------
# simulataneous CIs for GAMs
# estimated by resampling the 
# Baysian posterior estimates of
# the variance-covariance matrix
# assuming that it is multivariate normal
# the function below also estimates 
# the unconditional variance-covariance
# matrix, Vb=vcov(x,unconditional=TRUE), 
# which allows for undertainty in the actual
# estimated mean as well 
# (Marra & Wood 2012 Scandinavian Journal of Statistics, 
#  Vol. 39: 53–74, 2012, doi: 10.1111/j.1467-9469.2011.00760.x )
# simultaneous CIs provide much better coverage than pointwise CIs
# see: http://www.fromthebottomoftheheap.net/2016/12/15/simultaneous-interval-revisited/
#----------------------------------

gamCI <- function(m,newdata,nreps=10000) {
  require(mgcv)
  require(dplyr)
  Vb <- vcov(m,unconditional = TRUE)
  pred <- predict(m, newdata, se.fit = TRUE)
  fit <- pred$fit
  se.fit <- pred$se.fit
  BUdiff <- MASS::mvrnorm(n=nreps, mu = rep(0, nrow(Vb)), Sigma = Vb)
  Cg <- predict(m, newdata, type = "lpmatrix")
  simDev <- Cg %*% t(BUdiff)
  absDev <- abs(sweep(simDev, 1, se.fit, FUN = "/"))
  masd <- apply(absDev, 2L, max)
  crit <- quantile(masd, prob = 0.95, type = 8)
  pred <- data.frame(newdata,fit=pred$fit,se.fit=pred$se.fit)
  pred <- mutate(pred,
                 uprP = fit + (2 * se.fit),
                 lwrP = fit - (2 * se.fit),
                 uprS = fit + (crit * se.fit),
                 lwrS = fit - (crit * se.fit)
  )
  return(pred)
}


#-----------------------------
# fit cubic spline over age, separately for each 
# treatment arm and antigen
# estimate approximate simultaneous confidence intervals
# for the fits
# set the dummy to 0 to 
# zero out all of the random effects
# see posts on Stack Exchange for explanation:
# https://stats.stackexchange.com/questions/131106/predicting-with-random-effects-in-mgcv-gam/131116#131116
# https://stats.stackexchange.com/questions/189384/predicting-mean-smooth-in-gam-with-smooth-by-random-factor-interaction
#-----------------------------

library(mgcv)
pgamfits <- foreach(ab=levels(dlp$antigenf),.combine=rbind) %dopar% {
    pd <- filter(dlp, antigenf==ab)
    gfit <- mgcv::gam(seropos~s(age, bs="cr")+s(childid,bs="re",by=dummy)+tr, data=pd,family="binomial")
    newd0 <- pd %>% mutate(dummy=0,tr="Cont")
    newd1 <- pd %>% mutate(dummy=0,tr="Int")
    gsci0 <- gamCI(m=gfit,newdata=newd0,nreps=1000)
    gsci1 <- gamCI(m=gfit,newdata=newd1,nreps=1000)
    gsci <- bind_rows(gsci0,gsci1) %>% mutate(antigenf=ab)
    return(gsci)
}

# use antilogit to get back to prevalence from the logit link
pgamfits <- pgamfits %>%
  mutate(fit  = 1/(1+exp(-fit)),
         uprP = 1/(1+exp(-uprP)),
         lwrP = 1/(1+exp(-lwrP)),
         uprS = 1/(1+exp(-uprS)),
         lwrS = 1/(1+exp(-lwrS))) %>%
  mutate(antigenf=factor(antigenf,levels=levels(dlp$antigenf)),
         tr=factor(tr,levels=c("Cont","Int"),labels=c("Control","Intervention"))
  )

```

## Estimate seroprevalence by treatment arm
Estimate overall mean seroprevalence by treatment arm and compare arms. Standard errors account for repeated measurements within child using the influence curve.

```{r tmle means,message=FALSE}
#-----------------------------
# estimate means by treatment
# arm and difference between
# arms using targeted maximum
# likelihood estimation
# get 95% CIs through the
# influence curve clustered
# on child ID
# adjust for age with GLM
#-----------------------------
tmlemeans <- foreach(tri=levels(dlp$tr),.combine=rbind) %:% 
  foreach(ab=levels(dlp$antigenf),.combine=rbind) %dopar% {
    print(ab)
    pd <- filter(dlp,tr==tri & antigenf==ab)
    mu <- tmleAb(Y=pd$seropos,id=pd$childid,W=pd$age,
                 SL.library=c("SL.glm"),
                 family="binomial")
    di <- data.frame(tr=tri,antigenf=ab,mean=mu$psi,se=mu$se,lb=mu$lb,ub=mu$ub)
    return(di)
  }

tmlediffs <- foreach(ab=levels(dlp$antigenf),.combine=rbind) %dopar% {
    pd <- filter(dlp,antigenf==ab) %>% mutate(tr01=ifelse(tr=="Int",1,0))
    diffmu <- tmleAb(Y=pd$seropos,X=pd$tr01,W=pd$age,
                     id=pd$childid,SL.library=c("SL.glm"),
                     family="binomial")
    di <- data.frame(antigenf=ab,diff=diffmu$psi,lb=diffmu$lb,ub=diffmu$ub,p=diffmu$p)
    return(di)
  }

#-----------------------------
# report estimates in tables
#-----------------------------

# seroprevalence among intervention and control
tmlemeans %>%
  arrange(antigenf,tr) %>%
  knitr::kable(digits=2,caption="Seroprevalence by intervention group (Intervention, control)") %>%
  kable_styling(bootstrap_options = "striped",full_width = TRUE)
  

# difference between arms
knitr::kable(tmlediffs,digits=2,caption="Difference in seroprevalence (Intervention minus Control)")  %>%
  kable_styling(bootstrap_options = "striped",full_width = TRUE)

```

No evidence for lower enteric antibody seroprevalence due to intervention.

## Figure of seroprevlance curves

```{r age-seroprevalence figure, fig.width=9 , fig.height=10}
# bright color-blind palette defined in an earlier code chunk
pcols <- c(cblack,cteal)

ppcurve <- ggplot(data=pgamfits,aes(x=age,color=tr,fill=tr)) +
  # approximate simultaneous CI from spline fit
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA) +
  # spline fit
  geom_line(aes(y=fit)) +
  # plot aesthetics
  facet_wrap(~antigenf,nrow=7,ncol=2) +
  scale_x_continuous(limits=c(4,17),breaks=seq(4,16,by=2))+
  scale_y_continuous(breaks=seq(0,1,by=0.2),labels=sprintf("%1.0f",seq(0,1,by=0.2)*100))+
  coord_cartesian(ylim=c(0,1))+
  labs(x="Age in months",y="Seroprevalence (%)") +
  scale_fill_manual(values=pcols) +
  scale_color_manual(values=pcols) +
  theme_minimal() +
  theme(legend.position = "right")

ppcurve

```

# Kongwa, Tanzania Intervention Effects

In Kongwa, Tanzania, antibody measurements were among children ages 1-9 years old. Without children <1, it was difficult to identify seropositivity cutoff values based on mixture models for any of the enteric pathogens (see main text). Therefore, there were only 3 pathogens (Giardia, Crypto, E. histolytica) where we could classify children as seropositive or seronegative becuase out-of-sample specimens were used to estimate cutoffs using an ROC curve. 
In light of this limitation, below we compare the intervention and control groups using mean log MFI values, which is a reasonable alternative in this situation ([Arnold et al. 2017](http://journals.plos.org/plosntds/article?id=10.1371/journal.pntd.0005616).

_Salmonella_ and _Campylobacter_ antibodies were not measured after the first year enrollment survey in 2012.  No clusters were treated at enrollment, therefore we cannot assess the effect of azithromycin treatment on these pathogens. Therefore, limit comparisons to years 2013-2015 to avoid any potential time trends, including those due to a change in bead lots from 2012 until the later years.

```{r load kongwa data}
#-----------------------------
# load the formatted data
# for enteric pathogens
# with seropositivity indicators
#-----------------------------
dl <- readRDS(here("data","kongwa_analysis2.rds"))

dl <- dl %>%
  filter(!antigen %in% c("salb","sald","p18","p39","pgp3")) %>%
  filter(!year %in% c(2012)) %>%
  mutate(antigenf=droplevels(antigenf),
         pathogen=droplevels(pathogen),
         tr=factor(ifelse(aztr==0,"Control","Intervention"),levels=c("Control","Intervention"))
  )
```


```{r kongwa composite antigens, include=FALSE,eval=FALSE}
#-----------------------------
# create composite 
# seropositivity indicators
# that use information from
# multiple antigens
#-----------------------------
dl <- dl %>%
  mutate(seropos=posroc)

dgi <- dl %>% 
  filter(antigen %in% c("vsp3","vsp5")) %>%
  select(svy,year,vilid,clusterid,id,aztr,age,pathogen,antigen,seropos) %>%
  spread(antigen,seropos) %>%
  mutate(seropos=ifelse(vsp3==1|vsp5==1,1,0),
         antigen="vsp3vsp5",
         antigenf="Giardia VSP-3 or VSP-5") %>%
  select(-vsp3,-vsp5)

dcr <- dl %>% 
  filter(antigen %in% c("cp17","cp23")) %>%
  select(svy,year,vilid,clusterid,id,aztr,age,pathogen,antigen,seropos) %>%
  spread(antigen,seropos) %>%
  mutate(seropos=ifelse(cp17==1|cp23==1,1,0),
         antigen="cp17cp23",
         antigenf="Cryptosporidium Cp17 or Cp23") %>%
  select(-cp17,-cp23)


dlp <- bind_rows(dl,dgi,dcr) %>%
  mutate(antigenf=factor(antigenf,
                         levels=c("Giardia VSP-3 or VSP-5",
                                  "Cryptosporidium Cp17 or Cp23",
                                  "E. histolytica LecA")) ) %>%
  filter(!is.na(antigenf)) %>%
  select(pathogen,antigen,antigenf,
         svy,year,vilid,clusterid,id,aztr,age,
         mfi,logmfi,
         roccut,mixcut,
         posroc,posmix,seropos) %>%
  arrange(pathogen,antigenf,svy,year,vilid,clusterid)


```


## Estimate curves in each treatment arm
```{r kongwa age-curve by arm }
#-----------------------------
# fit cubic spline over age
# estimate approximate simultaneous confidence intervals
# for the fits
# include random effects for
# cluster to account for repeated
# observations
#-----------------------------
gamfits <- foreach(ab=levels(dl$antigenf),.combine=rbind) %dopar% {
    pd <- filter(dl, antigenf==ab) %>%
      mutate(dummy=1)
    gfit <- mgcv::gam(logmfi~s(age, bs="cr",k=9)+s(clusterid,bs="re",by=dummy)+tr ,data=pd,family="gaussian")
    newd0 <- pd %>% mutate(dummy=0,tr="Control")
    newd1 <- pd %>% mutate(dummy=0,tr="Intervention")
    gsci0 <- gamCI(m=gfit,newdata=newd0,nreps=1000)
    gsci1 <- gamCI(m=gfit,newdata=newd1,nreps=1000)
    gsci <- bind_rows(gsci0,gsci1) %>% mutate(antigenf=ab)
    return(gsci)
  }

gamfits <- gamfits %>%
  mutate(antigenf=factor(antigenf,levels=levels(dl$antigenf)),
         tr=factor(tr,levels=levels(dl$tr))
  )

```

## Estimate means by treatment arm
Estimate overall mean log MFI by treatment arm and compare arms. Standard errors account for repeated measurements within clusters using the influence curve.

```{r kongwa tmle means,message=FALSE}
#-----------------------------
# estimate means by treatment
# arm and difference between
# arms using targeted maximum
# likelihood estimation
# get 95% CIs through the
# influence curve clustered
# on child ID
# adjust for age with glm
#-----------------------------
ktmlemeans <- foreach(tri=levels(dl$tr),.combine=rbind) %:% 
  foreach(ab=levels(dl$antigenf),.combine=rbind) %dopar% {
    pd <- filter(dl,tr==tri & antigenf==ab)
    mu <- tmleAb(Y=pd$logmfi,id=pd$clusterid,W=select(pd,age),
                 SL.library=c("SL.glm"))
    di <- data.frame(tr=tri,antigenf=ab,mean=mu$psi,se=mu$se,lb=mu$lb,ub=mu$ub)
    return(di)
  }

ktmlediffs <- foreach(ab=levels(dl$antigenf),.combine=rbind) %dopar% {
    pd <- filter(dl,antigenf==ab) %>% mutate(tr01=ifelse(tr=="Control",0,1))
    diffmu <- tmleAb(Y=pd$logmfi,X=pd$tr01,W=select(pd,age),
                     id=pd$clusterid,SL.library=c("SL.glm"))
    di <- data.frame(antigenf=ab,diff=diffmu$psi,se=diffmu$se,lb=diffmu$lb,ub=diffmu$ub,p=diffmu$p)
    return(di)
  }

#-----------------------------
# report estimates in tables
#-----------------------------

# seroprevalence among intervention and control
knitr::kable(arrange(ktmlemeans,antigenf,tr),digits=2,caption="Mean MFI in the Intervention and Control groups")  %>%
  kable_styling(bootstrap_options = "striped",full_width = TRUE)

# difference between groups
knitr::kable(ktmlediffs,digits=2,caption="Difference in MFI (Intervention minus Control)")  %>%
  kable_styling(bootstrap_options = "striped",full_width = TRUE)

```
These results suggest that IgG values for _E. histolytica_ LecA were slightly _higher_ in intervention clusters than control clusters (difference = 0.25, 95% CI = 0.09, 0.42).  This is consistent with age dependent mean curves (below) that show the intervention clusters that received annual azithromycin distribution consistently had slightly higher mean LecA levels. 

Below, estimate the difference between arms in LecA seroprevalence, which suggests a higher seroprevalence in the intervention clusters that received annual azythromycin distribution (21% relative increase).

```{r kongwa leca seroprev, message=FALSE}
leca_serop <- foreach(tri=levels(dl$tr),.combine=rbind) %dopar% {
    pd <- filter(dl,tr==tri & antigenf=="E. histolytica LecA")
    mu <- tmleAb(Y=pd$posroc,id=pd$clusterid,W=select(pd,age),
                 SL.library=c("SL.glm"),family="binomial")
    di <- data.frame(tr=tri,antigenf="E. histolytica LecA", mean=mu$psi,se=mu$se,lb=mu$lb,ub=mu$ub)
    return(di)
}


pd <- dl %>%
  filter(antigenf=="E. histolytica LecA") %>%
  mutate(tr01=ifelse(tr=="Control",0,1))
leca_diff <- tmleAb(Y=pd$posroc,X=pd$tr01,W=select(pd,age),
                     id=pd$clusterid,SL.library=c("SL.glm"),
                    family="binomial")


```


## Figure of mean curves

```{r kongwa age-mean figure, fig.width=9, fig.height=10}
# bright color-blind palette defined in an earlier code chunk
pcols <- c(cblack,cteal)

mucurve <- ggplot(data=gamfits,aes(x=age,color=tr,fill=tr)) +
  # approximate simultaneous CI from spline fit
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA) +
  # spline fit
  geom_line(aes(y=fit)) +
  # plot aesthetics
  facet_wrap(~antigenf,nrow=7,ncol=2) +
  scale_x_continuous(limits=c(1,9),breaks=1:9)+
  scale_y_continuous(breaks=seq(0,4,by=1))+
  coord_cartesian(ylim=c(0,4.5))+
  labs(x="Age in years",y="log10 Luminex Response (MFI-bg)") +
  scale_fill_manual(values=pcols) +
  scale_color_manual(values=pcols) +
  theme_minimal() +
  theme(legend.position = "right")

mucurve

```


# Kongwa, Tanzania Bead Lot Effects

## Load and format data
```{r data format bead comp}
#-----------------------------
# load the formatted data
# created with 
# asembo-enteric-ab-data-format.Rmd
#-----------------------------
dl <- readRDS(here("data","kongwa_analysis.rds"))

# list the enteric antigens and formatted labels for them
mbavars <- c("vsp3","vsp5","cp17","cp23","leca","salb","sald","etec","cholera","p18","p39","pgp3")

mbalabs <- c("Giardia VSP-3","Giardia VSP-5","Cryptosporidium Cp17","Cryptosporidium Cp23","E. histolytica LecA","Salmonella LPS group B","Salmonella LPS group D","ETEC LT β subunit","Cholera toxin β subunit","Campylobacter p18","Campylobacter p39","C. trachomatis pgp3")

dl <- dl %>%
  mutate(antigen=as.character(antigen))

#-----------------------------
# subset to antigens measured
# in all years
#-----------------------------
dl <- dl %>%
  filter(antigen %in% c("vsp3","vsp5","cp17","cp23","leca","etec"))

#-----------------------------
# create a stratification
# variable for bead set
#-----------------------------
dl <- dl %>%
  mutate(beadset = factor(ifelse(year==2012,"Year 1","Years 2-4")),
         antigenf=factor(antigenf))


```

```{r roc cutoffs, echo=TRUE, message=FALSE, warning=FALSE}
#----------------------------
# Add ROC cutoffs 
# Estimated by Diana Martin
# for years 1 and 2-4 for antigens
# included in multiple years
#----------------------------
dl$roccut <- NA
dl$roccut[dl$antigen %in% "vsp3"] <- log10(116) 
dl$roccut[dl$antigen %in% "vsp5"] <- log10(109) 
dl$roccut[dl$antigen %in% "cp23"] <- log10(1185)
dl$roccut[dl$antigen %in% "cp17"] <- log10(321)
dl$roccut[dl$antigen %in% "leca"] <- log10(153)
dl$roccut[dl$antigen %in% "pgp3"] <- log10(818)


# different cutoffs for bead set used in years 2-4
dl$roccut[dl$antigen %in% "vsp3"  & dl$svy >1] <- log10(170)
dl$roccut[dl$antigen %in% "vsp5"  & dl$svy >1] <- log10(141)
dl$roccut[dl$antigen %in% "cp23"  & dl$svy >1] <- log10(378) 
dl$roccut[dl$antigen %in% "cp17"  & dl$svy >1] <- log10(180) 
dl$roccut[dl$antigen %in% "leca"  & dl$svy >1] <- log10(94)
dl$roccut[dl$antigen %in% "pgp3"  & dl$svy >1] <- log10(882)

```

## Age stratified antibody distributions

This is a beadset-stratified version of Figure 1 - supplement 1 in the article.

```{r age stratified distribution figure, fig.width=12,fig.height=6}
# create age categories
dlplot <- dl %>%
  mutate(agecat = cut(age,breaks=c(0,1,2,3,4,10),
                      labels=c("1 year","2 years","3 years","4 years","5-9 years")),
         antigenf=factor(antigenf,levels=mbalabs)
         ) %>%
  filter(antigen!="pgp3") %>% mutate(antigenf=droplevels(antigenf)) %>%
  group_by(antigen) %>%
  mutate(plotroc=min(roccut))


# custom color blind color palette is in the preamble chunck
pcols <- c(cteal,corange)



# custom log labels
log10labs <- c( 
  expression(10^0),
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)

# faceted over age and antigen
p2 <- ggplot(data=dlplot,aes(x=logmfi,group=beadset,fill=beadset,color=beadset)) +
  facet_grid(agecat~antigenf,scales='free')+
  geom_density(aes(y=..density..),color=NA,alpha=0.5)+
  geom_vline(aes(xintercept = roccut,color=beadset)) +
  coord_cartesian(ylim=c(0, 1.25))+
  scale_x_continuous(expand=c(0,0),limits=c(0,5),breaks=0:4,labels=log10labs) +
  scale_fill_manual(values=pcols)+
  scale_color_manual(values=pcols)+
  labs(title="",x="Luminex Response (MFI-bg)",y="Density") +
  theme_minimal()+
  theme(
    strip.text.x = element_text(size=9),
    strip.text.y=element_text(size=12,angle=0),
    legend.position="top"
  )

p2
```

The patterns are very similar across beadsets. Vertical lines are ROC-cutoffs for each beadset.

Below is a figure that contrasts the distribution of the Year 1 beadset (blue) with the overall distribution (black line); and then Years 2-4 beadset (orange) with the overall distribution (black line).  The plots  show that the combined analysis, which pools data over all years, essentially reflects the patterns observed in years 2-4 -- consistent with that bead set contributing >4 times more observations to the analysis (n=4,087 vs n=902).

```{r distribution comparison to overall, fig.width=12,fig.height=6}
# faceted over age and antigen
p3 <- ggplot(data=dlplot,aes(x=logmfi)) +
  facet_grid(agecat~antigenf,scales='free')+
  geom_density(data=filter(dlplot,beadset=="Year 1"), aes(y=..density..),fill=cteal,color=NA,alpha=0.5)+
  geom_density(aes(y=..density..),fill=NA,color="black")+
  # scale_y_continuous(expand=c(0,0),breaks=seq(0,40,by=10)) +
  coord_cartesian(ylim=c(0, 1.25))+
  scale_x_continuous(expand=c(0,0),limits=c(0,5),breaks=0:4,labels=log10labs) +

  labs(title="",x="Luminex Response (MFI-bg)",y="Density") +
  theme_minimal()+
  theme(
    strip.text.x = element_text(size=9),
    strip.text.y=element_text(size=12,angle=0),
    legend.position="top"
  )

p3

p4 <- ggplot(data=dlplot,aes(x=logmfi)) +
  facet_grid(agecat~antigenf,scales='free')+
  geom_density(data=filter(dlplot,beadset=="Years 2-4"), aes(y=..density..),fill=corange,color=NA,alpha=0.5)+
  geom_density(aes(y=..density..),fill=NA,color="black")+
  coord_cartesian(ylim=c(0, 1.25))+
  scale_x_continuous(expand=c(0,0),limits=c(0,5),breaks=0:4,labels=log10labs) +

  labs(title="",x="Luminex Response (MFI-bg)",y="Density") +
  theme_minimal()+
  theme(
    strip.text.x = element_text(size=9),
    strip.text.y=element_text(size=12,angle=0),
    legend.position="top"
  )

p4

```

## Joint distribution of antibody response

Summarize the joint distribution of antibody response, stratified by beadsets used in Y1 and Y2-4 for antigens included in both beadsets.

```{r group antigens and reshape}
#--------------------------------
# create antigen groupings for
# comparisons
# drop obs not contributing
#--------------------------------
dl2 <- dl %>%
  mutate(comp= ifelse(antigen %in% c("vsp3","vsp5"),"Giardia",NA),
         comp= ifelse(antigen %in% c("cp17","cp23"),"Cryptosporidium",comp),
         comp= factor(comp,levels=c("Giardia","Cryptosporidium"))
         ) %>% 
  mutate(xlab=ifelse(antigen %in% c("vsp3","vsp5"),"VSP-3",NA),
         ylab=ifelse(antigen %in% c("vsp3","vsp5"),"VSP-5",NA),
         xlab=ifelse(antigen %in% c("cp17","cp23"),"Cp17",xlab),
         ylab=ifelse(antigen %in% c("cp17","cp23"),"Cp23",ylab)
         ) %>%
  filter(!is.na(comp))

#--------------------------------
# for antigen each pair, 
# label it as "x" or "y" to 
# spread it to wide format
#--------------------------------
dw <- dl2 %>%
  mutate(xy=ifelse(antigen %in% c("vsp3","cp17"),"x","y")) %>%
  select(beadset,comp,xlab,ylab,id,xy,logmfi) %>%
  spread(xy,logmfi) %>%
  mutate(beadset=factor(beadset,levels=c("Year 1","Years 2-4")))

```

```{r est correlations}
#--------------------------------
# estimate spearman's correlation
# within each country, comparison
#--------------------------------
dcorr <- dw %>%
  group_by(beadset,comp) %>%
  mutate(corxy=cor(x,y,method="spearman",use="pairwise.complete.obs") ) %>%
  summarize(corxy=max(corxy,na.rm=T))
```

```{r est trimmed smooths}
#--------------------------------
# estimate smooths, trimmed to
# drop the bottom and top 1% of
# data in each comparison
# to avoid edge effects
#--------------------------------
dsmooths <- foreach(beadseti=levels(dw$beadset),.combine=rbind) %:%
                      foreach(compi=levels(dw$comp),.combine=rbind) %do% {
                        pd <- filter(dw,beadset==beadseti & comp==compi)
                        if(nrow(pd)>0) {
                          xqs <- quantile(pd$x,probs=c(0.01,0.99),na.rm=TRUE)
                          newd <- data.frame(x=seq(xqs[1],xqs[2],by=0.01))
                          lfit <- loess(y~x,data=pd)
                          return(data.frame(beadset=beadseti,comp=compi,x=newd,y=predict(lfit,newdata=newd)))
                        }
                        
  
}

```

```{r composite figure,fig.width=7,fig.height=9 }

pcols <- c(cteal,corange)

# grab labels
dlabs <- dw %>% select(beadset,comp,xlab,ylab) %>% group_by(beadset,comp) %>% slice(1)

complot <- ggplot(data=dw,aes(x=x,y=y,color=beadset)) +
  facet_grid(comp~beadset) +
  geom_point(pch=19,alpha=0.1) +
  geom_line(data=dsmooths,aes(x=x,y=y),col="black",size=1.2)+
  geom_text(data=dcorr,
            aes(x=0.5,y=4.3,label=paste("rho ==",sprintf("%1.2f",corxy)) ),
            parse=TRUE, col="black")   +
  geom_text(data=dlabs,aes(x=2.3,y=0.1,label=xlab),color="gray40",angle=0)+
  geom_text(data=dlabs,aes(x=0.1,y=2.3,label=ylab),color="gray40",angle=90)+
  scale_x_continuous(limits=c(0,4.6),breaks=0:4,labels = log10labs)+
  scale_y_continuous(limits=c(0,4.6),breaks=0:4,labels = log10labs)+
  scale_color_manual(values=pcols) +
  coord_equal() +
  labs(x="Luminex Response (MFI-bg)",y="Luminex Response (MFI-bg)") +
  theme_minimal(base_size=12) +
  theme(
    strip.text.x=element_text(size=12),
    strip.text.y=element_text(size=12,angle=0),
    legend.position="none"
  )


complot

```


Stratified version of Figure 2 in the main text.
The relationships and correlation estimates are nearly identical across beadsets.


## Age-dependent mean curves
Estimate age-dependent mean log10 MFI-bg in the Tanzania study, separately for the two different bead sets.
```{r age-curve mean gam}

# ---------------------------------------
# fit cubic spline over age, separately for each antigen
# and beadset
# estimate approximate simultaneous confidence intervals
# for the fits
# ---------------------------------------
library(mgcv)
gamfits <- foreach(ab=levels(dl$antigenf),.combine=rbind) %:%
  foreach(beadseti=levels(dl$beadset),.combine=rbind) %dopar% {
    pd <- dplyr::filter(dl, antigenf==ab & beadset==beadseti)
    gfit <- gam(logmfi~s(age, bs="cr",k=9),data=pd)
    gsci <- gamCI(m=gfit,newdata=pd,nreps=1000)
    gsci$antigen <- ab
    return(gsci)
  }

# ---------------------------------------
# get overall, pooled estimate
# ---------------------------------------
gamfits2 <- foreach(ab=levels(dl$antigenf),.combine=rbind) %dopar% {
    pd <- dplyr::filter(dl, antigenf==ab)
    gfit <- gam(logmfi~s(age, bs="cr",k=9),data=pd)
    gsci <- gamCI(m=gfit,newdata=pd,nreps=1000)
    gsci$antigen <- ab
    gsci$beadset <- factor("Combined",levels=c("Year 1","Years 2-4","Combined"))
    return(gsci)
}

gamfits <- gamfits %>%
  mutate(beadset=factor(beadset,levels=c("Year 1","Years 2-4","Combined"))) %>%
  bind_rows(gamfits2)

```



```{r age-curve mean figure tanzania, fig.height=12, fig.width=6}

log10labs <- c( 
  expression(10^0),
  expression(10^1),
  expression(10^2),
  expression(10^3),
  expression(10^4)
)


dplot <- data.frame(gamfits) %>%
  dplyr::filter(pathogen!="V. cholerae") %>%
  mutate(pathogen=factor(pathogen,levels=c("Giardia",
                                           "Cryptosporidium",
                                           "E. histolytica",
                                           "ETEC")),
         antigenf=factor(antigenf,levels=c("Giardia VSP-3","Giardia VSP-5",
                                           "Cryptosporidium Cp17","Cryptosporidium Cp23",
                                           "E. histolytica LecA",
                                           "ETEC LT β subunit")))

pcols <- c(cteal,corange,cblack)

pcurve <- ggplot(data=dplot,aes(x=age+0.5,group=beadset,color=beadset,fill=beadset)) +
  # approximate simultaneous CI from spline fit
  geom_ribbon(aes(ymin=lwrS,ymax=uprS),alpha=0.2,color=NA) +
  # spline fit
  geom_line(aes(y=fit)) +
  # plot aesthetics
  facet_wrap(~antigenf,nrow=6,ncol=2) +
  scale_x_continuous(limits=c(0,10),breaks=seq(0,10,by=2))+
  scale_y_continuous(breaks=0:4,labels=log10labs)+
  coord_cartesian(ylim=c(0,4.5))+
  labs(x="Age in years",y="Luminex Response (MFI-bg)") +
  scale_fill_manual(values=pcols) +
  scale_color_manual(values=pcols) +
  theme_minimal(base_size=16) +
  theme(legend.position = "top")

pcurve


```

The results show that in general the MFI-bg levels were slightly lower in year 1 than in years 2-4, but that the qualitative age-dependent patterns are similar. Since year is perfectly confounded by bead lot, there is no way to distinguish whether differences are due to between-year differences in exposure or due to bead lot effects.  The plots also show that the combined analysis, which pools data over all years, essentially reflects the patterns observed in years 2-4 -- consistent with that bead set contributing >4 times more observations to the analysis (n=4,087 vs n=902).

# Session Info
```{r session info}
sessionInfo()
```

