---
title: Enteropathogen seroepidemiology among children in low-resource settings
subtitle: Table 1, summary of antibody measurements and cutoffs by cohort
output: 
  html_document:
    highlight: haddock
    theme: default
    code_folding: show
    df_print: paged
    toc: yes
    toc_depth: '3'
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Notebook summary
This notebook summarizes the antigens included in each cohort, the number of children, number of measurements, and seropositivity cutoff values estimated through ROC-based cutoffs, mixture models, or distribution among presumed unexposed (see main article Methods for details).

This notebook draws on datasets created by the notebooks that estimate antibody distributions in each cohort:
  * Haiti: Fig1-haiti-ab-distributions.Rmd -> haiti_analysis2.rds
  * Kenya: Fig1sup3-asembo-ab-distributions.Rmd -> asembo_analysis2.rds
  * Tanzania: Fig1sup1sup2-kongwa-ab-distributions.Rmd -> kongwa_analysis2.rds


# Script preamble
```{r preamble}
#-----------------------------
# preamble
#-----------------------------
library(here)
here()
library(tidyverse)
library(kableExtra)


```

# Load the data

```{r load data}
#-----------------------------
# load the formatted data
#-----------------------------
dh <- readRDS(here("data","haiti_analysis2.rds"))
da <- readRDS(here("data","asembo_analysis2.rds"))
dk <- readRDS(here("data","kongwa_analysis2.rds"))

# re-order antigens in Kenya for consistency with other cohorts
da <- da %>%
  mutate(antigenf=factor(antigenf,levels=c("Giardia VSP-3","Giardia VSP-5","Cryptosporidium Cp17","Cryptosporidium Cp23","E. histolytica LecA","Salmonella LPS group B","Salmonella LPS group D","ETEC LT β subunit","Cholera toxin β subunit","Campylobacter p18","Campylobacter p39")))
```

# Summary counts and statistics for each study

## Haiti
```{r haiti summary, warning=FALSE}
#-----------------------------
# Haiti summary
#-----------------------------
dhsum <- dh %>%
  group_by(antigenf) %>%
  summarize(Nchild=length(unique(id)),
            Nobs=length(!is.na(mfi)),
            roccut=max(roccut,na.rm=T),
            mixcut=max(mixcut,na.rm=T),
            unexpcut=max(unexpcut,na.rm=T)
  )
```

## Kenya
```{r kenya summary, warning=FALSE}
#-----------------------------
# Kenya summary
#-----------------------------
dasum <- da %>%
  group_by(antigenf) %>%
  summarize(Nchild=length(unique(childid)),
            Nobs=length(!is.na(mfi)),
            roccut=max(roccut,na.rm=T),
            mixcut=max(mixcut,na.rm=T),
            unexpcut=max(unexpcut,na.rm=T)
  )
```

## Tanzania
```{r tanzania summary, warning=FALSE}
#-----------------------------
# Tanzania summary
# report ROC cutoffs from
# years 2-4
#-----------------------------
dksum <- dk %>%
  group_by(antigenf) %>%
  mutate(roc2=ifelse(year>2012,roccut,NA)) %>%
  summarize(Nchild=length(unique(id[is.na(mfi)==F])),
            Nobs=sum(is.na(mfi)==F),
            roccut=max(roc2,na.rm=T),
            mixcut=max(mixcut,na.rm=T)  ) 
```

# Print the table
```{r table, warning=FALSE, results="html"}
dtab <- bind_rows(dhsum,dasum,dksum) %>%
  mutate(roccut=ifelse(is.infinite(roccut),NA,roccut),
         mixcut=ifelse(is.infinite(mixcut),NA,mixcut),
         unexpcut=ifelse(is.infinite(unexpcut),NA,unexpcut)
         )

# exclude trachoma, since article is focused on enteropathogens
dtab <- dtab %>%
  filter(antigenf!="C. trachomatis pgp3")

# add custom footnotes (careful, this is fragile!)
dtab$antigenf[c(27,28,31,32)] <- paste0(dtab$antigenf[c(27,28,31,32)],footnote_marker_symbol(2))
dtab$antigenf[c(30)] <- paste0(dtab$antigenf[c(30)],footnote_marker_symbol(3))

options(knitr.kable.NA = "")
knitr::kable(dtab,digits=2,escape=FALSE,
             caption="Table 1. Number of children and samples tested, and estimated seropositivity cutoffs by country and antigen included in the seroepidemiologic analyses.",
             col.names=c("","N<br>children","N<br>samples","External<br>Reference","Mixture<br>Model","Presumed<br>Unexposed")) %>%
  kable_styling(bootstrap_options = c("striped","condensed"),full_width = TRUE) %>%
  add_header_above(c(" " = 3, "Seropositivity cutoff, log10 IgG (MFI-bg) *" = 3)) %>%
  group_rows("Leogane, Haiti", 1,10) %>%
  group_rows("Asembo, Kenya", 11,21) %>%
  group_rows("Kongwa, Tanzania", 22,32) %>%
  footnote(
    symbol=c("Seropositivity cutoffs determined using external reference samples (typically ROC curves except for Giardia and E. hystolitica in Haiti), finite Gaussian mixture models, or distribution among the presumed unexposed (see Methods for details). External reference cutoffs reported from years (2013-2015) in Tanzania, estimated among 4,087 samples. Cutoff values are missing if they could not be estimated in each method; cutoff values based on the presumed unexposed required longitudinal measurements within individual children and therefore could not be estimated for any antigen in the repeated cross-sectional design in Tanzania.",
             "Measured only in year 1  of the study (2012)",
             "Measured only in years 2-4 of the study (2013-2015)")
  )


```




# Session Info
```{r session info}
sessionInfo()
```

