---
title: Enteropathogen seroepidemiology among children in low-resource settings
subtitle: Haiti incidence analysis, Table 2
output: 
  html_document:
    highlight: haddock
    theme: default
    code_folding: show
    df_print: paged
    toc: yes
    toc_depth: '3'
    toc_float:
      collapsed: yes
      smooth_scroll: yes
---

# Notebook Summary

This script estimates the incidence rate of seroconversion and seroreversion in the Haitian cohort. It also estimates the incidence rate of antibody boosting and waning events, defined as a 4-fold change in antibody levels to above a seropositivity cutoff (boosting) or from above a seropositivity cutoff (waning).


# Script preamble
```{r preamble}
#-----------------------------
# preamble
#-----------------------------
library(here)
here()

# load packages
library(tidyverse)
library(kableExtra)

# set up for parallel computing
# configure for a laptop (use only 3 cores)
library(foreach)
library(doParallel)
registerDoParallel(cores=3)

# bright color blind palette:  https://personal.sron.nl/~pault/ 
cblack <- "#000004FF"
cblue <- "#3366AA"
cteal <- "#11AA99"
cgreen <- "#66AA55"
cchartr <- "#CCCC55"
cmagent <- "#992288"
cred <- "#EE3333"
corange <- "#EEA722"
cyellow <- "#FFEE33"
cgrey <- "#777777"

# safe color blind palette
# http://jfly.iam.u-tokyo.ac.jp/color/
# http://www.cookbook-r.com/Graphs/Colors_(ggplot2)/
cbPalette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00", "#CC79A7")

```


```{r load data}
#-----------------------------
# load the formatted data
# created with 
# haiti-enteric-ab-data-format.Rmd
#-----------------------------
dl <- readRDS(here("data","haiti_analysis2.rds"))

```

# Identify incident changes
Among children, identify those that changed status between enrollment and follow-up. Those who changed from negative to positive are seroconverters (`seroi` below), and those who changed from positive to negative are seroreverters (`seror` below).


```{r long incidence}

#-----------------------------
# identify seropositive measures
# hierarchy of information for 
# cutoffs:
# 1. ROC
# 2. mixture model based
# 3. estimated among presumed unexposed
#
# store the cutoff value used
# for figures
#-----------------------------
dl <- dl %>%
  mutate(seropos=ifelse(!is.na(posroc),posroc,posmix),
         serocut=ifelse(!is.na(posroc),roccut,mixcut),
         serocut_desc=ifelse(!is.na(posroc),"ROC","Mixture Model")) %>%
  mutate(seropos=ifelse(!is.na(seropos),seropos,posunex),
         serocut_desc=ifelse(!is.na(serocut),serocut_desc,"Unexp dist"),
         serocut=ifelse(!is.na(serocut),serocut,unexpcut)) 

```

# Seroincidence estimates

Estimates of the mean seroincidence rates. The rates assume that incident sero-conversions and sero-reversions happened at the mid-point between measurements for each child.

As with the age-prevalence curve analysis, use information across multiple recombinant antigens for each pathogen when identifying risk periods and incident seroconversions and seroreversions.



```{r identify 4-fold changes}
#-----------------------------
# Identify incident boosting/waning 
# of 4-fold MFI or more. For
# incident boosting, +4fold increase
# that ends above seropositivity cutoff
# for incident waning, -4fold decrease
# that starts above seropositivity cutoff
#-----------------------------
dl2 <- dl %>%
  group_by(antigenf,id) %>%
  arrange(antigenf,id,age) %>%
  mutate(nobs=n(),obsnum=row_number(),
         agediff=ifelse(obsnum==1,lead(age)-age,age-lag(age)),
         
    logmfi_lag  = lag(logmfi),
    logmfi_lead = lead(logmfi),
    logmfi_dlag  = logmfi - logmfi_lag,
    logmfi_dlead = logmfi_lead - logmfi,
    logmfi_d4fold = ifelse(logmfi_dlag>log10(4),1,0),
    # incident boosting based on a 4-fold increase in MFI
    # with a second measure above the seropositivity cutoff
    seroi4fold = ifelse(logmfi_dlag>log10(4) & logmfi>serocut,1,0),
    seroi4fold = ifelse(is.na(logmfi_dlag),NA,seroi4fold),
    # incident waning based on a 4-fold decrease in MFI
    # with the first measure above the seropositivity cutoff
    seror4fold = ifelse(logmfi_dlag< -log10(4) & logmfi_lag>serocut,1,0)
         ) %>%
  ungroup()

```

```{r composite antigens}
#-----------------------------
# create composite
# seroprevalence, boosting,
# and waning indicators
# that use information from
# multiple antigens if available
#-----------------------------
dl3 <- dl2 %>%
  ungroup() %>%
  mutate(pathogen2=as.character(pathogen),
         pathogen2=ifelse(antigenf=="Norovirus GI.4","Norovirus GI.4",pathogen2),
         pathogen2=ifelse(antigenf=="Norovirus GII.4.NO","Norovirus GII.4.NO",pathogen2))%>%
  group_by(pathogen2,id,obsnum) %>%
  select(pathogen=pathogen2,id,obsnum,age,agediff,seropos,seroi4fold,seror4fold) %>%
  # since incidence could fall at slightly different times across the antigens
  # for the same pathogen, use the earlier onset as the incidence measure
  arrange(pathogen,id,obsnum) %>%
  group_by(pathogen,id) %>%
  mutate(lagseroi=lag(seroi4fold),
         seroi4fold=ifelse(seroi4fold==1 & lagseroi==1 & !is.na(lagseroi),0,seroi4fold)) %>%
  # since incidence could fall at slightly different times across the antigens
  # for the same pathogen, use the later onset as the incident measure
  mutate(leadseror=lead(seror4fold),
         seror4fold=ifelse(seror4fold==1 & leadseror==1 & !is.na(leadseror),0,seror4fold)) %>%
  # now get the max for each pathogen at each timepoint
  group_by(pathogen,id,obsnum) %>%
  mutate(seropos=max(seropos),
         seroi4fold=max(seroi4fold),
         seror4fold=max(seror4fold)
         ) %>%
  slice(1) %>%
  mutate(antigenf = case_when(
    pathogen == "Giardia" ~ "Giardia VSP-3 or VSP-5",
    pathogen == "Cryptosporidium" ~ "Cryptosporidium Cp17 or Cp23",
    pathogen == "E. histolytica" ~ "E. histolytica LecA",
    pathogen == "Salmonella" ~ "Salmonella LPS groups B or D",
    pathogen == "ETEC" ~ "ETEC LT β subunit",
    pathogen == "Norovirus GI.4" ~ "Norovirus GI.4",
    pathogen == "Norovirus GII.4.NO" ~ "Norovirus GII.4.NO"),
    antigenf=factor(antigenf,levels=c("Giardia VSP-3 or VSP-5",
                                  "Cryptosporidium Cp17 or Cp23",
                                  "E. histolytica LecA",
                                  "Salmonella LPS groups B or D",
                                  "ETEC LT β subunit",
                                  "Norovirus GI.4",
                                  "Norovirus GII.4.NO"))
  ) %>%
  select(pathogen,antigenf,id,age,obsnum,seropos,seroi4fold,seror4fold)

#----------------------------
# there were 7 chilren that
# had a +4-fold increase in 
# one crypto antigen and a 
# 4-fold decrease in another
# at the same measurement
# this is unusual. In these
# seven cases, set the 
# incident waning indicator to 0
#----------------------------
table(dl3$seroi4fold,dl3$seror4fold)
indivp <- ggplot(data=filter(dl, 
                             id %in% c(2853,3119,3155,4637,5308,5344,5374) & 
                               antigen %in% c("cp17","cp23")), 
                 aes(x=age,y=logmfi,id=factor(id),color=antigenf)) +
  geom_line() +
  geom_point()+
  facet_wrap(~id,ncol=3) +
  theme_minimal() +
  theme(legend.position="top")
indivp

dl3 <- dl3 %>%
  mutate(seror4fold=ifelse(seroi4fold==1 & seror4fold==1,0,seror4fold))
table(dl3$seroi4fold,dl3$seror4fold)


```


```{r incidence}

#-----------------------------
# identify incident 
# seroconversions and reversions
#-----------------------------

# group the data by child and
# use lags to identify
# time in years between measurements,
# sero-conversions + sero-reversions 
# between measurements
# set the first measurement to 
# missing for the incidence indicators
# (using the is.na(age_diff) to identify them)
dl4 <- dl3 %>%
  group_by(antigenf,id) %>% 
  arrange(antigenf,id,obsnum) %>%
  mutate(age_min  = min(age),
         agediff = age - lag(age),
         
         # incident seroconversions and reversions
         # including cumulative numbers
         # based on crossing seropositivity cutoff
         seropos_lag  = lag(seropos),
         seroi = ifelse(seropos==1 & seropos_lag==0,1,0),
         seroi = ifelse(is.na(agediff),NA,seroi),
         seroin = cumsum(ifelse(is.na(seroi),0,seroi)),
         seroin = ifelse(seroi==1,seroin,0),
         seror = ifelse(seropos==0 & seropos_lag==1,1,0),
         seror = ifelse(is.na(agediff),NA,seror),
         serorn = cumsum(ifelse(is.na(seror),0,seror)),
         serorn = ifelse(seror==1,serorn,0)
         ) %>%
  select(
         id,obsnum,
         starts_with("age"),
         antigenf,
         seropos,seroi,seroin,seror,serorn,
         seroi4fold,seror4fold,
         -seropos_lag)

```

## Seroincidence rates based on seropositivity cutoff

```{r long incidence rate estimates, warning=FALSE}
#-----------------------------
# estimate sero-incidence rates
# for conversion and reversion
# estimate SEs with a bootstrap
#-----------------------------

#-----------------------------
# estimate time at risk
# for seroconversion and reversion
# assumed to be 1/2 of time
# between measurements
# if indivs are seropositive
# at measurement 1 they are
# not at risk for seroconversion
# (and vice-versa for seroreversion)
#-----------------------------
dl5 <- dl4 %>%
  mutate(ptc = ifelse(seropos==0,agediff,0),
         ptc = ifelse(seropos==1 & seroi==1,agediff/2,ptc),
         ptr = ifelse(seropos==1 & seroi==0,agediff,0),
         ptr = ifelse(seropos==0 & seror==1,agediff/2,ptr))

#-----------------------------
# estimate incidence rates
# for each antibody
# units are episodes per child-year
#-----------------------------
rate_ests <- dl5 %>%
  group_by(antigenf) %>%
  summarize(ni=sum(seroi,na.rm=T),
            nit=sum(ptc,na.rm=T),
            nr=sum(seror,na.rm=T),
            nrt=sum(ptr,na.rm=T)) %>%
  mutate(seroi=ni/nit,
         seror=nr/nrt)

#-----------------------------
# get bootstrap CIs
# resampling children with 
# replacement due to repeated
# measures
#-----------------------------
dboot <- dl5 %>%
  group_by(antigenf) %>%
  select(id,antigenf,seroi,seror,ptc,ptr) %>%
  mutate(id=as.character(id))

set.seed(123)
nreps <- 1000
ids <- unique(dboot$id)
bsamp <- matrix(sample(ids,size=length(ids)*nreps,replace=TRUE),
                nrow=length(ids),ncol=nreps)
bootests <- foreach(ab=levels(dboot$antigenf),.combine=rbind) %:%
  foreach(brep=1:nreps,.combine=rbind) %dopar% {
    di <- dboot %>% filter(antigenf==ab)
    di <- left_join(data.frame(id=bsamp[,brep]),di,by=c("id")) %>%
      select(-id) %>% group_by(antigenf) %>%
      summarize_all(function(x) sum(x,na.rm=TRUE))
    }

rate_cis <- bootests %>% 
  group_by(antigenf) %>%
  mutate(sero_c=seroi/ptc,sero_r=seror/ptr) %>%
  # restrict to finite values (some ETEC bootstrap replicates have a denominator of 0)
  summarize(seroi_lb = quantile(sero_c[is.finite(sero_c)],probs=c(0.025),na.rm=T),
            seroi_ub = quantile(sero_c[is.finite(sero_c)],probs=c(0.975),na.rm=T),
            seror_lb = quantile(sero_r[is.finite(sero_r)],probs=c(0.025),na.rm=T),
            seror_ub = quantile(sero_r[is.finite(sero_r)],probs=c(0.975),na.rm=T)
            )

rate_ests <- left_join(rate_ests,rate_cis,by="antigenf") %>%
  select(antigenf,ni,nit,starts_with("seroi"),nr,nrt,starts_with("seror"))


```

Confirm bootstrap distributions of the rates are smooth and reasonable
```{r bood dist figs}
bsum <- bootests %>% 
  group_by(antigenf) %>%
  mutate(sero_c=seroi/ptc,sero_r=seror/ptr)

ggplot(data=bsum,aes(x=sero_r)) + geom_density() + facet_wrap(~antigenf,ncol=2,scales="free") + theme_minimal()

ggplot(data=bsum,aes(x=sero_c)) + geom_density() + facet_wrap(~antigenf,ncol=2,scales="free") + theme_minimal()

```


Tables of results
```{r long incidence rate tables}
rate_ests %>%
  select(antigenf,nit,ni,starts_with("seroi")) %>%
  knitr::kable(digits=2,
             caption="Incidence rates of seroconversion per child year",
             col.names = c("Antigen","Years at risk","n events","Incidence per year","min95","max95"),
             row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped"),full_width = TRUE)


rate_ests %>%
  select(antigenf,nrt,nr,starts_with("seror")) %>%
  knitr::kable(digits=2,
             caption="Incidence rates of seroreversion per child year",
             col.names = c("Antigen","Years at risk","n events","Incidence per year","min95","max95"),
             row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped"),full_width = TRUE)

```


## Seroincidence rates based on 4-fold change

```{r long incidence rate estimates 4-fold, warning=FALSE}
#-----------------------------
# estimate sero-incidence rates
# for antibody boosting and waning 
# estimate SEs with a bootstrap
#-----------------------------

#-----------------------------
# estimate time at risk
# for seroconversion and reversion
# assumed to be 1/2 of time
# between measurements
# if indivs seroconvert
# 
# create indicators of at risk
# based on incidence boosting
# and waning
# at risk if seronegative OR experienced
# a 4-fold decrease in MFI in previous period
#
# the below code is opaque, but confirmed through 
# extensive vetting and inspection that it does this:
# children are at risk of boosting if they are seronegative or
# if they experience a 4-fold decrease in MFI between the previous
# two measurements
#-----------------------------
dl6 <- dl4 %>%
  mutate(seroi4fold=ifelse(obsnum==1,0,seroi4fold),
         seror4fold=ifelse(obsnum==1,0,seror4fold),
         maxobs=max(obsnum),
         # seronegative at obs 1: at risk
         # seropositive at obs 1: not at risk
         cntr = 0,
         cntr = ifelse(seroi4fold==1 & lag(seroi4fold)==0,-1,0), # counter -1 if first +4
         cntr = ifelse(obsnum==1 & seropos==0,1,cntr), # counter +1 if seroneg&obs1
         cntr = ifelse(seror4fold==1,1,cntr), # counter +1 if seror
         cntr = ifelse(seror4fold==1 & seropos==0 & lag(cntr)==1,cntr-1,cntr),
         
         # sum counter to help with recodes below
         cntrs = cumsum(cntr),
         
         cntr = ifelse(obsnum!=maxobs & lead(cntr)==-1 & lead(cntrs)==-1,1,cntr),
         cntr = ifelse(obsnum!=maxobs & lead(cntr)==-1 & lead(cntrs)==-2,1,cntr),
         cntr = ifelse(seror4fold==1 & cntrs==2,0,cntr),

         # indicator of whether a chlid is at risk for seroconversion
         atrisk = cumsum(cntr),
         atrisk = lag(atrisk),
         
         # fix a very small number of idiosyncratic situations
         # that don't fit neatly into the above algorithm (n=5)
         atrisk = ifelse(obsnum==1,0,atrisk),
         atrisk = ifelse(obsnum > 2 & atrisk==2 | lag(atrisk==2),atrisk-1,atrisk),
         
         # now calculate person time at risk for conversion and reversion
         atriskc = atrisk,
         atriskr = 1-atrisk,
         ptc = ifelse(atriskc==1,agediff,0),
         ptc = ifelse(atriskc==1 & seroi4fold==1,agediff/2,ptc),
         ptr = ifelse(atriskr==1,agediff,0),
         ptr = ifelse(atriskr==1 & seror4fold==1,agediff/2,ptr)
         
  )


#-----------------------------
# estimate incidence rates
# for each antibody
# units are episodes per child-year
#-----------------------------
rate_ests_4fold <- dl6 %>%
  group_by(antigenf) %>%
  summarize(ni=sum(seroi4fold,na.rm=T),
            nit=sum(ptc,na.rm=T),
            nr=sum(seror4fold,na.rm=T),
            nrt=sum(ptr,na.rm=T)) %>%
  mutate(seroi=ni/nit,seror=nr/nrt)

#-----------------------------
# get bootstrap CIs
# resampling children with 
# replacement due to repeated
# measures
#-----------------------------
dboot <- dl6 %>%
  group_by(antigenf) %>%
  select(id,antigenf,seroi4fold,seror4fold,ptc,ptr) %>%
  mutate(id=as.character(id))

set.seed(123)
nreps <- 1000
ids <- unique(dboot$id)
bsamp <- matrix(sample(ids,size=length(ids)*nreps,replace=TRUE),
                nrow=length(ids),ncol=nreps)
bootests <- foreach(ab=levels(dboot$antigenf),.combine=rbind) %:%
  foreach(brep=1:nreps,.combine=rbind) %dopar% {
    di <- dboot %>% filter(antigenf==ab)
    di <- left_join(data.frame(id=bsamp[,brep]),di,by=c("id")) %>%
      select(-id) %>% group_by(antigenf) %>%
      summarize_all(function(x) sum(x,na.rm=TRUE))
    }

rate_cis <- bootests %>% 
  group_by(antigenf) %>%
  mutate(sero_c=seroi4fold/ptc,sero_r=seror4fold/ptr) %>%
  # restrict to finite values (some ETEC bootstrap replicates have a denominator of 0)
  summarize(seroi_lb = quantile(sero_c[is.finite(sero_c)],probs=c(0.025),na.rm=T),
            seroi_ub = quantile(sero_c[is.finite(sero_c)],probs=c(0.975),na.rm=T),
            seror_lb = quantile(sero_r[is.finite(sero_r)],probs=c(0.025),na.rm=T),
            seror_ub = quantile(sero_r[is.finite(sero_r)],probs=c(0.975),na.rm=T)
            )

rate_ests_4fold <- left_join(rate_ests_4fold,rate_cis,by="antigenf") %>%
  select(antigenf,ni,nit,starts_with("seroi"),nr,nrt,starts_with("seror"))


```

Confirm bootstrap distributions of the rates are smooth and reasonable
```{r boot dist figs 4fold}
bsum <- bootests %>% 
  group_by(antigenf) %>%
  mutate(sero_c=seroi4fold/ptc,sero_r=seror4fold/ptr)

ggplot(data=bsum,aes(x=sero_r)) + geom_density() + facet_wrap(~antigenf,ncol=2,scales="free") + theme_minimal()

ggplot(data=bsum,aes(x=sero_c)) + geom_density() + facet_wrap(~antigenf,ncol=2,scales="free") + theme_minimal()

```

Table of results
```{r long incidence rate table 4-fold}
rate_ests_4fold %>%
  select(antigenf,nit,ni,starts_with("seroi")) %>%
  knitr::kable(digits=2,
             caption="Incidence rates of 4-fold antibody boosting per child year",
             col.names = c("Antigen","Years at risk","n events","Incidence per year","min95","max95"),
             row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped"),full_width = TRUE)


rate_ests_4fold %>%
  select(antigenf,nrt,nr,starts_with("seror")) %>%
  knitr::kable(digits=2,
             caption="Incidence rates of 4-fold antibody waning per child year",
             col.names = c("Antigen","Years at risk","n events","Incidence per year","min95","max95"),
             row.names = FALSE) %>%
  kable_styling(bootstrap_options = c("striped"),full_width = TRUE)


```

# Table 2
Create a single table to summarize and compare incidence rates by the two methods

```{r summary table}
tab1 <- rate_ests %>%
  mutate(seroiests1 = paste(sprintf("%1.2f",seroi)," (",sprintf("%1.2f",seroi_lb),", ",sprintf("%1.2f",seroi_ub),")",sep=""),
         serorests1 = paste(sprintf("%1.2f",seror)," (",sprintf("%1.2f",seror_lb),", ",sprintf("%1.2f",seror_ub),")",sep="") ) %>%
  select(antigenf,ni1=ni,nit1=nit,seroi1=seroi,seroiests1,nr1=nr,nrt1=nrt,seror1=seror,serorests1)

tab2 <- rate_ests_4fold %>%
  mutate(seroiests2 = paste(sprintf("%1.2f",seroi)," (",sprintf("%1.2f",seroi_lb),", ",sprintf("%1.2f",seroi_ub),")",sep=""),
         serorests2 = paste(sprintf("%1.2f",seror)," (",sprintf("%1.2f",seror_lb),", ",sprintf("%1.2f",seror_ub),")",sep="") ) %>%
  select(antigenf,ni2=ni,nit2=nit,seroi2=seroi,seroiests2,nr2=nr,nrt2=nrt,seror2=seror,serorests2)

# join estimates and calculate percent change
dtab <- left_join(tab1,tab2,by=c("antigenf")) %>%
  mutate(ratioi = sprintf("%1.1f",seroi2/seroi1),
         ratioic = sprintf("%1.1f",ni2/ni1),
         ratior = sprintf("%1.1f",seror2/seror1),
         ratiorc = sprintf("%1.1f",nr2/nr1)
  )

# reshape long (stack) seroconversion and seroreversion
dtabi <- dtab %>% 
  select(antigenf,nt1=nit1,n1=ni1,rate1=seroiests1,nt2=nit2,n2=ni2,rate2=seroiests2,ratiocase=ratioic,ratiorate=ratioi) %>%
  mutate(ratedesc="Seroconversion")

dtabr <- dtab %>% 
  select(antigenf,nt1=nrt1,n1=nr1,rate1=serorests1,nt2=nrt2,n2=nr2,rate2=serorests2,ratiocase=ratiorc,ratiorate=ratior) %>%
  mutate(ratedesc="Seroreversion")

dtablong <- bind_rows(dtabi,dtabr) %>% select(ratedesc,everything())


knitr::kable(select(dtablong,-ratedesc),digits=1,
             caption="Incidence rates of seroconversion and seroreversion per child year among children ages 0-11 years in Haiti, 1990-1999. ",
             col.names = c("Pathogen","Child-years", "Incident cases","Rate (95% CI)","Child-years", "Incident cases","Rate (95% CI)","Ratio of cases","Ratio of rates")) %>%
  kable_styling(bootstrap_options = c("striped"),full_width = TRUE) %>%
  add_header_above(c(" " = 1, "Seropositivity cutoff *" = 3, "4 Fold Change in IgG levels †" = 3, " " = 2)) %>%
  group_rows("Seroconversion / boosting", 1,7) %>%
  group_rows("Seroreversion / waning", 8, 14) %>%
  footnote(symbol = c("Incident changes in serostatus defined by crossing seropositivity cutoffs.","Incident changes in serostatus defined by a 4-fold increase or decrease in IgG levels (MFI-bg), with incident boosting episodes restricted to changes that ended above the seropositivity cutoff and incident waning episodes restricted to changes that started from above the seropositivity cutoff.")
           )
```

# Compare seroincidence rates with seroprevalence
For pathogens without seroreversion, age-dependent seroprevalence is a measure of the cumulative distribution of exposure. In turn, this is a measure of the cumulative incidence of seroconversion. Since the seroreversion rate was substantial for many pathogens studied, we would hypothesize there would not necessarily be a strong relationship between seroprevalence and the average seroincidence rate over the age period.  Examine this relationship for each pathogen.  

```{r seroprevalence}
serop <- dl6 %>%
  group_by(antigenf) %>%
  summarize(seroprev=mean(seropos))

d_seropi <- left_join(rate_ests,serop,by="antigenf") %>%
  mutate(method="Seropositivity cutoff")
d_serop4fold <- left_join(rate_ests_4fold,serop,by="antigenf")  %>%
  mutate(method="4-fold increase in MFI")

d_serop <- bind_rows(d_seropi,d_serop4fold) %>%
  mutate(method=factor(method))


pcols <- c(cred,corange,cgreen,cteal,cblue,cmagent,cgrey)

pserop <- ggplot(data=d_serop,aes(x=seroprev,y=seroi,color=antigenf)) +
  geom_pointrange(aes(ymin=seroi_lb,ymax=seroi_ub),size=0.7) +
  # geom_smooth(aes(x=seroprev,y=seroi),se=FALSE,color="gray40",lwd=0.25,method="glm")+
  # geom_smooth(aes(x=seroprev,y=seroi),se=FALSE,color="gray40",lwd=0.25,lty="dashed",method="loess")+
  # geom_text(aes(label=antigenf),nudge_x = d_serop$nudge,hjust=d_serop$just,size=2)+
  facet_grid(.~method)+
  scale_x_continuous(breaks=seq(0.6,1,by=0.2),labels=seq(60,100,by=20))+
  scale_y_continuous(breaks=seq(0,2,by=0.5))+
  scale_color_manual(values=pcols,guide=guide_legend(title="Pathogen")) +
  coord_cartesian(xlim=c(0.6,1),ylim=c(0,2))+
  labs(x="Seroprevalence (%)",y="Seroincidence rate per child-year")+
  theme_minimal() 
  # theme(legend.position="none")

pserop


```
Interestingly, there is a strong relationship between the two measures, in spite of significant levels of seroreversion. This figure will be combined with results from Kenya in a separate notebook (hence, results saved in output for re-use).

# Save incidence estimates
```{r save estimates}
saveRDS(rate_ests,file=here("output","haiti-enteric-ab-ests-incidencerates.rds"))
saveRDS(rate_ests_4fold,file=here("output","haiti-enteric-ab-ests-incidencerates-4fold.rds"))
saveRDS(d_serop,file=here("output","haiti-enteric-ab-ests-seroprev-vs-rates.rds"))


```

# Sensitivity analysis of fold-changes in IgG

The alternative analysis above assumed a 4-fold change in IgG to define a change in serostatus. A 4-fold increase is a magnitude of change that vaccine immunogenicity studies often use to define seroconversion and was the basis for our choice. Below, we examine the influence of choosing fold-changes in IgG on seroconversion and seroreversion rates in the study. In all cases, seroconversion was defined as a fold-increase in IgG to above a seropositivity cutoff, and seroreversion was defined as a fold-decrease in IgG from above a seropositivity cutoff. We examined from 1 to 10-fold changes.  All estimates include bootstrapped 95% confidence intervals, as in the primary analyses above.

Regretably, the code below is fairly opaque because we have not had time to streamline it. It essentially repeats the workflow created above in each X-fold boosting or waning scenario. Since changing the fold-change definition requires re-creating the incidence dataset, combining information temporally across multiple antibody responses in some cases, it is a bit involved in each simulation scenario.

```{r fold change sensitivity analysis}

#-----------------------------
# loop over pathogens and
# fold-changes in IgG
#-----------------------------
dsens <- dl %>%
  group_by(antigenf,id) %>%
  arrange(antigenf,id,age) %>%
  mutate(nobs=n(),obsnum=row_number(),
         agediff=ifelse(obsnum==1,lead(age)-age,age-lag(age)),
         logmfi_lag  = lag(logmfi),
         logmfi_dlag  = logmfi - logmfi_lag
  ) %>%
  # Keep Norovirus separate as a pathogen
  mutate(pathogen2=as.character(pathogen),
         pathogen2=ifelse(antigenf=="Norovirus GI.4","Norovirus GI.4",pathogen2),
         pathogen2=ifelse(antigenf=="Norovirus GII.4.NO","Norovirus GII.4.NO",pathogen2)) %>%
  select(-pathogen) %>% rename(pathogen=pathogen2)


#-----------------------------
# loop over pathogens and
# fold-changes in IgG
# 
# Note: there are 7x91x1000=637,000
# bootstrap replicates in this
# so even though the estimator
# is simple it still takes a few mins
# because it has to re-create
# the incident dataset 637 times
#-----------------------------
set.seed(12)
dsens_res <- foreach(ab=unique(dsens$pathogen),.combine=rbind) %:%
  foreach(Xfold=seq(1,10,by=0.1),.combine=rbind) %dopar% {
    
    #-----------------------------
    # Identify incident boosting/waning 
    # of X-fold MFI or more. For
    # incident boosting, +X-fold increase
    # that ends above seropositivity cutoff
    # for incident waning, -X-fold decrease
    # that starts above seropositivity cutoff
    #-----------------------------
    di <- dsens %>%
      filter(pathogen==ab) %>%
      group_by(antigenf,id) %>%
      arrange(antigenf,id,obsnum) %>%
      mutate(seroi = ifelse(logmfi_dlag>log10(Xfold) & logmfi>serocut,1,0),
             seroi = ifelse(is.na(logmfi_dlag),NA,seroi),
             seror = ifelse(logmfi_dlag< -log10(Xfold) & logmfi_lag>serocut,1,0)
      ) %>%
      ungroup()
    
    #-----------------------------
    # create composite 
    # seropositivity, boosting,
    # and waning indicators
    # that use information from
    # multiple antigens if avail
    #-----------------------------
    di2 <- di %>%
      group_by(pathogen,id,obsnum) %>%
      # since incidence could fall at slightly different times across the antigens
      # for the same pathogen, use the earlier onset as the incidence measure
      arrange(pathogen,id,obsnum) %>%
      group_by(pathogen,id) %>%
      mutate(lagseroi=lag(seroi),
             seroi=ifelse(seroi==1 & lagseroi==1 & !is.na(lagseroi),0,seroi)) %>%
      # since incidence could fall at slightly different times across the antigens
      # for the same pathogen, use the later onset as the incident measure
      mutate(leadseror=lead(seror),
             seror=ifelse(seror==1 & leadseror==1 & !is.na(leadseror),0,seror)) %>%
      # now get the max for each pathogen at each timepoint
      group_by(pathogen,id,obsnum) %>%
      mutate(seropos=max(seropos),
             seroi=max(seroi),
             seror=max(seror)
      ) %>%
      slice(1) %>%
      select(pathogen,id,obsnum,agediff,seropos,seroi,seror) %>%
      # in unusual cases for multiple-antigen pathogens some children could in theory
      # boost in one antigen and wane in another. In the analysis of 4-fold changes
      # this occurred 7 times across >4,000 incident periods among
      # children with Cryptosporidium Cp17 and Cp23 antigens. In these cases,
      # recode the incident waning to 0 (assumption is that they had incident boosting)
      mutate(seror=ifelse(seroi==1 & seror==1,0,seror))
    
    #-----------------------------
    # estimate time at risk
    # for incident boosting and waning
    # assumed to be 1/2 of time
    # between measurements
    # if indivs change serostatus
    # 
    # create indicators of at risk
    # based on incidence boosting
    # and waning
    # at risk if seronegative OR experienced
    # a X-fold decrease in MFI in previous period
    #
    # the below code is opaque, but confirmed through 
    # extensive vetting and inspection that it does this:
    # children are at risk of boosting if they are seronegative or
    # if they experience a X-fold decrease in MFI between the previous
    # two measurements
    #-----------------------------
    di3 <- di2 %>%
      group_by(id) %>%
      mutate(seroi=ifelse(obsnum==1,0,seroi),
             seror=ifelse(obsnum==1,0,seror),
             maxobs=max(obsnum),
             # seronegative at obs 1: at risk
             # seropositive at obs 1: not at risk
             cntr = 0,
             cntr = ifelse(seroi==1 & lag(seroi)==0,-1,0), # counter -1 if first boost X-fold
             cntr = ifelse(obsnum==1 & seropos==0,1,cntr), # counter +1 if seroneg & obs1
             cntr = ifelse(seror==1,1,cntr), # counter +1 if seror
             cntr = ifelse(seror==1 & seropos==0 & lag(cntr)==1,cntr-1,cntr),
             
             # sum counter to help with recodes below
             cntrs = cumsum(cntr),
             
             cntr = ifelse(obsnum!=maxobs & lead(cntr)==-1 & lead(cntrs)==-1,1,cntr),
             cntr = ifelse(obsnum!=maxobs & lead(cntr)==-1 & lead(cntrs)==-2,1,cntr),
             cntr = ifelse(seror==1 & cntrs==2,0,cntr),
             
             # indicator of whether a chlid is at risk for incident boosting
             atrisk = cumsum(cntr),
             atrisk = lag(atrisk),
             
             # fix a very small number of idiosyncratic situations
             # that don't fit neatly into the above algorithm (n=5)
             atrisk = ifelse(obsnum==1,0,atrisk),
             atrisk = ifelse(obsnum > 2 & atrisk==2 | lag(atrisk==2),atrisk-1,atrisk),
             
             # now calculate person time at risk for boosting and waning
             atriskc = atrisk,
             atriskr = 1-atrisk,
             ptc = ifelse(atriskc==1,agediff,0),
             ptc = ifelse(atriskc==1 & seroi==1,agediff/2,ptc),
             ptr = ifelse(atriskr==1,agediff,0),
             ptr = ifelse(atriskr==1 & seror==1,agediff/2,ptr)
             
      )
    
    
    #-----------------------------
    # estimate incidence rates
    # for each antibody
    # units are episodes per child-year
    #-----------------------------
    dsensi_rates <- di3 %>% 
      ungroup() %>%
      summarize(ni=sum(seroi,na.rm=T),
                nit=sum(ptc,na.rm=T),
                nr=sum(seror,na.rm=T),
                nrt=sum(ptr,na.rm=T)) %>%
      mutate(seroi=ni/nit,seror=nr/nrt)
    
    #-----------------------------
    # get bootstrap CIs
    # resampling children with 
    # replacement due to repeated
    # measures
    #-----------------------------
    dboot <- di3 %>%
      ungroup() %>%
      select(id,seroi,seror,ptc,ptr)
    
    nreps <- 1000
    ids <- unique(dboot$id)
    bsamp <- matrix(sample(ids,size=length(ids)*nreps,replace=TRUE),
                    nrow=length(ids),ncol=nreps)
    dsens_bootests <- matrix(NA,nrow=nreps,ncol=2)
    i <- 1
    while(i <= nreps) {
      bi <- left_join(data.frame(id=bsamp[,i]),dboot,by=c("id")) %>%
        select(-id) %>% 
        summarize_all(function(x) sum(x,na.rm=TRUE))
      dsens_bootests[i,] <- c(bi$seroi/bi$ptc, bi$seror/bi$ptr)
      i <- i+1
    }
    
    dsensi_cis <- data.frame(bseroi=dsens_bootests[,1],bseror=dsens_bootests[,2]) %>% 
      # restrict to finite values (some ETEC bootstrap replicates have a denominator of 0)
      summarize(seroi_lb = quantile(bseroi[is.finite(bseroi)],probs=c(0.025),na.rm=T),
                seroi_ub = quantile(bseroi[is.finite(bseroi)],probs=c(0.975),na.rm=T),
                seror_lb = quantile(bseror[is.finite(bseror)],probs=c(0.025),na.rm=T),
                seror_ub = quantile(bseror[is.finite(bseror)],probs=c(0.975),na.rm=T)
      )
    
    # post results
    dsensi_res <- bind_cols(dsensi_rates,dsensi_cis) %>%
      mutate(pathogen=ab,foldchange=Xfold) %>%
      select(pathogen,foldchange,ni,nit,starts_with("seroi"),nr,nrt,starts_with("seror")) 
    dsensi_res
    

  }    

# reshape the results for ggplot
dsens_res2 <- dsens_res %>%
  mutate(antigenf = case_when(
    pathogen == "Giardia" ~ "Giardia VSP-3 or VSP-5",
    pathogen == "Cryptosporidium" ~ "Cryptosporidium Cp17 or Cp23",
    pathogen == "E. histolytica" ~ "E. histolytica LecA",
    pathogen == "Salmonella" ~ "Salmonella LPS groups B or D",
    pathogen == "ETEC" ~ "ETEC LT β subunit",
    pathogen == "Norovirus GI.4" ~ "Norovirus GI.4",
    pathogen == "Norovirus GII.4.NO" ~ "Norovirus GII.4.NO"),
    antigenf=factor(antigenf,levels=c("Giardia VSP-3 or VSP-5",
                                  "Cryptosporidium Cp17 or Cp23",
                                  "E. histolytica LecA",
                                  "Salmonella LPS groups B or D",
                                  "ETEC LT β subunit",
                                  "Norovirus GI.4",
                                  "Norovirus GII.4.NO"))
  )
dsens_res_scr <- dsens_res2 %>%
  select(antigenf,foldchange,starts_with("seroi")) %>%
  mutate(outcome="incident boosting") %>%
  rename(rate=seroi,rate_lb=seroi_lb,rate_ub=seroi_ub)
dsens_res_srr <- dsens_res2 %>%
  select(antigenf,foldchange,starts_with("seror")) %>%
  mutate(outcome="incident waning") %>%
  rename(rate=seror,rate_lb=seror_lb,rate_ub=seror_ub)
dsens_plot <- bind_rows(dsens_res_scr,dsens_res_srr) 

```

## Figure of results
```{r foldchange sens figure,fig.height=10,fig.width=6}
# color palette
pcols <- cbPalette[c(2,6)]

# rate estimates for horizontal lines
scr_plot <- rate_ests %>% mutate(outcome="incident boosting")
srr_plot <- rate_ests %>% mutate(outcome="incident waning")

dsensp <- ggplot(data=dsens_plot,aes(x=foldchange,color=outcome,fill=outcome))+
  # facet_wrap(~antigenf,ncol=2,nrow=4)+
  facet_grid(antigenf~outcome)+
  geom_vline(xintercept = 4,alpha=0.5,linetype="dashed")+
  geom_hline(data=scr_plot,aes(yintercept=seroi),linetype="dashed",alpha=0.5)+
  geom_hline(data=srr_plot,aes(yintercept=seror),linetype="dashed",alpha=0.5)+
  geom_ribbon(aes(ymin=rate_lb,ymax=rate_ub),alpha=0.2,color=NA)+
  geom_line(aes(y=rate))+
  
  scale_color_manual(values=pcols)+
  scale_fill_manual(values=pcols)+
  scale_x_continuous(breaks=1:10)+
  labs(x="fold change in IgG used to define change in serostatus",y="rate per child-year")+
  coord_cartesian(ylim=c(0,2),x=c(1,10))+
  theme_minimal() +
  theme(
    # legend.position = c(0.38,0.94),
    legend.position="none",
    legend.title=element_blank(),
    strip.text=element_text(hjust=0),
    strip.text.y=element_text(angle=0)
  )

dsensp

```

**Figure caption:** Estimates of prospective rates of incident boosting and waning by pathogen across a range of fold-changes in IgG used to define a change in serostatus. Estimates derived from 142 children ages 0-11 years in Leogane, Haiti measured prospectively over a median of 5 years, 1990-1999. Shaded regions indicate bootstrapped 95% confidence intervals.  Horizontal dashed lines mark each pathogen's seroconversion rate (boosting panels) or seroreversion rate (waning panels) estimated in the cohort using seropositivity cutoffs. Vertical dashed lines mark a 4-fold increase in IgG, commonly used to define seroconversion in challenge studies and vaccine immunogenicity studies.

The results show that incident boosting rate (force of infection) estimates were relatively insensitive to choice of fold-change in IgG in this study, particularly over a range of 2-4 fold changes. In general, incident boosting rates declined with more stringent fold-change criteria. At a 4-fold increase, the only pathogen for which the incident boosting rate was significantly different from the seroconversion rate was _Cryptosporidium_. 

# Session Info
```{r session info}
sessionInfo()
```




